### 2016-11-24 proposal/obs.tex

In set theory, a function is usually defined as a mapping that relates
each element in its domain to an element in its range. For example,
the function $a^b$ where $a$ is a positive natural number and $b$ is a
natural number denotes the following infinite set whose elements are
of the form $((a,b),a^b)$ for appropriate values of $a$ and $b$:
\[
\{ ((1,0),1), ~((1,1),1),~((2,0),1) ~\ldots~,~((2,3),8),~\ldots~ \}
\]
There is no notion of this mapping being ``calculated'' in any sense:
it just is. One ``knows'' from the mapping that $2^3$ is $8$: no time,
energy, or any other resource needs to be spent to acquire this
knowledge.

The mathematical idealization of computation founded on Turing
machines~\cite{citeulike:321509} or the
$\lambda$-calculus~\cite{Barendregt:Lambda} maintains this
understanding of computations as mathematical functions. One can use
this mathematical framework to reason about computations and as the
basis for realizing computations on actual physical machines. But
ultimately, programming practice deviates (sometimes significantly)
from the set-theoretic view of functions or the mathematical
idealization of the $\lambda$-calculus. For one thing, an infinite set
cannot be directly stored in a computer; a function is therefore
represented using rules which, given enough time, energy, memory, and
other resources, can be interpreted by a computer to generate any
particular output on demand. It is these rules that we must now
discuss in some detail. For concreteness in this section, we express
them using the notation of the Haskell programming
language~\cite{peytonjones:h98}.

As a reasonable starting point, the set-theoretic function \textit{pow}
can be realized in Haskell using the following declaration:
\begin{quote}
\begin{verbatim} 
pow :: Int -> Int -> Int
pow a 0 = 1
pow a b = a * pow a (b-1)
\end{verbatim}
\end{quote}
The code fragment declares a function called \verb|pow| that takes two values from the domain of Haskell integers \verb.Int. and returns another Haskell integer whose value is the first input raised to the power of the second input. The definition relies on several pre-defined abstractions provided by Haskell. In an appropriate environment, i.e., a computer with an installed Haskell compiler and runtime system, the function declaration can be used to calculate the value of \verb|pow 2 3|. This calculation will proceed in elementary steps that require time, energy, memory, and other computational resources on the computing platform to produce the result \verb|8|.

It is of course possible to write other definitions that realize the same function, e.g:
\begin{quote}
\begin{verbatim} 
pow' :: Int -> Int -> Int
pow' a 0 = 1
pow' a 1 = a
pow' a b | even b = pow' (a * a) (b `div` 2)
pow' a b | odd b = a * pow' a (b-1)
\end{verbatim}
\end{quote}
This definition is often much more efficient. Consider for the example the calculation of $2^{9}$ using the two approaches. The first definition induces the following sequence of elementary calculations:
\begin{quote}
\begin{verbatim} 
pow 2 9 = 2 * pow 2 8
        = 2 * (2 * pow 2 7)  
        = 2 * (2 * (2 * pow 2 6))  
        = 2 * (2 * (2 * (2 * pow 2 5)))  
        = 2 * (2 * (2 * (2 * (2 * pow 2 4)))) 
        = 2 * (2 * (2 * (2 * (2 * (2 * pow 2 3))))) 
        = 2 * (2 * (2 * (2 * (2 * (2 * (2 * pow 2 2)))))) 
        = 2 * (2 * (2 * (2 * (2 * (2 * (2 * (2 * pow 2 1))))))) 
        = 2 * (2 * (2 * (2 * (2 * (2 * (2 * (2 * (2 * pow 2 0)))))))) 
        = 2 * (2 * (2 * (2 * (2 * (2 * (2 * (2 * (2 * 1)))))))) 
        = 2 * (2 * (2 * (2 * (2 * (2 * (2 * (2 * 2)))))))
        = 2 * (2 * (2 * (2 * (2 * (2 * (2 * 4))))))
        = 2 * (2 * (2 * (2 * (2 * (2 * 8)))))
        = 2 * (2 * (2 * (2 * (2 * 16))))
        = 2 * (2 * (2 * (2 * 32))) 
        = 2 * (2 * (2 * 64))
        = 2 * (2 * 128)
        = 2 * 256
        = 512
\end{verbatim}
\end{quote}
The second instead induces the following sequence of elementary calculations:
\begin{quote}
\begin{verbatim} 
pow' 2 9 = 2 * pow' 2 8
         = 2 * pow' (2 * 2) (8 `div` 2) 
         = 2 * pow' 4 4
         = 2 * pow' (4 * 4) (4 `div` 2) 
         = 2 * pow' 16 2 
         = 2 * pow' (16 * 16) (2 `div` 2) 
         = 2 * pow' 256 1
         = 2 * 256
         = 512
\end{verbatim}
\end{quote}

Despite the mismatches between the idealized set-theoretic function
and its two Haskell realizations, it is common to consider all three
definitions to be equivalent at a certain level of
abstraction. Specifically, if we constrain the \emph{observers} to
only ask questions regarding the input-output behavior of a function
(known as the function's \emph{extensional} behavior) then all three
definitions are indeed equivalent. If we however tolerate observers
asking questions regarding timing, memory usage, battery consumption,
or even electromagnetic radiation then the situation is different. For
the set-theoretic function, these observations are non-sensical. For
the Haskell functions, we have for example that one execution of
\verb|pow 2 10000| took 27.31109 ms on a modern laptop while an
execution of \verb|pow' 2 10000| under similar circumstances only took
11.72614 ms, less than half the time. The two functions \verb|pow| and
\verb|pow'| can now be distinguished by observations.

\paragraph*{Observational Equivalence.} There is a well-established
theory that formalizes equivalence of programs up to
observations~\cite{Barendregt:Lambda,Milner:1999:CMS:329902}. In the
conventional application of this theory, the only allowed observations
are the \emph{extensional} ones, i.e., the input-output relations and
other ``semantic'' communication among agents. According to this
conventional situation, the functions \verb|pow| and \verb|pow'| would
be equivalent. As we have seen, by allowing observations of timing
behavior, the two functions \verb|pow| and \verb|pow'| can be
distinguished but formalizing this appears difficult: the precise
timing depends on the particular compiler, runtime system, processor
model, available memory, the resolution of the hardware clock, the
operating system services that manage the hardware clock, and even the
ambient temperature of the room in which the experiment is set up. In
more detail, we know that in order to measure the time elapsed during
the execution of a program, the operating system must execute
additional code that itself takes time to execute and hence distorts
the results to some extent. Moreover the precision of the timing
results is not arbitrary: it is limited by the precision of the
hardware clock, the speed of the code that must execute to record the
timing, the exact memory locations in which the timing results are
stored, and the speed of common operations like addition and
subtraction that must be performed to calculate the elapsed time. In a
\emph{classical worldview}, all these factors are acknowledged to
exist but are not formalized mathematically. Instead, the experimenter
takes great care to minimize their impact by striving to communicate
reproducible results consisting of mean execution times under
carefully managed conditions. Thus, the precise timings are treated as
an objective (ontic) property of the functions and measurements are
viewed as operations that simply reveal such properties with minimal
disturbance.