### 2016-11-24 proposal/collab.tex

\documentclass[11pt]{article}
\usepackage{fullpage,times,url}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\Large {\bf Collaborators}}
\end{center}

\begin{itemize}
\item Recent collaborators:
  Thorsten Altenkirch (University of Nottingham),
  Zena M. Ariola (University of Oregon),
  Aaron Bohannon (University of Pennsylvania),
  Jacques Carette (McMaster University),
  Ant\^onio Carlos da Rocha Costa (Univ. Cat\'olica de Pelotas, Brazil),
  Andr\'e Rauber Du Bois (Heriot-Watt University),
  Ronald Garcia (University of British Columbia),
  Jonathan Grattage (Universit\'e de Grenoble),
  Hugo Herbelin (INRIA, Futurs),
  Oleg Kiselyov,
  Giovani Rubert Librelotto (Universidade Federal de Santa Maria),
  Andrew Lumsdaine (Indiana University),
  Chung-chieh Shan (Rutgers),    
  Juliana Kaizer Vizzotto (Universidade Federal de Santa Maria).

\item Graduate and Postdoctoral Advisors: Matthias Felleisen (Northeastern
  University), Corky Cartwright (Rice University), and John Hughes (Chalmers
  University, Sweden)

\item Thesis Advisor and Postgraduate-Scholar Sponsor: 
    Roshan James (Jane Street) 
    John Fiskio-Lasseter (Lewis and Clark College), 
    Yevgeniy Makarov,
    Bart Massey (Portland State University),
    Juliana Kaizer Vizzotto (Universidade Federal de Santa Maria),
    Bennett Yates.

\end{itemize}

\end{document}