### 2016-11-25 proposal/amr-bio.tex
amr-bio.tex Finish putting everything into bibLaTeX...

\documentclass[11pt]{article}
\usepackage{fullpage,times,url}
\usepackage[utf8]{inputenc}
\usepackage[ backend=biber]{biblatex}
\addbibresource{prop.bib}
\ExecuteBibliographyOptions{sorting=none,maxbibnames=5,doi=false,isbn=false,url=false}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\Large {\bf Biographical Sketch for Amr Sabry}}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Professional Preparation}

\begin{itemize}
\item Cairo University, Egypt, Electronics \& Computer Engineering, B.Sc.
  (1986)
\item Cairo University, Egypt, Computer Engineering, M.Sc. (1989)
\item Rice University, Computer Science, M.Sc. (1993) 
\item Rice University, Computer Science, Ph.D. (1994) 
\item Chalmers University, Sweden, Programming Languages, Postdoctoral
  researcher (1994-1995)
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Appointments}

\begin{itemize}
\item{2007-Present} Professor, School of Informatics and Computing,
Indiana University.
\item{2000-2007} Associate Professor, Computer Science Department,
Indiana University.
\item{Spring 2004} Visiting Professor, Genova University, Italy.
\item{Fall 2003} Visiting Researcher, Microsoft Research, Cambridge, UK.
\item{1995-2000} Assistant Professor, Department of Computer and
Information Science, University of Oregon.
\item{1994-1995} Postdoctoral Researcher, Department of Computer Science,
  Chalmers University, G\"oteborg, Sweden.
\item{1993-1994} Research Assistant, Rice University, visiting Carnegie 
Mellon University.
\item{1989-1993} Research and Teaching Assistant, Rice University.
\item{1987} Teaching Assistant, Polytechnic Military School, Egypt. 
\item{1986-1989} Research and Teaching Assistant, Cairo University, Egypt.
\item{1986-1989} Programmer and Consultant, Pacer, Egypt.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{refsection}
\nocite{DQT2014, geometry2013, VDS2009, AGVS2005, VAS2006}
\printbibliography[title={Closely Related Publications}]
\end{refsection}

\begin{refsection}
\nocite{CaretteSabry2016, JamesSabry2012, Sabry2009, MoggiSabry2001, FSDF1993}
\printbibliography[title={Other Significant Publications}]
\end{refsection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Synergetic Activities}

Paul Purdom and Amr Sabry, CNF Generator for Factoring Problems.  Software
system to generate satisfiability problems used in the SAT competition
organized in conjunction with the Sixth International Conference on the
Theory and Applications of Satisfiability Testing, March 2003, and
continually used by designers of SAT solvers worldwide. Available online at
\url{http://www.cs.indiana.edu/cgi-pub/sabry/cnf.html}.

\end{document}