### 2016-11-30 proposal/fac.tex
done

\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{amsfonts, verbatim}
\usepackage{proof}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{comment}
\usepackage{natbib}
\usepackage{url}
\usepackage{cmll}
\usepackage{bussproofs}
\usepackage{bbold}
\bibpunct();A{},
\bibliographystyle{plainnat}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Environments

\newenvironment{code}{\verbatim}{\endverbatim}
\newenvironment{smallcode}{\verbatim}{\endverbatim}
\newenvironment{verysmallcode}{\scriptsize\verbatim}{\endverbatim\normalsize}
\newtheorem{thm}{Theorem}[section]
\newtheorem{definition}[thm]{Definition}
\newtheorem{proposition}[thm]{Proposition}
\newtheorem{example}[thm]{Example}

%% Commands

\newcommand{\hide}[1]{}
\newcommand{\zeroH}{\mathbb{0}}
\newcommand{\oneH}{\mathbb{1}}
\newcommand{\alt}{~|~} 
\newcommand{\liftt}[1]{\llbracket #1 \rrbracket}
\newcommand{\ie}{\textit{i.e.}\xspace}
\newcommand{\eg}{\textit{e.g.}\xspace}
\newcommand{\QM}{QM}
\newcommand{\blob}{\rule[.2ex]{1ex}{1ex}}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ip}[2]{\langle #1 \alt #2 \rangle}
\newcommand{\op}[2]{|#1\rangle\langle #2|}
\newcommand{\expect}[2]{\langle #1 \alt #2 \alt #1 \rangle}
\newcommand{\Nat}{\mathbb{N}}
\newcommand{\inleft}[1]{\textsf{inL}~#1}
\newcommand{\inright}[1]{\textsf{inR}~#1}
\newcommand{\fst}[1]{\textsf{fst}~#1}
\newcommand{\snd}[1]{\textsf{snd}~#1}
\newcommand{\scalarZ}{\texttt{impossible}}
\newcommand{\scalarO}{\texttt{possible}}
\newcommand{\norm}[1]{||#1||}
\newcommand{\scalarPlus}{\ensuremath{\oplus}}
\newcommand{\scalarTimes}{\ensuremath{\wedge}}
\newcommand{\vect}[1]{\overline{#1}}
\newcommand{\dvect}[1]{\overline{\overline{#1}}}
\newcommand{\ltree}{\swarrow}
\newcommand{\rrtree}{\begin{array}{c@{\hspace{-0pt}}c}\searrow\\&\searrow\end{array}}
\newcommand{\rltree}{\begin{array}{c@{\hspace{-7pt}}c}\searrow\\&\swarrow\end{array}}

\newcommand{\todo}[1]{\textbf{TODO}: #1}
\newcommand{\note}[1]{\textsc{\textbf{NOTE: #1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Facilities, Equipment, and Other Resources at Indiana University}
\author{}
\date{}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\noindent There are substantial central computing resources offered by
the University: our cyberinfrastructure is rated in the top 10 in the
U.S, and includes the Big Red cluster (30 TeraFlops), the Quarry
cluster (9 TeraFlops, general-purpose Unix computing), and supports
TeraGrid, I-Light, and the Open Science Grid. Dr. Ortiz (Department of
Physics) and Dr. Sabry and Dr. Hanson (School of Informatics and
Computing) have access to departmental license software, as well as
access to most electronic and hard copy journals through Indiana
University Libraries, and access to conference facilities for
meetings.

\bigskip
\centerline{\Large{\underline{Department of Physics}}}
\bigskip

\noindent A computer server facility specifically serving the
Departments of Physics and Astronomy is installed in the basement of
Swain Hall. This $\sim 110~m^2$ room has sufficient power and cooling
capacity to house 24 racks of hardware in a single area for ease of
security and maintenance resulting in a high-quality, small-scale
facility meeting local needs. In addition to the ATLAS/D0 Tier-3
Analysis Cluster (22 eight-core 2.66/2.88 GHz Xeon nodes, $>$ 60 TB of
raid disk storage), it also houses experimental computing for lattice
gauge and related calculations, the special partial-wave computing
facility for the GlueX experiment (including both regular nodes and an
experimental GPU cluster), computing access for the neutrino group,
and servers for the Physics and Astronomy Departments. Finally, it
also provides an appropriate home for interfaces to hardware such as
IU's Data Capacitor, a large-scale data storage/access system
developed by Indiana University IT Services and used heavily by
Astronomy.  Indiana University and the University of Chicago jointly
manage the Midwest Tier2 (MWT2) computing center, with hardware split
approximately equally between them. The Indiana University portion of
the MWT2 is located primarily in Indianapolis on the campus of Indiana
University Purdue University Indianapolis (IUPUI). Currently the IUPUI
portion of the MWT2 has 1.8M kSi2k of computing power and $\sim$ 200
TB of storage. The CPUs are various AMD and Intel processors purchased
from several vendors, a total of 768 cores at IUPUI with approximately
2 GB of memory per core. A 1 Gb/s link connects each compute node to
the Indiana University backbone while storage nodes are connected to
the backbone with 10 Gb/s connections. The MWT2 is connected to the
outside world with a 10 Gb/s link.  In addition to the usual desktop
units, a dedicated room is available containing projection units and a
high-definition Polycom videoconferencing unit recently acquired using
DoE stimulus funds. The Department has also requested renovation of a
different space to house this system as well as other equipment for
use in distance learning.

\bigskip
\centerline{\Large{\underline{School of Informatics and Computing}}}
\bigskip

\noindent The School of Informatics and Computing maintains a highly
distributed, heterogeneous, networked computing environment consisting of
UNIX/Linux, Windows, and Macintosh workstations and servers across 3 primary
facilities (2 auxiliary). This includes over 250 Linux workstations/servers,
over 450 Windows workstations/servers and over 150 Apple
workstations/servers. There are 3 Linux clusters (128-node dual Opteron,
16-node dual Opteron, and 8-node dual Xeon) to support School
research. Additionally, there are 2 dedicated virtual machine servers
available to host VMs for both research and instruction.

The network infrastructure at the School provides 1000Mbps gigabit ethernet
connections to all servers and most workstations and as well as 802.11a/b/g
wireless connectivity. The facilities have 1 gigabit or greater connectivity
(one facility has 10 gigabit) with campus and research backbones which
provides high-speed access to university and worldwide computing resources.

Special-purpose teaching and instructional labs are provided, including 2
electronic classrooms (24-seat, 48-seat), hardware laboratories, and 4
multi-purpose Linux and Windows laboratories to support the course work for
both undergraduate and graduate students.  There are 4 rooms (2 classrooms)
within the School enabled with video conferencing technology that can be used
for both distance education and research collaboration.

The School maintains 10 full time staff members dedicated to IT
infrastructure and facilities with the mission of enabling world class
research and instruction.

\end{document}