### 2017-10-29 prop_iop.tex
prop_letter.tex \section{The Kochen-Specker Theorem and Contextuality} Updating...

%% LyX 2.2.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt]{iopart}
\usepackage[T1,T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{array}
\usepackage{verbatim}
\usepackage{booktabs}
\usepackage{mathrsfs}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{amstext}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\usepackage{iopams}
\usepackage{setstack}
 \newlist{casenv}{enumerate}{4}
 \setlist[casenv]{leftmargin=*,align=left,widest={iiii}}
 \setlist[casenv,1]{label={{\itshape\ \casename} \arabic*.},ref=\arabic*}
 \setlist[casenv,2]{label={{\itshape\ \casename} \roman*.},ref=\roman*}
 \setlist[casenv,3]{label={{\itshape\ \casename\ \alph*.}},ref=\alph*}
 \setlist[casenv,4]{label={{\itshape\ \casename} \arabic*.},ref=\arabic*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage[ backend=biber]{biblatex}
\addbibresource{prop.bib}
\ExecuteBibliographyOptions{sorting=none,maxbibnames=10,doi=false,isbn=false,url=false}
\DeclareUnicodeCharacter{0229}{\c{e}}
\usepackage{mmap}
\usepackage{fullpage}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{bbold}
\usepackage{wesa}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{iopams}
\usepackage{setstack}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}
\newtheorem{lemma}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[thm]{Definition}
\newtheorem{example}[thm]{Example}
\theoremstyle{remark}
\newtheorem{claim}[thm]{Claim}

\newcommand{\Hilb}{\mathcal{H}}
\newcommand{\events}{\ensuremath{\mathcal{E}}}
\newcommand{\qevents}{\ensuremath{\mathcal{E}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\imposs}{{\mbox{\wesa{impossible}}}}
\newcommand{\likely}{{\mbox{\wesa{likely}}}}
\newcommand{\unlikely}{{\mbox{\wesa{unlikely}}}}
\newcommand{\necess}{{\mbox{\wesa{certain}}}}
\newcommand{\unknown}{{\mbox{\wesa{unknown}}}}
\newcommand{\midd}{{\mbox{\wesa{middle}}}}
\usepackage[braket,qm]{qcircuit}
\newcommand{\fket}[1]{{|#1\rangle}}
\newcommand{\fproj}[1]{|#1\rangle\langle #1|}
\newcommand{\proj}[1]{\op{#1}{#1}}
\newcommand{\ps}{\texttt{+}}
\newcommand{\minus}{\texttt{-}}

\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\newcommand{\amr}[1]{\fbox{\begin{minipage}{0.9\textwidth}\color{green}{Amr says: #1}\end{minipage}}}
\newcommand{\yutsung}[1]{\fbox{\begin{minipage}{0.9\textwidth}\color{purple}{Yu-Tsung says: #1}\end{minipage}}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}~\middle|~{#2}\right\} }}
\def\C{{\mathbb{C}}}
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{\mathrm{L{#1}}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{\mathrm{R{#1}}}}}
% https://tex.stackexchange.com/questions/345284/can-i-have-an-almost-non-breaking-space-in-latex
% Change non-breaking space to \nolinebreak[1] 
% Hence, it will not hyphenate too much surrounding words.
\newcommand{\nb}{\nolinebreak[3] }
\newcommand{\iotal}[1][]{\ensuremath{\iota^{\mathrm{L{#1}}}}}
\newcommand{\iotar}[1][]{\ensuremath{\iota^{\mathrm{R{#1}}}}}
\renewcommand{\maketitle}{}
\newcommand{\tfrac}[2]{\textstyle\frac{#1}{#2}}

\makeatother

 \providecommand{\casename}{Case}

\begin{document}

\title{Quantum Interval-Valued Probability: Measurement and Probability}

\author{Andrew J Hanson$^{1}$, Gerardo Ortiz$^{2}$, Amr Sabry$^{1}$ and
Yu-Tsung Tai$^{3,1}$}

\address{$^{1}$ School of Informatics and Computing, Indiana University,
Bloomington, IN 47405, USA}

\address{$^{2}$ Department of Physics, Indiana University, Bloomington, IN
47405, USA}

\address{$^{3}$ Department of Mathematics, Indiana University, Bloomington,
IN 47405, USA}

\ead{ortizg@indiana.edu}
\begin{abstract}
\today 
\end{abstract}

\pacs{03.65.-w, 03.65.Ta, 02.40.Dr, 02.50.Cw, 02.50.Le}

\noindent{\it Keywords\/}: {\noindent Finite precision measurement, the Born rule, Gleason's
theorem, interval-valued probability, discrete quantum theory}

\submitto{\jpa}

\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

Textbook quantum mechanics asserts that, given a quantum spin system
described by the wave function $\ket{\psi}=\frac{\sqrt{\pi}}{2}\fket{\ps\frac{1}{2}}+\frac{\sqrt{4-\pi}}{2}\fket{\minus\frac{1}{2}}$
and given the observable $S_{z}=\frac{\hbar}{2}(\fproj{\ps\frac{1}{2}}-\fproj{\minus\frac{1}{2}})$,
then the probability of observing the eigenvalue $\ps\frac{\hbar}{2}$
(with post-measurement state $\fket{\ps\frac{1}{2}}$) is $\frac{\pi}{4}$
and the probability of observing the eigenvalue $\minus\frac{\hbar}{2}$
(with post-measurement state $\fket{\minus\frac{1}{2}}$) is $1-\frac{\pi}{4}$.
According to the frequency interpretation of probability, the above
assertion implies that when performing a very large number of experiments
on quantum systems described by wave function $\ket{\psi}$, a fraction
$\frac{\pi}{4}$ of the experiments will result in the observation
$\ps\frac{\hbar}{2}$ and the remaining fraction $1-\frac{\pi}{4}$
will result in the observation $\minus\frac{\hbar}{2}$.

In reality, i.e., in an actual experiment, simulation, or computation,
bounded by space, time, and energy resources, the situation is much
fuzzier: (i) we cannot \emph{exactly} prepare the state $\ket{\psi}$;
(ii) we cannot build a measurement apparatus corresponding \emph{exactly}
to observable $S_{z}$; and (iii) we cannot prepare \emph{exact} replicas
of any state. In other words, we can only perform a limited number
of experiments on quantum systems closely related to wave function
$\ket{\psi}$ and using an apparatus that closely approximates observable
$S_{z}$. Given any fixed resources, these experiments will \emph{not}
result in the \emph{exact} fraction $\frac{\pi}{4}$ of observations
$\ps\frac{\hbar}{2}$ and $1-\frac{\pi}{4}$ of observations $\minus\frac{\hbar}{2}$.
In fact, these experiments can be viewed as a method (algorithm) to
calculate the value $\pi$ and it is known that the state of the art
algorithms for computing the $n$th binary digit of~$\pi$ require
on the order of $O\left(n\log^{O\left(1\right)}\left(n\right)\right)$
operations~\cite{journals/moc/BaileyBP97}. In other words, calculating
the $n$th digit of $\pi$ appears to require more and more physical
resources as~$n$ gets larger.

All of this is well-known and holds in both the classical and quantum
worlds. In the classical world, these observations are believed to
be well-understood approximations to an idealized independent reality
that can continuously approached with more and more resources. The
question in the quantum world is more subtle; it is not evident at
all that this resource-aware perspective does not alter the very foundations
of quantum mechanics. Indeed taking such resource bounds into consideration
is what founded computer science as a discipline and has become crucial
for understanding the very nature of computation. Following Feynman~\cite{Feynman1982Simulating},
Landauer~\cite{Landauer1996188}, and others, one might argue that
this resource-bounded perspective is also crucial for understanding
the very nature of physical processes.

In previous work, we analyzed the wave function description of quantum
systems from a resource-bounded perspective. Technically, instead
of postulating that quantum states are rays in a Hilbert space, i.e.,
we instead used special finite fields whose size reflected the available
resources. \ldots{}

\yutsung{If the core is non-empty, could the core characterize $\bar{\mu}$
comprehensively?}

\newpage{}

This is a \emph{theoretical} investigation of \emph{experimental}
physics using \emph{computational} methods. All experiments and computations
are processes bounded in space, time, energy, and other resources~\cite{Jaeger2007,Piccinini2015}.
Yet, for centuries, the mathematical formalization of such processes
has been founded on the infinitely precise real or complex numbers~\cite{Ziegler2007,weihrauch2012computable,blum2012complexity}.
Our purpose here is to exploit the consequences of replacing infinitely-specified
quantum probabilities by finite number of intervals used in interval-valued
probability measures~(IVPM); in particular, we show that a quantum
IVPM may not be induced by a infinitely-precise state, but it might
be more and more likely to be identified to a particular state when
the measurement resources increase.

Indeed, almost every description of quantum mechanics, quantum computation,
or quantum experiments refers to entities such as $\rme$, $\pi$,
$\sqrt{2}$, etc (see, e.g., \cite{Redhead1987-REDINA,544199,Mermin2007}).
From a computational perspective, such numbers do not exist in their
entirety ``for free~\cite{Kent1999,CliftonKent2000}.'' For example,
the state of the art algorithms for computing the $n$th binary digit
of~$\pi$ require on the order of $O\left(n\log^{O\left(1\right)}\left(n\right)\right)$
operations~\cite{journals/moc/BaileyBP97}. In other words, simply
referring to the $n$th digit of $\pi$ requires more and more resources
as $n$ gets larger. Taking such resource bounds into consideration
is what founded computer science as a discipline and is crucial for
understanding the very nature of computation and, following Feynman~\cite{Feynman1982Simulating},
Landauer~\cite{Landauer1996188}, and others, for understanding the
very nature of physical processes.

We have been revisiting quantum mechanics, quantum information, and
quantum computation from this resource-aware perspective. Our initial
results in that domain showed how subtle the issues can be~\cite{usat,geometry2013,DQT2014}:
a straightforward replacement of the complex numbers by a finite field
yields a variant of quantum mechanics in which computationally hard
problems like UNIQUE-SAT (which decides whether a given Boolean formula
is unsatisfiable or has exactly one satisfying assignment~\cite{Valiant198685,Papadimitriou1993,AroraBarak2009})
can be deterministically solved in constant time. To eliminate such
unrealistic theories requires delicate analysis of the structure of
the Hilbert space, the process of observation, and the notion of probability
teasing apart their reliance on the infinitely precise real numbers~\cite{geometry2013,DQT2014}.

In this paper, we shift focus from the infinitely-specified but not
directly observable quantum states, to observable measurable properties
of quantum systems and their probabilities. Furthermore, we insist
that our theories of measurement and probability only refer to finitely
communicable evidence within feasible computational bounds. It follows
that states, observations, and probabilities all become ``fuzzy'',
i.e., specified by intervals of confidence that can only increase
in precision if the available resources increase proportionally. Our
notion of ``fuzzy quantum mechanics'' is related to existing work~\cite{GranikCaulfield1996,Pykacz2013,SNL2009,Gudder2005,aerts1993physical}
but, as will be explained in more detail, is distinguished by its
unique computational character.

We will begin by reviewing existing work that recasts classical probability
spaces in a resource-aware setting and move to aim at recasting quantum
probability and quantum measurement. In particular, we develop a measurement
framework based on quantum IVPM. Surprisingly, we found a quantum
IVPM which can not be induced by a state, while Shapley proved a classical
convex IVPM can always be induced by a classical ``state''~\cite{Shapley1971,Grabisch2016},
and Gleason proved a infinitely-specified quantum probabilities measure
can always be induced by a quantum state~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.
Nevertheless, a quantum IVPM could still correspond to many possible
states if it is broken into pieces. Our examples also suggest that
a quantum IVPM could be more and more likely induced by a state while
the intervals become sharper, i.e., the measurement resources increase.

Since Gleason's theorem supports the idea of von Neumann to define
a state as a density matrix~\cite{Varadarajan2008}, a density matrix
might not be a adequate imprecise quantum state. Instead, a imprecise
quantum state might be more like an interactive system suggested by
Quantum Bayesianism or QBism~\cite{FuchsSchack2013,FuchsMerminSchack2014,VonBaeyer2016},
which would interact differently with different clients. Furthermore,
the validity the fundamental theorems of quantum mechanics followed
up by Gleason's theorem, such as Bell~\cite{Bell1964,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
and Kochen-Specker~\cite{kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007},
might need to be reassessed based on our imprecise measurement. Our
research then might provide a new insight to the old debate among
Meyer, Mermin, and others about whether finite precision measurement
would nullify the Kochen-Specker theorem or not~\cite{PhysRevLett.83.3751,Mermin1999,BarrettKent2004}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Classical Probability}

A \emph{probability space} specifies the necessary conditions for
reasoning coherently about collections of uncertain events~\cite{Kolmogorov1950,Shafer1976,Griffiths2003,Swart2013}.
We review the conventional presentation of probability spaces and
then discuss the computational resources needed to estimate probabilities.

%%%

\subsection{Classical Probability Spaces}

The conventional definition of a probability space builds upon the
field of real numbers. In more detail, a probability space consists
of a \emph{sample space} $\Omega$, a space of \emph{events}~$\events$,
and a \emph{probability measure}~$\pmeas$ mapping events in $\events$
to the real interval $[0,1]$. We will only consider \emph{finite}
sets of events and restrict our attention to non-empty finite sets
$\Omega$ as the sample space. The space of events $\events$ includes
every possible subset of $\Omega$: it is the powerset~$2^{\Omega}=\set{E}{E\subseteq\Omega}$.
For future reference, we emphasize that events are the primary notion
of interest and that the sample space is a convenient artifact that
allows us to treat events as sets obeying the laws of Boolean algebra~\cite{Boole1948,Redhead1987-REDINA,Griffiths2003}.

\begin{definition}[Probability Measure]\label{def:ClassicalProbabilitySpace}
Given the set of events $\events$, a \emph{probability measure} is
a function $\pmeas:\events\rightarrow[0,1]$ such that: 
\begin{itemize}
\item $\pmeas(\emptyset)=0$, 
\item $\pmeas(\Omega)=1$, 
\item for every event $E$, $\pmeas\left(\Omega\backslash E\right)=1-\pmeas\left(E\right)$
where $\Omega\backslash E$ is the complement event of $E$, and 
\item for every collection $\left\{ E_{i}\right\} _{i=1}^{N}$ of pairwise
disjoint events, $\pmeas\left(\bigcup_{i=1}^{N}E_{i}\right)=\sum_{i=1}^{N}\pmeas(E_{i})$. 
\end{itemize}
\end{definition} There is some redundancy in the definition that
will be useful when moving to quantum probability spaces.

\noindent \begin{example}[Two-coins experiment]\label{ex1} Consider
an experiment that tosses two coins. We have four possible outcomes
that constitute the sample space $\Omega=\{HH,HT,TH,TT\}$. There
are 16 total events including the event $\{HH,HT\}$ that the first
coin lands heads up, the event $\{HT,TH\}$ that the two coins land
on opposite sides, and the event $\{HT,TH,TT\}$ that at least one
coin lands tails up. Here is a possible probability measure for these
events: 
\begin{equation}
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) & = & 0\\
\pmeas(\{HH\}) & = & \frac{1}{3}\\
\pmeas(\{HT\}) & = & 0\\
\pmeas(\{TH\}) & = & \frac{2}{3}\\
\pmeas(\{TT\}) & = & 0\\
\pmeas(\{HH,HT\}) & = & \frac{1}{3}\\
\pmeas(\{HH,TH\}) & = & 1\\
\pmeas(\{HH,TT\}) & = & \frac{1}{3}
\end{array} & \begin{array}{rcl}
\pmeas(\{HT,TH\}) & = & \frac{2}{3}\\
\pmeas(\{HT,TT\}) & = & 0\\
\pmeas(\{TH,TT\}) & = & \frac{2}{3}\\
\pmeas(\{HH,HT,TH\}) & = & 1\\
\pmeas(\{HH,HT,TT\}) & = & \frac{1}{3}\\
\pmeas(\{HH,TH,TT\}) & = & 1\\
\pmeas(\{HT,TH,TT\}) & = & \frac{2}{3}\\
\pmeas(\{HH,HT,TH,TT\}) & = & 1
\end{array}\end{array}
\end{equation}
\end{example}

\noindent It is useful to think that this probability measure is completely
determined by the ``reality'' of the two coins in question and their
characteristics, in the sense that each pair of coins induces a measure,
and each measure must correspond to some pair of coins. The measure
above would be induced by two particular coins such that the first
coin is twice as likely to land tails up than heads up and the second
coin is double-headed.

%%%%%

\subsection{Measuring Probabilities: Buffon's Needle Problem\label{subsec:Measuring-Probabilities:-Buffon}}

In a strict computational or experimental setting, one should question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval~$[0,1]$~\cite{Turing_1937,Ziegler2007,weihrauch2012computable}.
This interval includes numbers like $0.h_{1}h_{2}h_{3}\ldots$ where
$h_{i}$ is 1 or 0 depending on whether Turing machine $\mathit{TM}_{i}$
halts or not. Such numbers cannot be computed. This interval also
includes numbers like $\pi/4$ which can only be computed with increasingly
large resources as the precision increases. Therefore, in a resource-aware
computational or experimental setting, it is more appropriate to consider
probability measures that map events to a set of elements computable
with a fixed set of resources. We expand on this observation in the
next section and then recall its formalization using interval-valued
probability measures~\cite{Weichselberger2000,JamisonLodwick2004}.

In the previous example, we assumed the probability~$\pmeas(E)$
of each event~$E$ is known a priori. In reality, although each event
is assumed to have a probability, the exact value of $\pmeas(E)$
may not be known. According to the \emph{frequency interpretation
of probability} (which we will revisit when moving to the quantum
case)~\cite{Venn1876,Hajek2012}, to determine the probability of
an event, we run $M$ independent trials which gives us an approximation
of the (assumed) ``true'' or ``real'' probability. Let $x_{i}$
be 1 or 0 depending on whether the event~$E$ occurs in the $i$-th
trial or not, then $\pmeas(E)$ could be approximated to given accuracy~$\epsilon>0$
by the relative frequency~$\sum_{i=1}^{M}x_{i}/M$ with the probability
converging to one as $M$ goes to infinity, i.e., 
\begin{equation}
\forall\epsilon>0,\lim_{M\rightarrow\infty}\pmeas\left(\left|\pmeas(E)-\frac{1}{M}\sum_{i=1}^{M}x_{i}\right|<\epsilon\right)=1\,.
\end{equation}
This fact is called the law of large numbers~\cite{Bernoulli2006,Kolmogorov1950,Uspensky1937,Shafer1976,544199}.

Let's look at a concrete example. Suppose we drop a needle of length
$\ell$ onto a floor made of equally spaced parallel lines a distance
$h$ apart, where $\ell<h$. It is a known fact that the probability
of the needle crossing a line is $2\ell/\left(\pi h\right)$~\cite{Buffon1777,DeMorgan1872,Hall1873,Uspensky1937}.
Consider an experimental setup consisting of a collection of $M$
identical needles of length $\ell$. We throw the $M$ needles one
needle at a time, and observe the number $M_{c}$ of needles that
cross a line, thus estimating the probability of a needle crossing
a line to be $M_{c}/M$. In an actual experiment with $500$ needles
and the ratio $\ell/h=0.75$~\cite{Hall1873}, it was found that
$236$ crossed a line so the relative frequency is $0.472$ whereas
the idealized mathematical probability is $0.4774\ldots$. In a larger
experiment with $5000$ needles and the ratio $\ell/h=0.8$~\cite{Uspensky1937},
the relative frequency was calculated to be $0.5064$ whereas the
idealized mathematical probability is $0.5092\ldots$. We see that
the observed probability approaches $2\ell/\left(\pi h\right)$ but
only if \emph{larger and larger resources} are expended. These resource
considerations suggest that it is possible to replace the real interval
$[0,1]$ with rational numbers up to a certain precision related to
the particular experiment in question. There is clearly a connection
between the number of needles and the achievable precision: in the
hypothetical experiment with 3 needles, it is not sensible to retain
100 digits in the expansion of $2\ell/\left(\pi h\right)$.

There is another more subtle assumption of unbounded computational
power in the experiment. We are assuming that we can always determine
with certainty whether a needle is crossing a line. But ``lines''
on the the floor have thickness, their distance apart is not exactly
$h$, and the needles' lengths are not all absolutely equal to $\ell$.
These perturbations make the events ``fuzzy.'' Thus, in an experiment
with limited resources, it is not possible to talk about the idealized
event that exactly $M_{c}$ needles cross lines as this would require
the most expensive needles built to the most precise accuracy, laser
precision for drawing lines on the floor, and the most powerful microscopes
to determine if a needle does cross a line. Instead we might talk
about the event that $M_{c}-\delta$ needles evidently cross lines
and $M_{c}+\delta'$ needles plausibly cross lines where $\delta$
and $\delta'$ are resource-dependent approximations. This fuzzy notion
of events leads to probabilities being only calculable within intervals
of confidence reflecting the certainty of events and their plausibility.
This is indeed consistent with published experiments: in an experiment
with $3204$ needles and the ratio $\ell/h=0.6$~\cite{DeMorgan1872},
$1213$ needles clearly crossed a line and $11$ needles were close
enough to plausibly be considered as crossing the line: we would express
the probability in this case as the interval $\left[\frac{1213}{3204},\frac{1224}{3204}\right]$
expressing that we are certain that the event has probability at least
$\frac{1213}{3204}$ but it is possible that it would have probability
$\frac{1224}{3204}$.

Another way to re-express the above observations is that the calculus
of probabilities requires all probabilities to be \emph{coherent}
(e.g., the probability of the union of two disjoint events is the
sum of their probabilities). But given limited resources, the calculation
of a probability measure can only be approximate and hence the probabilities
cannot be absolutely guaranteed to be coherent in the ideal sense.

%%%%%

\subsection{Classical Interval-Valued Probability Measures}

As motivated above, an event $E_{1}$ may have an interval of probability
$[l_{1},r_{1}]$. Assume that another disjoint event $E_{2}$ has
an interval of probability $[l_{2},r_{2}]$, what is the interval
of probability for the event $E_{1}\cup E_{2}$? The answer is somewhat
subtle: although it is possible to use the sum of the intervals $[l_{1}+l_{2},r_{1}+r_{2}]$
as the combined probability, one can find a much tighter interval
if information \emph{against} the event (i.e., information about the
complement event) is also taken into consideration. Formally, for
a general event $E$ with probability $[l,r]$, the evidence that
contradicts $E$ is evidence supporting the complement of $E$. The
complement of $E$ must therefore have probability $\left[1-r,1-l\right]$
which we abbreviate $\left[1,1\right]-\left[l,r\right]$. In general,
the scalar multiplication of an interval is defined by 
\begin{equation} 
x[l,r]=\cases{ 
[xl,xr] & for $x\ge0$ ;\\
[xr,xl] & for $x\le0$ .
}
\end{equation}
Given a sample space~$\Omega$, its set of events~$\events$, and
a collection of intervals $\mathscr{I}$, a function~$\bar{\mu}:\events\rightarrow\mathscr{I}$
is a classical interval-valued probability measure if and only if
$\bar{\mu}$ satisfies the following conditions~\cite{JamisonLodwick2004}
where the last line uses $\subseteq$ to allow for tighter intervals
that exploit the complement event: 
\begin{itemize}
\item $\bar{\mu}(\emptyset)=[0,0]$. 
\item $\bar{\mu}(\Omega)=[1,1]$. 
\item For any event $E$, 
\begin{equation}
\bar{\mu}\left(\Omega\backslash E\right)=\left[1,1\right]-\bar{\mu}\left(E\right)\,.\label{eq:classical-complement}
\end{equation}
\item For a collection $\left\{ E_{i}\right\} _{i=1}^{N}$ of pairwise disjoint
events, we have 
\begin{equation}
\bar{\mu}\left(\bigcup_{i=1}^{N}E_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(E_{i}\right)\,.\label{eq:classical-include}
\end{equation}
\end{itemize}
\begin{example}[Two-coin experiment with interval probability]\label{ex3}
We split the unit interval $[0,1]$ in the following four closed sub-intervals:
$[0,0]$ which we call \imposs, $[0,\frac{1}{2}]$ which we call
\unlikely, $[\frac{1}{2},1]$ which we call \likely, and $[1,1]$
which we call \necess. Using these new values, we can modify the
probability measure of example~\ref{ex1} by mapping each numeric
value to the smallest sub-interval containing it to get the following:
\begin{equation}
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\bar{\mu}(\emptyset) & = & \imposs\\
\bar{\mu}(\{HH\}) & = & \unlikely\\
\bar{\mu}(\{HT\}) & = & \imposs\\
\bar{\mu}(\{TH\}) & = & \likely\\
\bar{\mu}(\{TT\}) & = & \imposs\\
\bar{\mu}(\{HH,HT\}) & = & \unlikely\\
\bar{\mu}(\{HH,TH\}) & = & \necess\\
\bar{\mu}(\{HH,TT\}) & = & \unlikely
\end{array} & \begin{array}{rcl}
\bar{\mu}(\{HT,TH\}) & = & \likely\\
\bar{\mu}(\{HT,TT\}) & = & \imposs\\
\bar{\mu}(\{TH,TT\}) & = & \likely\\
\bar{\mu}(\{HH,HT,TH\}) & = & \necess\\
\bar{\mu}(\{HH,HT,TT\}) & = & \unlikely\\
\bar{\mu}(\{HH,TH,TT\}) & = & \necess\\
\bar{\mu}(\{HT,TH,TT\}) & = & \likely\\
\bar{\mu}(\{HH,HT,TH,TT\}) & = & \necess
\end{array}\end{array}
\end{equation}
Despite the absence of infinitely precise numeric information, the
probability measure is quite informative: it reveals that the second
coin is double-headed and that the first coin is biased. To understand
the $\subseteq$-condition, consider the following calculation: 
\begin{equation}\eqalign{ 
& \bar{\mu}(\{HH\})+\bar{\mu}(\{HT\})+\bar{\mu}(\{TH\})+\bar{\mu}(\{TT\})\\
=& \imposs+\unlikely+\imposs+\likely\\
=& \left[0,0\right]+\left[0,\frac{1}{2}\right]+\left[0,0\right]+\left[\frac{1}{2},1\right]= \left[\frac{1}{2},\frac{3}{2}\right]\,. 
}\end{equation}
If we were to equate $\bar{\mu}(\Omega)$ with the sum of the individual
probabilities, we would get that $\bar{\mu}(\Omega)=\left[\frac{1}{2},\frac{3}{2}\right]$.
However, using the fact that $\bar{\mu}(\emptyset)=\imposs$, we have
$\bar{\mu}\left(\Omega\right)=1-\bar{\mu}\left(\emptyset\right)=\necess=[1,1]$.
This interval is tighter and a better estimate for the probability
of the event $\Omega$, and of course it is contained in $[\frac{1}{2},\frac{3}{2}]$.
However it is only possible to exploit the information about the complement
when all four events are combined. Thus the $\subseteq$-condition
allows us to get an estimate for the combined event from each of its
constituents and then gather more evidence knowing the aggregate event.
\end{example}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Quantum Probability}

The mathematical framework of classical probability above assumes
that there exists a predetermined set of events that are independent
of the particular experiment — classical physics is non-contextual~\cite{kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007}.
However, even in classical situations, the structure of the event
space is often only partially known and the precise dependence of
two events on each other cannot, a priori, be determined with certainty.
In the quantum framework, this partial knowledge is compounded by
the fact that there exist non-commuting events which cannot happen
simultaneously. To accommodate these more complex situations, conventional
approaches to quantum probability abandon the sample space~$\Omega$
and reason directly about events which are generalized from plain
sets to projection operators. A quantum probability space therefore
consists of just two components: a set of events $\qevents$ often
formalized as projection operators and a probability measure $\mu:\qevents\rightarrow[0,1]$
formalized using the Born rule~\cite{Born1983,Mermin2007,Jaeger2007}.

%%%%%

\subsection{Quantum Events}

\begin{definition}[Projection Operators; Orthogonality~\cite{10.2307/2308516,Redhead1987-REDINA,peres1995quantum,Griffiths2003,Swart2013}]\label{def:Projection}
Given a Hilbert space $\Hilb$, an event (an experimental proposition~\cite{BirkhoffVonNeumann1936},
a question~\cite{10.2307/2308516,Abramsky2012}, or an elementary
quantum test~\cite{peres1995quantum}) is represented as a (self-adjoint
or orthogonal~\cite{Griffiths2003,Maassen2010}) projection operator
$P:\Hilb\rightarrow\Hilb$ onto a linear subspace of $\Hilb$. The
following define projections and list some of their properties: 
\begin{itemize}
\item $\mathbb{0}$ is a projection. 
\item For any pure state~$\ket{\psi}$, $\proj{\psi}$ is a projection
operator. 
\item Projection operators $P_{0}$ and $P_{1}$ are \emph{orthogonal} if
$P_{0}P_{1}=P_{1}P_{0}=\mathbb{0}$. The sum of two projection operators~$P_{0}+P_{1}$
is also a projection operator if and only if they are orthogonal. 
\item Conversely, every projection~$P$ can be expressed as $\sum_{j=1}^{N}\proj{\psi_{j}}$,
where $P$ actually projects onto the linear subspace with orthonormal
basis~$\left\{ \ket{\psi_{j}}\right\} _{j=1}^{N}$. 
\item A set of projections $\left\{ P_{i}\right\} _{i=1}^{N}$ is called
an \emph{ideal measurement} if it is a partition of the identity,
i.e., $\sum_{i=1}^{N}P_{i}=\mathbb{1}$~\cite{Swart2013}. In this
case, projections $\left\{ P_{i}\right\} _{i=1}^{N}$ must be mutually
orthogonal~\cite{Griffiths2003,Halmos1957}, and $N$ must be less
or equal to the dimension of the Hilbert space. 
\item If $P$ is a projection operator, then $\mathbb{1}-P$ is also a projection
operator, called its \emph{complement}. It is orthogonal to $P$,
and corresponds to the complement event~$\Omega\backslash E$ in
classical probability~\cite{Griffiths2003}. 
\item Projection operators $P_{0}$ and $P_{1}$ \emph{commute} if $P_{0}P_{1}=P_{1}P_{0}$.
The product of two projection operators~$P_{0}P_{1}$ is also a projection
operator if and only if they commute. This corresponds to the classical
intersection between events~\cite{peres1995quantum,Griffiths2003}. 
\item For two commuting projection operators $P_{0}$ and $P_{1}$, their
\emph{disjunction}~$P_{0}\vee P_{1}$ is defined to be $P_{0}+P_{1}-P_{0}P_{1}$~\cite{Griffiths2003}. 
\end{itemize}
\end{definition}

\begin{example}[One-qubit quantum probability space] Consider a
one-qubit Hilbert space with each event interpreted as a possible
post-measurement state~\cite{peres1995quantum,Mermin2007,Jaeger2007}.
For example, the event $\proj{0}$ indicates that the post-measurement
state will be $\ket{0}$; the probability of such an event depends
on the current state; the event $\proj{1}$ indicates that the post-measurement
state will be $\ket{1}$; the event $\proj{\ps}$ where $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$
indicates that the post-measurement state will be $\ket{\ps}$; the
event $\mathbb{1}=\proj{0}+\proj{1}$ indicates that the post-measurement
state will be a linear combination of $\ket{0}$ and $\ket{1}$ and
is clearly certain; finally the empty event $\mathbb{0}$ states that
the post-measurement state will be the empty state and is impossible.
As in the classical case, a probability measure is a function that
maps events to $[0,1]$. Here is a partial specification of a possible
probability measure that would be induced by a system whose current
state is $\ket{0}$, $\mu\left(\mathbb{0}\right)=0$, $\mu\left(\mathbb{1}\right)=1$,
$\mu\left(\proj{0}\right)=1$, $\mu\left(\proj{1}\right)=0$, $\mu\left(\proj{\ps}\right)=\frac{1}{2}$,
\ldots{}. Note that, similarly to the classical case, the probability
of $\mathbb{1}$ is 1 and the probability of collections of orthogonal
events (e.g., $\proj{0}+\proj{1}$) is the sum of the individual probabilities.
A collection of non-orthogonal events (e.g., $\proj{0}$ and $\proj{\ps}$)
is however not even a valid event. In the classical example, we argued
that each probability measure is uniquely determined by two actual
coins. A similar (but much more subtle) argument is valid also in
the quantum case. By postulates of quantum mechanics and Gleason's
theorem, it turns out that for large enough quantum systems, each
probability measure is uniquely determined by an actual quantum state
as discussed next. \end{example}

%%%%%

\subsection{Quantum Probability Measures}

Given our setup, the definition of a quantum probability measure is
a small variation on the classical definition.

\begin{definition}[Quantum Probability Measure~\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010}]\label{def:QuantumProbabilitySpace}
Given a Hilbert space $\Hilb$ with its set of events~$\events$,
a \emph{quantum probability measure} is a function~$\mu:\events\rightarrow[0,1]$
such that: 
\begin{itemize}
\item $\mu(\mathbb{0})=0$. 
\item $\mu(\mathbb{1})=1$. 
\item For any projection $P$, $\mu\left(\mathbb{1}-P\right)=1-\mu\left(P\right)$. 
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
we have $\mu\left(\sum_{i=1}^{N}P_{i}\right)=\sum_{i=1}^{N}\mu\left(P_{i}\right)$. 
\end{itemize}
\end{definition}

\noindent A quantum probability measure can be easily constructed
if one knows the current state of the quantum system by using the
Born rule. Specifically, for each pure normalized quantum state $\ket{\phi}$,
the Born rule induces a probability measure $\mu_{\phi}^{\mathrm{B}}$
defined as $\mu_{\phi}^{\mathrm{\mathrm{B}}}(P)=\ip{\phi}{P\phi}$.
The situation generalizes to mixed states $\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}$,
where $\sum_{j=1}^{N}q_{j}=1$ in which case the generalized Born
rule induces a probability measure $\mu_{\rho}^{\mathrm{\mathrm{B}}}$
defined as $\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(P\right)=\Tr\left(\rho P\right)=\sum_{j=1}^{N}q_{j}\mu_{\phi_{j}}^{\mathrm{\mathrm{B}}}\left(P\right)$~\cite{peres1995quantum,544199,Jaeger2007}.
Conversely every probability measure must be of this form.

\begin{thm}[Gleason's theorem~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}]\label{cor:Gleason's}In
a Hilbert space $\Hilb$ of dimension $d\geq3$, given a quantum probability
measure~$\mu:\events\rightarrow[0,1]$, there exists a unique mixed
state~$\rho$ such that $\mu=\mu_{\rho}^{\mathrm{\mathrm{B}}}$.
\end{thm}

%%%%%

\subsection{Measuring Quantum Probabilities}

Similarly to the classical case, it is possible to estimate quantum
probabilities by utilizing the frequentist approach of the previous
section, assuming identical measurements conditions in each repeated
experiment~\cite{peres1995quantum}. For instance, if one wants to
determine the probability that the spin of a given silver atom is
$+\hbar/2$, a Stern-Gerlach apparatus is built where ideally an inhomogeneous
magnetic field is generated along, let's say, the quantization axis
$z$. One then produces a collimated beam of identically prepared
(neutral) silver atoms that is directed between the poles of the magnet
where a predetermined field-gradient along the $z$ direction has
been established. Under appropriate experimental conditions we will
observe that the beam, after traversing the magnetic-field region,
will be deflected towards two regions identified by distinguished
spots on a detector situated behind the apparatus~\cite{Stern1988,peres1995quantum,544199,Griffiths2003}.
Each of the two discrete values is associated to either $+\hbar/2$
or $-\hbar/2$, commonly called ``spin up'' and ``spin down'',
respectively. By ``counting'' the number of atoms that are deflected
in the ``spin up'' region one can, in principle, estimate the probability
that the prepared state of the silver atom state has spin $+\hbar/2$.
Notice that a real experiment does not necessarily represent an ``ideal
measurement''. For example, not all silver atoms will be identically
prepared, or the field-gradient could not be large enough to distinguish
between the spin up and down situations simply producing a large single
blot. In other words, the closer we get to an ideal measurement the
better we determine those probabilities at the cost of significantly
increasing the number of resources. It is not very well appreciated
in the literature that Bohr attempted to argue against the measurability
of the spin of a free electron~\cite{Bohr1985,MartensDeMuynck1994,McEvoy2001}.
Essentially, Bohr argued (and Mott later on justified his assertion
by an elegant use of uncertainty relations~\cite{10.2307/j.ctt7ztxn5.15})
that a Stern-Gerlach experiment could not succeed in establishing
the spin of an unbound electron because the Lorentz force would blur
the detected pattern. This example illustrates the case of a fundamental
physical limitation that not even infinite resources could mitigate.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Cryptodeterminism}

Although a real or interval-valued probability measure could model
whether we believe an event will happen or not, the classical world
could be modeled in a simpler way. If the complete initial conditions
for tossing a coin or throwing a needle is known, we can determine
whether an event will happen or not for sure~\cite{DiaconisHolmesMontgomery2007}.
In another words, Newtonian physics could be modeled by a cryptodeterministic\footnote{The word ''cryptodeterminism'' is used by Peres which means
\begin{quote}
We now introduce the cryptodeterministic hypothesis: It is possible
to specify all the details of a preparation, so that the result of
any measurement becomes completely deterministic.
\end{quote}
in~\cite{PeresRon1988}, and
\begin{quote}
... any purported cryptodeterministic theory which would attribute
a definite result to each quantum measurement, and still reproduce
the statistical properties of quantum theory ...
\end{quote}
in~\cite{Peres1991,peres1995quantum}.} measure defined as follows.

\begin{definition}[Classical Cryptodeterministic Measure]\label{def:CryptodeterministicMeasures}Given
the set of events~$\events$, a \emph{classical cryptodeterministic
measure} is a function~$\mu^{\mathrm{D}}:\events\rightarrow\left\{ 0,1\right\} $
such that: 
\begin{itemize}
\item $\mu^{\mathrm{D}}(\emptyset)=0$, 
\item $\mu^{\mathrm{D}}(\Omega)=1$, 
\item for every event $E$, $\mu^{\mathrm{D}}\left(\Omega\backslash E\right)=1-\mu^{\mathrm{D}}\left(E\right)$
where $\Omega\backslash E$ is the complement event of $E$, and 
\item for every collection $\left\{ E_{i}\right\} _{i=1}^{N}$ of pairwise
disjoint events, $\mu^{\mathrm{D}}\left(\bigcup_{i=1}^{N}E_{i}\right)=\sum_{i=1}^{N}\mu^{\mathrm{D}}(E_{i})$. 
\end{itemize}
\end{definition}

\begin{example}[Two-coin experiment with cryptodeterministic measure]\label{ex4}Except
the prefect prediction, a cryptodeterministic measure can model the
belief of events which has already happened. For example, if we tossed
the coin in example~\ref{ex1}, and saw $HH$, our belief of which
event has happened is represented by the following example.
\begin{equation}
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\mu^{\mathrm{D}}(\emptyset) & = & 0\\
\mu^{\mathrm{D}}(\{HH\}) & = & 1\\
\mu^{\mathrm{D}}(\{HT\}) & = & 0\\
\mu^{\mathrm{D}}(\{TH\}) & = & 0\\
\mu^{\mathrm{D}}(\{TT\}) & = & 0\\
\mu^{\mathrm{D}}(\{HH,HT\}) & = & 1\\
\mu^{\mathrm{D}}(\{HH,TH\}) & = & 1\\
\mu^{\mathrm{D}}(\{HH,TT\}) & = & 1
\end{array} & \begin{array}{rcl}
\mu^{\mathrm{D}}(\{HT,TH\}) & = & 0\\
\mu^{\mathrm{D}}(\{HT,TT\}) & = & 0\\
\mu^{\mathrm{D}}(\{TH,TT\}) & = & 0\\
\mu^{\mathrm{D}}(\{HH,HT,TH\}) & = & 1\\
\mu^{\mathrm{D}}(\{HH,HT,TT\}) & = & 1\\
\mu^{\mathrm{D}}(\{HH,TH,TT\}) & = & 1\\
\mu^{\mathrm{D}}(\{HT,TH,TT\}) & = & 0\\
\mu^{\mathrm{D}}(\{HH,HT,TH,TT\}) & = & 1
\end{array}\end{array}
\end{equation}
\end{example}

The ability to predict classical events unambiguously rather than
probabilistically inspired Einstein, Podolsky, and Rosen to ask whether
quantum mechanics has any chance to be improved and predict in the
same way~\cite{EPR1935,peres1995quantum,544199,Mermin2007}. This
question can be formulated by whether quantum mechanics could be modeled
by quantum cryptodeterministic measures.

\begin{definition}[Quantum Cryptodeterministic Measure]\label{def:QuantumCryptodeterministicMeasures}Given
a Hilbert space $\Hilb$ with its set of events~$\events$, a \emph{quantum
cryptodeterministic measure} is a function~$\mu^{\mathrm{D}}:\events\rightarrow\left\{ 0,1\right\} $
such that: 
\begin{itemize}
\item $\mu^{\mathrm{D}}(\mathbb{0})=0$, 
\item $\mu^{\mathrm{D}}(\mathbb{1})=1$, 
\item for every event $P$, $\mu^{\mathrm{D}}\left(\Omega\backslash P\right)=1-\mu^{\mathrm{D}}\left(P\right)$
where $\Omega\backslash P$ is the complement event of $P$, and 
\item for every collection $\left\{ P_{i}\right\} _{i=1}^{N}$ of pairwise
disjoint events, $\mu^{\mathrm{D}}\left(\bigcup_{i=1}^{N}P_{i}\right)=\sum_{i=1}^{N}\mu^{\mathrm{D}}(P_{i})$. 
\end{itemize}
\end{definition}

The answer is ``no'' due to the Kochen-Specker theorem whose proof
can also be formulated in the language of cryptodeterministic measure
as in section~\ref{subsec:Quantum-I_D-valued-Probability}.

\begin{thm}[Kochen-Specker~\cite{kochenspecker1967,peres1995quantum,Redhead1987-REDINA}]\label{thm:Kochen-Specker}Given
a Hilbert space of dimension~$d\ge3$, there is no cryptodeterministic
measure~$\mu^{\mathrm{D}}:\events\rightarrow\left\{ 0,1\right\} $.\end{thm}

Although sharing a lot of similarity, the Kochen-Specker theorem highlight
the difference between quantum and classical measures. When we think
a cryptodeterministic measure as a model of pasted events, the Kochen-Specker
theorem also consistent with the usual interpretation of the quantum
theory: we should not ask whether an event happened or not if an event
is incompatible with the event really happened; events compatible
with the event really happened are the only askable events.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Quantum Interval-valued Probability}

%%%%%

\subsection{Quantum Interval-valued Probability Measures}

As argued in the previous sections, given fixed finite resources,
it is only possible to estimate the quantum probabilities within an
interval of confidence. It is therefore natural to propose the notion
of a ``quantum interval-valued probability measure'' that combines
the definitions of conventional quantum probability measures with
classical interval-probabilities.

\begin{definition}[Quantum Interval-valued Probability Measure]\label{def:QuantumInterval-valuedProbability}
Given a Hilbert space $\Hilb$ with quantum events (projections)~$\events$,
and a collection of intervals~$\mathscr{I}$, a \emph{quantum $\mathscr{I}$-interval-valued
probability measure} is a function~$\bar{\mu}:\events\rightarrow\mathscr{I}$
such that: 
\begin{itemize}
\item $\bar{\mu}(\mathbb{0})=\left[0,0\right]$. 
\item $\bar{\mu}(\mathbb{1})=\left[1,1\right]$. 
\item For any projection $P$, 
\begin{equation}
\bar{\mu}\left(\mathbb{1}-P\right)=\left[1,1\right]-\bar{\mu}\left(P\right)\,.\label{eq:QuantumInterval-valuedProbability-Complement}
\end{equation}
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
we have 
\begin{equation}
\bar{\mu}\left(\sum_{i=1}^{N}P_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(P_{i}\right)\,.\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{itemize}
\end{definition}

\noindent It is easy to establish that quantum interval-valued probability
measures generalize conventional quantum probability measures. In
particular, any quantum probability measure can be recast as an interval-valued
measure using arbitrary small intervals as discussed in \ref{sec:Real-and-Interval}.
However, to simplify our discussion, we will focus on only three intervals
$\left[0,0\right]$, $\left[1,1\right]$ and \emph{$\left[0,1\right]$},
where $\left[0,0\right]$ and $\left[1,1\right]$ are called \imposs~and
\necess~as before, and \emph{$\left[0,1\right]$} is called \unknown~because
it provides no information. Given a quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$,
we define a quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
by $\bar{\mu}(P)=\iota\left(\mu(P)\right)$, where $\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is defined by 
\begin{equation} 
\iota(x)=\cases{ 
\necess & if $x=1$ ;\\ 
\imposs & if $x=0$ ;\\ 
\unknown & otherwise. 
}\label{eq:homomorphism} \end{equation}
\begin{figure}
\begin{centering}
\includegraphics[scale=0.38]{figure1} 
\par\end{centering}
\caption{\label{fig:three-dimensional-3-value}This figure illustrates the
1-dimensional projectors mapped by $\bar{\mu}_{\ket{0}}^{\mathrm{\mathrm{B}}}(P)=\iota\left(\mu_{\ket{0}}^{\mathrm{\mathrm{B}}}(P)\right)$
plotted in $\mathbb{R}^{3}$. The red dotted vector is $\ket{0}$,
and $\bar{\mu}_{\ket{0}}^{\mathrm{\mathrm{B}}}\left(\proj{0}\right)=\necess$.
$\ket{0}$ is the normal vector of the plane so that we have $\bar{\mu}_{\ket{0}}^{\mathrm{\mathrm{B}}}\left(\proj{\psi}\right)=\imposs$
for any vector~$\ket{\psi}$ in the plane. For any other vector~$\ket{\psi}$,
we have $\bar{\mu}_{\ket{0}}^{\mathrm{\mathrm{B}}}\left(\proj{\psi}\right)=\unknown$.}
\end{figure}
For example, figure~\ref{fig:three-dimensional-3-value} plots the
1-dimensional projectors mapped by $\iota\left(\mu_{\ket{0}}^{\mathrm{\mathrm{B}}}(P)\right)$.
This measure represents the beliefs of an experimenter with no prior
knowledge about the particular quantum system in question.\footnote{\yutsung{Actually, $\iota\left(\mu_{\ket{0}}^{\mathrm{\mathrm{B}}}(P)\right)$
is quite informative. Since $\iota\left(\mu_{\ket{0}}^{\mathrm{\mathrm{B}}}(P)\right)=\necess$
if and only if $P=\proj{0}$, we know the core of $\iota\left(\mu_{\ket{0}}^{\mathrm{\mathrm{B}}}(P)\right)$
is exactly $\left\{ \mu_{\ket{0}}^{\mathrm{\mathrm{B}}}\right\} $.
The quantum interval-valued probability measure for an experimenter
with no prior knowledge should be $\bar{\mu}(P)=\unknown$ for every
projection~$P$.}}

Similarly, we can recast the result of cryptodeterminism to quantum
interval-valued probability measures as well. In another word, for
any classically event space~$\events$,there is always a classical
interval-valued probability measure mapping to $\mathscr{I}_{\mathrm{D}}=\left\{ \imposs,\necess\right\} $.
However, given a Hilbert space of dimension~$d\ge3$, there is no
quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\mathrm{D}}$
by corollary~\ref{cor:Kochen-Specker-IVPM} of the Kochen-Specker
theorem. The Kochen-Specker theorem provides an example that quantum
interval-valued probability measures behave differently from the classical
one, and the convexity and core in the next section provides another
example that quantum interval-valued probability measures behave unexpectedly.

%%%%%

\subsection{Convexity and Core}

\noindent Formally, we can ask: what can we deduce about the state
of a quantum system given a quantum interval-valued probability measure,
i.e., given observations done with finite resources. In the case that
the intervals are infinitely precise, the question reduces to Gleason's
theorem which states that the state of the quantum system is uniquely
determined by the probability measure. But surely the less resources
are available, the less precise the intervals, and the less we expect
to know about the state of the system. To formally state and answer
this question we begin with defining the \emph{core} and \emph{convexity}
of a probability measure as follows.

\begin{definition} Given a quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$: 
\begin{itemize}
\item A \emph{core} of $\bar{\mu}$ is the set $\mathrm{core}\left(\bar{\mu}\right)=\set{\pmeas:\events\rightarrow[0,1]}{\forall P\in\events.~\pmeas\left(P\right)\in\bar{\mu}\left(P\right)}$. 
\item $\bar{\mu}$ is called convex if 
\begin{equation}
\bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\label{eq:QuantumInterval-valuedProbability-Convex}
\end{equation}
for all commuting $P_{0},P_{1}\in\events$. 
\end{itemize}
\end{definition}

\noindent The core of an interval-valued probability measure is the
set of real-valued infinitely-precise probability measures it approximates.
An interval-valued probability measure is convex if whenever certain
intervals exist then combinations of these intervals must also exist,
guaranteeing we can add and manipulate probabilities coherently. In
the classical world, every convex measure has a non-empty core, which
means that the interval-valued probability measure must approximate
at least one infinitely-precise measure.

\begin{thm}[Shapley~\cite{Shapley1971,Grabisch2016}]\label{thm:Shapley}
Every (classical) convex interval-valued probability measure has a
non-empty core. \end{thm}

In informal terms, this result states that although measurements done
with limited resources and poor precision may not uniquely identify
the true state of the system, there is always at least one system
that is consistent with the measurements. Surprisingly, we discovered
a counterexample to this statement in the quantum case, and have:

\begin{thm}[Empty core for general QIVPMs]\label{thm:Non-extensible-of-Gleason's}In
a Hilbert space of dimension~$d\ge3$, there exists a convex QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
such that $\coreBorn\left(\bar{\mu},\events\right)=\emptyset$.\end{thm}

\begin{proof}To prove the theorem, we first need to construct a QIVPM,
and then verify that no states consistent with it on all projectors.
Given an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2},\ldots,\ket{d-1}\right\} $,
let $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$, $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$,
and $\mathscr{I}_{0}=\left\{ \necess,\imposs,\unknown\right\} $.
Consider the function~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$
defined in Table~\ref{tab:non-Born-QIVPM}
\begin{table}
\caption{\label{tab:non-Born-QIVPM}Define a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$
on a Hilbert space of dimension~$d\ge3$. The projectors mapping
to the same interval are listed in the column labeled by $P$ and
separated by commas.}
\centering{}%
\begin{tabular}{cc}
\toprule 
\addlinespace
$P$ & $\bar{\mu}\left(P\right)$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
$\mathbb{0}$, $\proj{0}$, $\proj{\ps}$, and $\proj{\ps'}$ & $\imposs$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\mathbb{1}$, $\mathbb{1}-\proj{0}$, $\mathbb{1}-\proj{\ps}$, and
$\mathbb{1}-\proj{\ps'}$ & $\necess$\tabularnewline\addlinespace
\midrule 
\addlinespace
All other projectors & $\unknown$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
The situation when $d=3$ is illustrated in figure~\ref{fig:three-dimensional-3-value-2}.
\begin{figure}
\begin{centering}
\includegraphics[scale=0.38]{figure2_1} 
\par\end{centering}
\caption{\label{fig:three-dimensional-3-value-2}This figure illustrates theorem~\ref{thm:Non-extensible-of-Gleason's}
plotted in $\mathbb{R}^{3}$. The red, green, and blue vectors are
$\ket{0}$, $\ket{\ps}$, and $\ket{\ps'}$ respectively. }
\end{figure}
To verify $\bar{\mu}$ be a convex QIVPM, it is sufficient to verify
equation~(\ref{eq:QuantumInterval-valuedProbability-Convex}). There
are several cases:
\begin{itemize}
\item When $P_{0}=\mathbb{0}$, equation~(\ref{eq:QuantumInterval-valuedProbability-Convex})
holds by lemma~\ref{lemma:quantum-convex-condition-lemma}. This
lemma also implies equation~(\ref{eq:QuantumInterval-valuedProbability-Convex})
is true when when $P_{0}=\mathbb{1}$.
\item When $P_{0}=\proj{0}$, if $P_{1}=\proj{0}+P_{2}$, then equation~(\ref{eq:QuantumInterval-valuedProbability-Convex})
holds by lemma~\ref{lemma:quantum-convex-condition-lemma}. Otherwise,
we have $P_{0}P_{1}=\mathbb{0}$, and $\bar{\mu}\left(P_{0}P_{1}\right)=\bar{\mu}\left(P_{0}\right)=\imposs$.
There are two cases:
\begin{itemize}
\item When $P_{1}=\mathbb{1}-\proj{0}$, we have
\begin{equation}\eqalign{ 
 & \bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)=\bar{\mu}\left(\mathbb{1}\right)+\bar{\mu}\left(\mathbb{0}\right)\\
={} & \bar{\mu}\left(\mathbb{1}-\proj{0}\right)+\bar{\mu}\left(\proj{0}\right)=\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\,.
}\end{equation}
\item When $P_{1}\ne\mathbb{1}-\proj{0}$, we have
\begin{equation}
\bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)=\unknown+\imposs=\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\,.
\end{equation}
\end{itemize}
Therefore, equation~(\ref{eq:QuantumInterval-valuedProbability-Convex})
holds when $P_{0}=\proj{0}$. Similarly, equation~(\ref{eq:QuantumInterval-valuedProbability-Convex})
holds when $P_{0}=\proj{\ps}$ or $P_{0}=\proj{\ps'}$. By lemma~\ref{lemma:quantum-convex-condition-lemma},
equation~(\ref{eq:QuantumInterval-valuedProbability-Convex}) is
also true when when $P_{0}=\mathbb{1}-\proj{0}$, $\mathbb{1}-\proj{\ps}$,
or $\mathbb{1}-\proj{\ps'}$.
\item If neither the above two cases holds, we have $\bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\unknown+\unknown=\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)$.
\end{itemize}
Next we will prove $\coreBorn\left(\bar{\mu},\events\right)$ is an
empty set by contradiction. Suppose there is a state $\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}\in\coreBorn\left(\bar{\mu},\events\right)$,
where $\sum_{j=1}^{N}q_{j}=1$ and $q_{j}>0$. Since $\mu_{\rho}^{\mathrm{B}}(P)\in\bar{\mu}(P)$
for any~$P$, the state~$\rho$ must satisfy: 
\begin{equation}
\mu_{\rho}^{\mathrm{B}}(\proj{0})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps'})=0\,,
\end{equation}
which implies $\ip{\phi_{j}}{0}=\ip{\phi_{j}}{\ps}=\ip{\phi_{j}}{\ps'}=0$
for all $j$. The last equation never holds proving the theorem.\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Why Empty Core?}

The possibility that a convex quantum interval-valued probability
measure has an empty core departs from the classical case. To understand
the meaning of it, we will first study how to recover the non-empty
core with extra conditions. We will then consider having an empty
core as an legitimate result, and understand its consequence.

\subsection{Convexity}

One possibility for having an empty core is that the definition of
quantum convexity is too weak. On one hand, strengthening equation~(\ref{eq:QuantumInterval-valuedProbability-Convex})
to non-commuting projectors~$P_{0}$ and $P_{1}$ is unlikely to
succeed because the product of non-commuting projectors is not a projector.
On the other hand, strengthening equation~(\ref{eq:QuantumInterval-valuedProbability-Convex})
among commuting projectors cannot guarantee a non-empty core as well
if we follow what people strengthen convexity classically. Equation~(\ref{eq:QuantumInterval-valuedProbability-Convex})
classically is a special case of 
\begin{equation}
\bar{\mu}'\left(\bigcup_{i=1}^{N}E_{i}\right)+\sum_{{{\scriptstyle I\subseteq\left\{ 1,\ldots,N\right\} }\atop {\scriptstyle \left|I\right|\textrm{ is odd}}}}\bar{\mu}'\left(\bigcap_{i\in I}E_{i}\right)\subseteq\sum_{{{\scriptstyle I\subseteq\left\{ 1,\ldots,N\right\} }\atop {\scriptstyle \left|I\right|\textrm{ is even}}}}\bar{\mu}'\left(\bigcap_{i\in I}E_{i}\right)\label{eq:Dempster-Shafer-game}
\end{equation}
when $N=2$, where $\bar{\mu}'$ is a classical interval-valued probability
measure. If $\bar{\mu}'$ satisfies equation~(\ref{eq:Dempster-Shafer-game})
for all set of event~$\left\{ E_{i}\right\} _{i=1}^{N}\subseteq2^{\Omega}$,
then the left-end point of $\bar{\mu}'$ is called a Dempster-Shafer
belief function. It is reasonable to call a quantum interval-valued
probability~$\bar{\mu}$ Dempster-Shafer if $\bar{\mu}$ satisfies
\begin{equation}
\bar{\mu}\left(\bigvee_{i=1}^{N}P_{i}\right)+\sum_{{{\scriptstyle I\subseteq\left\{ 1,\ldots,N\right\} }\atop {\scriptstyle \left|I\right|\textrm{ is odd}}}}\bar{\mu}\left(\prod_{i\in I}P_{i}\right)\subseteq\sum_{{{\scriptstyle I\subseteq\left\{ 1,\ldots,N\right\} }\atop {\scriptstyle \left|I\right|\textrm{ is even}}}}\bar{\mu}\left(\prod_{i\in I}P_{i}\right)\,.\label{eq:Dempster-Shafer-game-quantum}
\end{equation}
for all set of mutually orthogonal projections~$\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\events$.
However, the quantum interval-valued probability measure in example~\ref{ex:three-dimensional-three-value}
is Dempster-Shafer and has an empty core. In summary, it is unlikely
to guarantee a non-empty core by strengthening convexity.

\subsection{Commuting Sub-Event Space}

Another reason for having an empty core is the existence of non-commuting
observables. Non-commutativity is one of the difference between quantum
and classical since Heisenberg's uncertainty principle states there
is an intrinsic uncertainty measuring non-commuting observables in
quantum mechanics~\cite{Heisenberg1983,peres1995quantum,544199,Griffiths2003,Jaeger2007}.
Similarly, we argue that the quantum phenomenon disappears when we
restrict the domain of a quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
on a mutually commuting subspace~$\events'$. As proved in theorem~\ref{thm:sub-event-space},
the restricted $\bar{\mu}|_{\events'}$ is essentially a classical
interval-valued probability measure. Furthermore, if $\bar{\mu}$
is convex, then $\bar{\mu}|_{\events'}$ will be convex, and have
a non-empty core classically.

When $\bar{\mu}$ is restricted to different commuting $\events'$,
the core elements of each $\bar{\mu}|_{\events'}$ may not be glued
together, and gives a core element of $\bar{\mu}$. The inability
of gluing core elements among commuting $\events'$ to non-commuting
$\events$ is consistent with the proof of the Kochen-Specker theorem,
where each triple of commuting observables could be colored well,
but we cannot color all observables which are not all commuting. This
is also consistent with the assertion that mutually commuting observables
could be measured simultaneously~\cite{Mermin_1993}. In another
word, if Alice and Bob in Bell's theorem can only pick from a set
of mutually commuting observables, they could never observe any ``quantum,``
and prove the quantum theory cannot be simulated by a local hidden
variable theory.

\subsection{Increase the Measurement Resource}

This section extend the idea of restricting the event space, but instead
of commuting sub-event space, we will try to find some subset~$\events'\subseteq\events$
such that $\bar{\mu}|_{\events'}$ has an non-empty core. This could
help us understand the core because of the following identity
\begin{equation}
\mathrm{core}\left(\bar{\mu}\right)=\bigcap_{\events'\subseteq\events}\mathrm{core}\left(\bar{\mu}|_{\events'}\right)\label{eq:core-intersection}
\end{equation}
if we know more about the right-hand side of the equation, we can
have more idea about the left-hand side. This idea can be applied
to example~\ref{ex:three-dimensional-three-value}, and the following
are three of all possibilities:
\begin{itemize}
\item When $\events_{1}=\events\backslash\left\{ \proj{\ps'}\right\} $,
because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$, we have
$\mathrm{core}\left(\bar{\mu}|_{\events_{1}}\right)=\left\{ \mu_{\ket{2}}^{\mathrm{\mathrm{B}}}\right\} $.
\item When $\events_{2}=\events\backslash\left\{ \proj{\ps}\right\} $,
because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps'})=\imposs$, we
have $\mathrm{core}\left(\bar{\mu}|_{\events_{2}}\right)=\left\{ \mu_{\ket{1}}^{\mathrm{\mathrm{B}}}\right\} $.
\item When $\events_{3}=\events\backslash\left\{ \proj{0},\proj{\ps},\proj{\ps'}\right\} $,
because $\bar{\mu}(\proj{\phi})\ne\necess$ for all $\ket{\phi}$,
we have $\mu_{\frac{\mathbb{1}}{3}}^{\mathrm{\mathrm{B}}}\in\mathrm{core}\left(\bar{\mu}|_{\events_{3}}\right)$
as well as $\mu_{\ket{2}}^{\mathrm{\mathrm{B}}}$ and $\mu_{\ket{1}}^{\mathrm{\mathrm{B}}}\in\mathrm{core}\left(\bar{\mu}|_{\events_{3}}\right)$.
\end{itemize}
Each of these possibilities is compatible with some but not all of
the observations. In another word, 
\begin{equation}
\mathrm{core}\left(\bar{\mu}\right)=\bigcap_{\events'\subseteq\events}\mathrm{core}\left(\bar{\mu}|_{\events'}\right)\subseteq\mathrm{core}\left(\bar{\mu}|_{\events_{1}}\right)\cap\mathrm{core}\left(\bar{\mu}|_{\events_{2}}\right)=\left\{ \mu_{\ket{2}}^{\mathrm{\mathrm{B}}}\right\} \cap\left\{ \mu_{\ket{1}}^{\mathrm{\mathrm{B}}}\right\} =\emptyset\,.
\end{equation}
The first two possibilities are particular interesting since each
two 1-dimensional projectors\nb$\proj{\psi_{0}}$ and $\proj{\psi_{1}}$
mapped to $\imposs$ suggests there is a subset~$\events'\subseteq\events$
such that $\mathrm{core}\left(\bar{\mu}|_{\events'}\right)=\left\{ \mu_{\ket{\psi_{0}}\times\ket{\psi_{1}}}^{\mathrm{\mathrm{B}}}\right\} $
in the Hilbert space of dimension 3, where the cross product of complex
vectors is defined in definition~\ref{def:cross-product}.

As we illustrated in figure~\ref{fig:three-dimensional-3-value},
if $\bar{\mu}$ has a non-empty core, the states mapping to $\imposs$
is contained in a plane, which contains no area in the sphere. On
the other hand, if the states mapping to $\imposs$ contain some area
in the sphere as in figure~\ref{fig:three-dimensional-real-frame-1},
and this area intersects a plane~$\events'$, then $\mathrm{core}\left(\bar{\mu}|_{\events'}\right)=\left\{ \mu_{\ket{\psi}}^{\mathrm{\mathrm{B}}}\right\} $,
where $\ket{\psi}$ is the normal vector of $\events'$. Since different
planes have different normal vectors, $\bar{\mu}$ has an empty core.
Furthermore, the larger the area, the more state vectors which are
inconsistent with any particular Born rule probability measure. In
another word, there are more different singleton sets in the right-hand
side of equation~(\ref{eq:core-intersection}). Therefore, the area
of the 1-dimensional projectors mapping to $\imposs$ could indicate
how much $\bar{\mu}$ deviate from the Born rule probability measures.

In this sense, we found that the area of the 1-dimensional projectors
mapping to $\imposs$ depends on the maximal length of the intervals,
which represents the ability of measurement equipment. In particular,
we have the following results whose details are described in \ref{sec:Real-Interval-valued-Frame}. 
\begin{itemize}
\item The radius of any disk inside the area of the 1-dimensional projectors
mapping to $\imposs$ must smaller than $\pi/4$ (theorem~\ref{thm:pi-div-4}). 
\item When the radius of any disk inside the area of the 1-dimensional projectors
mapping to $\imposs$ between $\arcsin\left(1/\sqrt{3}\right)$ and
$\pi/4$, we have $\unknown\in\mathscr{I}$, i.e., $\underset{\left[l,r\right]\in\mathscr{I}}{\max}r-l=1$
(theorem~\ref{thm:arcsin-one-over-sqrt-three}). 
\item When the radius of any disk inside the area of the 1-dimensional projectors
mapping to $\imposs$ between $\pi/6$ and $\arcsin\left(1/\sqrt{3}\right)$,
we have $\underset{\left[l,r\right]\in\mathscr{I}}{\max}r-l=\frac{1}{3}$
(to be typed...). 
\end{itemize}
In general, we conjecture that if the measurement equipment is made
more and more precise, the corresponding interval-valued probability
measure will be closer and closer to the Born rule. In the limit case,
$\mathscr{I}=\set{\left\{ a\right\} }{a\in\left[0,1\right]}$ we do
indeed recover the conventional Gleason's theorem and the Born rule.

%%%%%

\subsection{Wigner's Friends}

The situation of empty core might not be that strange if we think
in terms of Wigner's ``friends.'' In the original story of Wigner
and his friend~\cite{Wigner1961,FuchsMerminSchack2014}, the friend
makes a measurement in a closed laboratory and experiences an outcome.
Wigner, outside the laboratory, doesn't experience an outcome. If
he believes what his friend has told him about her plans in her laboratory
he will assign an entangled state to her, her apparatus, and the system
on which she is making her intervention. If Wigner goes on to ask
his friend about her experience, then the entangled state to her,
her apparatus, and the system will change into one in which they even
have separate wave functions, and Wigner and his friend will use the
same wave function to describe the system.

The question is what happens if Wigner has more than one friend in
the laboratory. Suppose Wigner has three friends: E, P, and R in the
laboratory. E, P, and R measure the spin-1 system described in example~\ref{ex:three-dimensional-three-value}.
After the experiment, E tells Wigner $\proj{0}$ never happened, P
tells Wigner $\proj{\ps}$ never happened, and R tells Wigner $\proj{\ps'}$
never happened. Wigner may then need to use the quantum Dempster-Shafer
rule(?) or other rules(?) to combine the belief told by his friends,
and the combined belief might be the quantum interval-valued probability
measure in example~\ref{ex:three-dimensional-three-value} which
has an empty core.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion and Discussion}

If we insist that probabilities cannot be computed to infinite precision
and are bound to be approximations represented by intervals of confidence,
then quantum states themselves can only be discussed within intervals
of confidence. Despite the fact that classically a IVPM could always
correspond to a real-valued probability measure, we found that there
may be no ``real'' infinitely-precise quantum state approximated
by a quantum IVPM as a whole. However, if a quantum IVPM is decomposed
into pieces, each piece might still be induced by a quantum state.
Moreover, our examples and Gleason's theorem summarized in table~\ref{table_long}
suggests that as the measurement resources increase, an entire quantum
IVPM could more and more likely be identified to a particular state.
\begin{table}
\caption{\label{table_long}Relation between measurement resources and interval-valued
probability measures.}

\begin{indented}
\item\begin{tabular}{c|cccc} 
\br 
Measurement resources & Lowest & \multicolumn{2}{c}{$\longleftrightarrow$} & Highest\tabularnewline 
\mr 
Count of $\mathscr{I}$ & $3$ & $4$ & \ldots{} & $\infty$\tabularnewline 
$\sup_{\left[l,r\right]\in\mathscr{I}}\left|r-l\right|$ & $1$ & $\frac{1}{2}$ & \ldots{} & $0$\tabularnewline 
Count of $\mathrm{core}\left(\bar{\mu}\right)$ & 0 & 0 &  & 1\tabularnewline 
Count of states correspond to pieces of $\bar{\mu}$ & $3$ & $2$ & \ldots{} & $1$\tabularnewline 
\mr 
How precise we could identify a state? & Coarse & \multicolumn{2}{c}{$\longleftrightarrow$} & Precise\tabularnewline 
\br 
\end{tabular} 
\end{indented}

\end{table}

Additional work is in progress on the following research questions: 
\begin{description}
\item [{Research Question.}] Confirm that there is always a quantum IVPM
with an empty-core if the length of the intervals is bigger than zero. 
\item [{Research Question.}] Find a general method to determine how a
quantum IVPM is close to the Born rule. 
\item [{Research Question.}] Confirm that as the number and precision
of the intervals increase the quantum IVPM converges to the measure
induced by the Born rule. 
\item [{Research Question.}] Investigate the status of the theorems of
Bell and Kochen-Specker for quantum IVPM. 
\end{description}
Although we discussed how a quantum IVPM could be induced from a real
quantum state, our investigation leaves an open question of whether
there exists a ``real'' entity that exists independently of measurements
and probabilities. The possibility of no ``real'' underlying state
is consistent with the elegantly recent work on Quantum Bayesianism
or QBism~\cite{FuchsSchack2013,FuchsMerminSchack2014,VonBaeyer2016},
which suggests that the quantum state is more like an interactive
system in computer science parlance. In another word, the quantum
state is subjective: each observer has a different view of the quantum
system that is consistent with their previous observations and that
allows that observer, independently of other observers, to assign
beliefs (i.e., probabilities) to possible future interactions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\ack

Anyone who has discussed this with us?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix
%dummy comment inserted by tex2lyx to ensure that this paragraph is not empty%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Interval-valued Frame Functions\label{sec:Interval-valued-Frame-Functions}}

Although quantum IVPMs in example~\ref{ex:three-dimensional-three-value}
and \ref{ex:three-dimensional-5-value} have empty cores, we might
still think that the probability measure~$\bar{\mu}$ is induced
by some states. In example~\ref{ex:three-dimensional-three-value},
the probability measure~$\bar{\mu}$ might be induced by the state
$\ket{2}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$,
or induced by the state $\ket{1}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps'})=\imposs$.
In example~\ref{ex:three-dimensional-5-value}, the probability measure~$\bar{\mu}$
might be induced by the state $\ket{2}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$.
In general, each two real 1-dimensional projectors~$\proj{\psi_{0}}$
and $\proj{\psi_{1}}$ mapping to impossible suggests $\bar{\mu}$
might be induced by the pure state~$\ket{\psi_{0}}\times\ket{\psi_{1}}$
in the Hilbert space of dimension 3.

Since a quantum IVPM is completely determined by its value on 1-dimensional
projectors in the Hilbert space of dimension 3, we could define the
interval-valued frame function to simplify the later discussion.

\begin{definition}\cite{Hatcher2001,dryden2005}~The unit sphere
in $\mathbb{R}^{d}$ is denoted by $S^{d-1}=\set{\ket{\psi}\in\mathbb{R}^{d}}{\ip{\psi}{\psi}=1}$,
while the unit sphere in $\mathbb{C}^{d}$ is denoted by $\mathbb{C}S^{d-1}=\set{\ket{\psi}\in\mathbb{C}^{d}}{\ip{\psi}{\psi}=1}$.
Since $\mathbb{C}$ is one-to-one correspondence to $\mathbb{R}^{2}$~\cite{GAMELIN2003},
$\mathbb{C}S^{d-1}$ has the same structure as $S^{2d-1}$. However,
different symbols emphasis that they are embedded in different spaces.\end{definition}

\begin{definition}\cite{gleason1957,peres1995quantum,RichmanBridges1999}~For
any quantum probability measure~$\mu$ on the Hilbert space of dimension~$d$,
we can define its \emph{frame function}~$f:\mathbb{C}S^{d-1}\rightarrow\left[0,1\right]$
by $f\left(\ket{\psi}\right)=\mu\left(\proj{\psi}\right)$. For a
quantum probability measure~$\mu_{\phi}^{\mathrm{\mathrm{B}}}$ induced
by the Born rule, its frame function is denoted by $f_{\phi}^{\mathrm{\mathrm{B}}}$.\end{definition}

\begin{thm}\label{thm:frame-function}\cite{gleason1957,peres1995quantum,RichmanBridges1999}~A
frame function~$f:\mathbb{C}S^{d-1}\rightarrow\left[0,1\right]$
satisfies the following properties: 
\begin{itemize}
\item $f\left(\lambda\ket{\psi}\right)=f\left(\ket{\psi}\right)$ for any
$\ket{\psi}\in\mathbb{C}S^{d-1}$ and $\lambda\in\mathbb{C}S^{0}$. 
\item For any orthonormal basis~$\left\{ \ket{\psi_{i}}\right\} _{i=0}^{2}$,
we have $1=\sum_{i=0}^{2}f\left(\ket{\psi_{i}}\right)$. 
\end{itemize}
Conversely, if a function~$f:\mathbb{C}S^{d-1}\rightarrow\left[0,1\right]$
satisfies the above properties, $f$ must be a frame function.\end{thm}

\begin{definition}For any quantum IVPM~$\bar{\mu}$ on the Hilbert
space of dimension 3, we can define an \emph{interval-valued frame
function}~$\bar{f}:\mathbb{C}S^{2}\rightarrow\mathscr{I}$ by $\bar{f}\left(\ket{\psi}\right)=\bar{\mu}\left(\proj{\psi}\right)=\left[f^{l}\left(\ket{\psi}\right),f^{r}\left(\ket{\psi}\right)\right]$,
where $f^{l}:\mathbb{C}S^{2}\rightarrow\left[0,1\right]$ and $f^{r}:\mathbb{C}S^{2}\rightarrow\left[0,1\right]$
are the left-end and the right-end of $\bar{f}$, respectively. Also,
if $f$ is the frame function of $\mu$, $\bar{f}$ is the interval-valued
frame function of $\bar{\mu}$, and $\mu$ is in the core of $\bar{\mu}$,
then we say $f$ is in the core of $\bar{f}$, and denoted by $f\in\mathrm{core}\left(\bar{f}\right)$.\end{definition}

\begin{thm}\label{thm:interval-valued-frame-function}An interval-valued
frame function~$\bar{f}:\mathbb{C}S^{2}\rightarrow\mathscr{I}$ satisfies
the following properties:
\begin{itemize}
\item $\bar{f}\left(\lambda\ket{\psi}\right)=\bar{f}\left(\ket{\psi}\right)$
for any $\ket{\psi}\in\mathbb{C}S^{2}$ and $\lambda\in\mathbb{C}S^{0}$. 
\item For any orthonormal basis~$\left\{ \ket{\psi_{i}}\right\} _{i=0}^{2}$,
we have $\left[1,1\right]-\bar{f}\left(\ket{\psi_{i}}\right)\subseteq\bar{f}\left(\ket{\psi_{j}}\right)+\bar{f}\left(\ket{\psi_{k}}\right)$,
where $i$, $j$, and $k$ are any permutations among $0$, $1$,
and $2$. 
\end{itemize}
Conversely, if a function~$\bar{f}:\mathbb{C}S^{2}\rightarrow\mathscr{I}$
satisfies the above properties, $\bar{f}$ must be an interval-valued
frame function.\end{thm}

\begin{proof} Given a quantum IVPM~$\bar{\mu}$, it is straightforward
to verify that its interval-valued frame function~$\bar{f}:\mathbb{C}S^{2}\rightarrow\mathscr{I}$
satisfies the above properties.

Conversely, given a function~$\bar{f}:\mathbb{C}S^{2}\rightarrow\mathscr{I}$
satisfying the above properties, consider $\bar{\mu}:\events\rightarrow\mathscr{I}$
defined by 
\begin{itemize}
\item $\bar{\mu}(\mathbb{0})=\left[0,0\right]$ and $\bar{\mu}(\mathbb{1})=\left[1,1\right]$. 
\item For any $\ket{\psi}\in\mathbb{C}S^{2}$, we have $\bar{\mu}\left(\proj{\psi}\right)=\bar{f}\left(\ket{\psi}\right)$
and $\bar{\mu}\left(\mathbb{1}-\proj{\psi}\right)=\left[1,1\right]-\bar{f}\left(\ket{\psi}\right)$. 
\end{itemize}
Since other conditions are trivial, we only need to verify equation~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}).
First notice that $\bar{\mu}\left(P_{0}+P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)$
implies equation~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
by induction. Second, equation~(\ref{eq:QuantumInterval-valuedProbability-Complement})
implies $\bar{\mu}\left(\mathbb{1}\right)\subseteq\bar{\mu}\left(P\right)+\bar{\mu}\left(\mathbb{1}-P\right)$.
Therefore, to prove equation~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}),
it is sufficient to check 
\begin{equation}
\bar{\mu}\left(\proj{\psi_{0}}+\proj{\psi_{1}}\right)\subseteq\bar{\mu}\left(\proj{\psi_{0}}\right)+\bar{\mu}\left(\proj{\psi_{1}}\right)
\end{equation}
for orthogonal $\ket{\psi_{0}}$ and $\ket{\psi_{1}}$,which always
holds because the properties of $\bar{f}$.\end{proof}

\begin{definition}\label{def:cross-product}For $\ket{\psi}=\alpha_{0}\ket{0}+\alpha_{1}\ket{1}+\alpha_{2}\ket{2}=\left(\begin{array}{c}
\alpha_{0}\\
\alpha_{1}\\
\alpha_{2}
\end{array}\right)$ and $\ket{\phi}=\beta_{0}\ket{0}+\beta_{1}\ket{1}+\beta_{2}\ket{2}=\left(\begin{array}{c}
\beta_{0}\\
\beta_{1}\\
\beta_{2}
\end{array}\right)$, their cross product is defined as follow: 
\begin{equation}
\ket{\psi\times\phi}=\ket{\psi}\times\ket{\phi}=\left|\begin{array}{cc}
\alpha_{1}^{*} & \beta_{1}^{*}\\
\alpha_{2}^{*} & \beta_{2}^{*}
\end{array}\right|\ket{0}+\left|\begin{array}{cc}
\alpha_{2}^{*} & \beta_{2}^{*}\\
\alpha_{0}^{*} & \beta_{0}^{*}
\end{array}\right|\ket{1}+\left|\begin{array}{cc}
\alpha_{0}^{*} & \beta_{0}^{*}\\
\alpha_{1}^{*} & \beta_{1}^{*}
\end{array}\right|\ket{2}\,,
\end{equation}
where $\left|\begin{array}{cc}
\alpha & \beta\\
\alpha' & \beta'
\end{array}\right|=\alpha\beta'-\alpha'\beta$. Also, even if $\ket{\psi}$ and $\ket{\phi}$ are normalized, their
cross product~$\ket{\psi}\times\ket{\phi}$ need not be normalized
as usual. We will use $\ket{\psi}\times\ket{\phi}$ and $\ket{\psi}\times\ket{\phi}/\left\Vert \ket{\psi}\times\ket{\phi}\right\Vert $
interchanged when there is no confusion.\end{definition}

\begin{definition}For $\ket{\psi}$ and $\ket{\phi}\in\mathbb{C}^{3}$,
if $\ket{\psi}=\lambda\ket{\phi}$ for some~$\lambda\ne0$, we call
$\ket{\psi}$ is parallel to $\ket{\phi}$ denoted by $\ket{\psi}\parallel\ket{\phi}$.
In this case, $\ket{\psi}$ and $\ket{\phi}$ represent the same physical
state. ``$\ket{\psi}$ is not parallel to $\ket{\phi}$'' is denoted
by $\ket{\psi}\nparallel\ket{\phi}$.\end{definition}

\begin{lemma}Given $\ket{\psi}$, $\ket{\phi}$ and $\ket{\varphi}\in\mathbb{C}^{3}$
such that $\ket{\psi}\nparallel\ket{\phi}$. Then, $\ket{\varphi}\parallel\left(\ket{\psi}\times\ket{\phi}\right)$
if and only if $\ip{\psi}{\varphi}=\ip{\phi}{\varphi}=0$.\end{lemma}
\begin{proof} To prove that $\ket{\varphi}\parallel\left(\ket{\psi}\times\ket{\phi}\right)\Rightarrow\ip{\psi}{\varphi}=\ip{\phi}{\varphi}=0$,
we can just verify 
\begin{equation}\eqalign{ 
&\ip{\psi}{\psi\times\phi}=\ip{\psi\times\phi}{\psi}^{*}\\ 
=&\left[\left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\alpha_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\alpha_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\alpha_{2}\right]^{*}=0\,,\\ 
&\ip{\phi}{\psi\times\phi}=\ip{\psi\times\phi}{\phi}^{*}\\ 
=&\left[\left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\beta_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\beta_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\beta_{2}\right]^{*}=0\,. 
}\end{equation}

To prove the other direction, consider the linear transformation~$T:\mathbb{C}^{3}\rightarrow\mathbb{C}^{2}$
defined by $T\left(\ket{\varphi}\right)=\left(\begin{array}{c}
\ip{\psi}{\varphi}\\
\ip{\phi}{\varphi}
\end{array}\right)$. Since $\ket{\psi}\nparallel\ket{\phi}$, we have $\mathrm{rank}\left(T\right)=2$.
Together with the rank equation~\cite{FraleighBeauregard1995}, we
have 
\begin{equation}
\dim\left(\ker T\right)=\mathrm{nullity}\left(T\right)=3-\mathrm{rank}\left(T\right)=3-2=1\,.
\end{equation}
By the other direction of the proof, we have already known $\ket{\psi}\times\ket{\phi}\in\ker T$.
Therefore, $\ket{\varphi}\parallel\left(\ket{\psi}\times\ket{\phi}\right)$.
\end{proof} Since interval-valued probability measures~$\bar{\mu}$
and interval-valued frame functions~$\bar{f}$ are one-to-one correspondence,
studying 1-dimensional projectors mapping to impossible by $\bar{\mu}$
is the same as studying the points on $\mathbb{C}S^{2}$ mapping to
impossible by $\bar{f}$. These points, denoted by $\bar{f}^{-1}\left(\imposs\right)=\set{\ket{\psi}\in\mathbb{C}S^{2}}{\bar{f}(\ket{\psi})=\imposs}$,
play an important role to understand $\bar{f}$. There are 4 different
cases when analyzing $\bar{f}^{-1}\left(\imposs\right)$: 
\begin{enumerate}
\item $\bar{f}^{-1}\left(\imposs\right)=\emptyset$. If $f_{\phi}^{\mathrm{\mathrm{B}}}\in\mathrm{core}\left(\bar{f}\right)$
for some pure state~$\ket{\phi}$, we could say nothing about $\ket{\phi}$. 
\item $\bar{f}^{-1}\left(\imposs\right)=\left\{ \ket{\psi_{0}}\right\} $.
If $f_{\phi}^{\mathrm{\mathrm{B}}}\in\mathrm{core}\left(\bar{f}\right)$
for some pure state~$\ket{\phi}$, $\ket{\phi}$ could be any state
orthogonal to $\ket{\psi_{0}}$. 
\item Consider $\left\{ \ket{\psi_{0}},\ket{\psi_{1}}\right\} \subseteq\bar{f}^{-1}\left(\imposs\right)$.
If $\left(\ket{\psi_{0}}\times\ket{\psi_{1}}\right)\parallel\left(\ket{\psi_{0}'}\times\ket{\psi_{1}'}\right)$
for any $\left\{ \ket{\psi_{0}'},\ket{\psi_{1}'}\right\} \subseteq\bar{f}^{-1}\left(\imposs\right)$,
then $\ket{\psi_{0}}\times\ket{\psi_{1}}$ is the only pure state
which might induce $\bar{f}$. In another word, if $f_{\phi}^{\mathrm{\mathrm{B}}}\in\mathrm{core}\left(\bar{f}\right)$
for some pure state~$\ket{\phi}$, then $\ket{\phi}\parallel\left(\ket{\psi_{0}}\times\ket{\psi_{1}}\right)$.
For example, example~\ref{ex:three-dimensional-5-value} is in this
case when the core is empty; an interval-valued frame function defined
by $\bar{f}\left(\ket{\psi}\right)=\iota\left(f_{\phi}^{\mathrm{\mathrm{B}}}\left(\ket{\psi}\right)\right)$
is also in this case when the core is the singleton set~$\left\{ f_{\phi}^{\mathrm{\mathrm{B}}}\right\} $,
where $\iota$ is defined in equation~(\ref{eq:homomorphism}). 
\item If there exist two pair of states and $\left\{ \ket{\psi_{0}},\ket{\psi_{1}}\right\} $
and $\left\{ \ket{\psi_{0}'},\ket{\psi_{1}'}\right\} \subseteq\bar{f}^{-1}\left(\imposs\right)$
such that $\left(\ket{\psi_{0}}\times\ket{\psi_{1}}\right)\nparallel\left(\ket{\psi_{0}'}\times\ket{\psi_{1}'}\right)$,
then both $\ket{\psi_{0}}\times\ket{\psi_{1}}$ and $\ket{\psi_{0}'}\times\ket{\psi_{1}'}$
might induce $\bar{f}$. However, neither $f_{\psi_{0}\times\psi_{1}}^{\mathrm{\mathrm{B}}}$
nor $f_{\psi_{0}'\times\psi_{1}'}^{\mathrm{\mathrm{B}}}$ is in the
core of $\bar{f}$ because $f_{\psi_{0}\times\psi_{1}}^{\mathrm{\mathrm{B}}}$
is inconsistent with $\bar{f}\left(\ket{\psi_{0}'}\right)=\bar{f}\left(\ket{\psi_{1}'}\right)=\imposs$
and $f_{\psi_{0}'\times\psi_{1}'}^{\mathrm{\mathrm{B}}}$ is inconsistent
with $\bar{f}\left(\ket{\psi_{0}}\right)=\bar{f}\left(\ket{\psi_{1}}\right)=\imposs$.
Actually, $\bar{f}$ has an empty core, and this is the case for example~\ref{ex:three-dimensional-three-value}. 
\end{enumerate}
In summary, if we have an idea of $\bar{f}^{-1}\left(\imposs\right)$,
we could have some idea of the pure states in core.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Real Interval-valued Frame Functions\label{sec:Real-Interval-valued-Frame}}

In order to simplify the discussion, we call a function\nb$\bar{f}:S^{2}\rightarrow\mathscr{I}$
satisfying conditions in theorem~\ref{thm:interval-valued-frame-function}
a real interval-valued frame function. Obviously, any results in appendix~\ref{sec:Interval-valued-Frame-Functions}
still hold if we replace every ``complex'' by ``real,'' in particular,
replace every $\mathbb{C}S^{2}$ by $S^{2}$. Furthermore, since restricting
the domain of a complex interval-valued frame function to $S^{2}$
gives a real interval-valued frame function, if we prove there is
no real interval-valued frame function satisfying a condition, then
there must be no complex interval-valued frame function satisfying
the corresponding condition. From now on, when we write an interval-valued
frame function~$\bar{f}$ without other specification, we would always
means a real interval-valued frame function~$\bar{f}:S^{2}\rightarrow\mathscr{I}$.

If $\bar{f}$ has a non-empty core, $\bar{f}$ must be in the case
1 - 3 in the end of appendix~\ref{sec:Interval-valued-Frame-Functions}
so that $\bar{f}^{-1}\left(\imposs\right)$ contains no area in $S^{2}$.
In contrast, if $\bar{f}^{-1}\left(\imposs\right)$ contains some
area in $S^{2}$, it must be the situation of case 4, and $\bar{f}$
has an empty core. Furthermore, the larger the area, the more state
vectors which are inconsistent with any particular Born rule frame
function. Therefore, the area of $\bar{f}^{-1}\left(\imposs\right)$
could indicate how much $\bar{f}$ deviate from the Born rule frame
functions. In the following paragraphs, we will establish that if
the area of $\bar{f}^{-1}\left(\imposs\right)$ is big, then the interval~$\mathscr{I}$
must be coarse. In another word, if the interval~$\mathscr{I}$ is
finer, then the area of $\bar{f}^{-1}\left(\imposs\right)$ would
be smaller so that $\bar{f}$ is closer to a Born rule frame function.

In order to simplify the discussion, consider the center of $\bar{f}^{-1}\left(\imposs\right)$
as the north pole, and assume $\bar{f}$ maps every state whose latitude
is larger than $\theta_{0}$ to impossible.

\begin{definition}Given a vector~$\ket{\psi}\in S^{2}$ with a specified
north pole~$\ket{0}\in S^{2}$, then $\angle\left(\ket{\psi}\right)=\arccos\ip{0}{\psi}$
denotes the latitude of $\ket{\psi}$. Conversely, $\angle^{-1}\left(I\right)=\set{\ket{\psi}\in S^{2}}{\angle\left(\ket{\psi}\right)\in I}$.
For example, the north pole has the latitude $0$, the equator has
the the latitude $\pi/2$, and $\angle^{-1}\left(\left[0,\theta_{0}\right]\right)=\set{\ket{\psi}\in S^{2}}{\theta_{0}\le\angle\left(\ket{\psi}\right)\le\pi/2}$
contains all states whose latitude is larger than $\theta_{0}$.\end{definition}

\begin{lemma}\label{lemma:imposs-imply-same-value}Given $\theta_{0}\in\left[0,\pi/2\right]$
and an interval-valued frame function~$\bar{f}:S^{2}\rightarrow\mathscr{I}$,
if $\bar{f}\left(\ket{\psi}\right)=\imposs$ for every $\ket{\psi}\in\angle^{-1}\left(\left[0,\theta_{0}\right]\right)$,
then there is $l_{0}\in\left[0,\frac{1}{2}\right]$ such that $\bar{f}\left(\ket{\psi_{0}}\right)=\left[l_{0},1-l_{0}\right]\in\mathscr{I}$
for any $\ket{\psi_{0}}\in\angle^{-1}\left(\left[\pi/2-\theta_{0},\pi/2\right]\right)$.\end{lemma}
\begin{figure}
\begin{centering}
\includegraphics[scale=0.38]{figureb1} 
\par\end{centering}
\caption{\label{fig:three-dimensional-real-frame-1}This figure is used in
the proof of lemma~\ref{lemma:imposs-imply-same-value}, where every
vector above the light green circle maps to $\imposs$, and we want
to prove that every vector between dark green and light blue maps
to the same interval.}
\end{figure}

\begin{proof} We could illustrate the situation in figure~\ref{fig:three-dimensional-real-frame-1}.
Let the yellow circle is the states whose latitude is $\theta_{0}$.
Then, every states above the yellow circle maps to $\imposs$ by $\bar{f}$.
In particular, $\bar{f}\left(\ket{0}\right)=\bar{f}\left(\ket{0'}\right)=\imposs$.
Since $\left\{ \ket{0'},\ket{1},\ket{2'}\right\} $ is an orthonormal
basis, by theorem~\ref{thm:interval-valued-frame-function}, we have

\begin{equation} 
\eqalign{\left[1,1\right]-\bar{f}\left(\ket{1}\right)\subseteq\bar{f}\left(\ket{0'}\right)+\bar{f}\left(\ket{2}\right)=\bar{f}\left(\ket{2'}\right)\\ 
\left[1,1\right]-\bar{f}\left(\ket{2'}\right)\subseteq\bar{f}\left(\ket{0'}\right)+\bar{f}\left(\ket{1}\right)=\bar{f}\left(\ket{1}\right) 
} 
\label{eq:includes} 
\end{equation}
so that $\bar{f}\left(\ket{2'}\right)=\left[1,1\right]-\bar{f}\left(\ket{1}\right)$.
By the same reason, for any state~$\ket{\psi}$ in the dotted black
circle between $\ket{2'}$ and $\ket{2}$, we have $\bar{f}\left(\ket{\psi}\right)=\left[1,1\right]-\bar{f}\left(\ket{1}\right)=\bar{f}\left(\ket{2'}\right)$.
Since the angle between $\ket{0}$ and $\ket{0'}$ is the same as
the angle between $\ket{2}$ and $\ket{2'}$, for any $\ket{\psi_{0}}$
and $\ket{\psi_{1}}\in\angle^{-1}\left(\left[0,\theta_{0}\right]\right)$,
if $\ket{\psi_{0}}$ and $\ket{\psi_{1}}$ are in the same longitude,
then $\bar{f}\left(\ket{\psi_{0}}\right)=\bar{f}\left(\ket{\psi_{1}}\right)$.

Moreover, the same argument can apply on the great circle which is
not longitude. Since $\ket{\bar{1}}$ is the normal vector of the
gray circle which includes $\ket{2}$ and $\ket{\bar{2}}$, we have
$\bar{f}\left(\ket{2}\right)=\left[1,1\right]-\bar{f}\left(\ket{\bar{1}}\right)=\bar{f}\left(\ket{\bar{2}}\right)$.
In another words, for any $\ket{\psi_{0}}$ and $\ket{\psi_{1}}\in\angle^{-1}\left(\left[\pi/2-\theta_{0},\pi/2\right]\right)$,
even if $\ket{\psi_{0}}$ and $\ket{\psi_{1}}$ are not in the same
longitude, then $\bar{f}\left(\ket{\psi_{0}}\right)$ and $\bar{f}\left(\ket{\psi_{1}}\right)$
are still the same.

Thus, we could plug $\bar{f}\left(\ket{1}\right)=\bar{f}\left(\ket{2'}\right)=\left[l_{0},r_{0}\right]$
into equation~(\ref{eq:includes}), and get 
\begin{equation}
\left[1-r_{0},1-l_{0}\right]=\left[1,1\right]-\left[l_{0},r_{0}\right]=\left[l_{0},r_{0}\right]\Rightarrow l_{0}+r_{0}=1\,.
\end{equation}
Also, we have 
\begin{equation}
1\in\bar{f}\left(\ket{0'}\right)+\bar{f}\left(\ket{1}\right)+\bar{f}\left(\ket{2'}\right)=\left[2l_{0},2r_{0}\right]\Rightarrow l_{0}\le\frac{1}{2}\le r_{0}\,.
\end{equation}
Thus the lemma is proved. \end{proof} \begin{thm}\label{thm:pi-div-4}Given
$\theta_{0}\in\left[0,\pi/2\right]$ and an interval-valued frame
function~$\bar{f}:S^{2}\rightarrow\mathscr{I}$, if $\bar{f}\left(\ket{\psi}\right)=\imposs$
for every $\ket{\psi}\in\angle^{-1}\left(\left[0,\theta_{0}\right]\right)$,
then $\theta_{0}<\pi/4$.\end{thm} \begin{proof} We are going to
prove by contradiction. Suppose $\theta_{0}\ge\pi/4$, we have $\left[0,\theta_{0}\right]\cap\left[\pi/2-\theta_{0},\pi/2\right]=\left[\pi/2-\theta_{0},\theta_{0}\right]\ne\emptyset$.
Pick any $\ket{\psi_{0}}\in\angle^{-1}\left(\left[\pi/2-\theta_{0},\theta_{0}\right]\right)$,
by lemma~\ref{lemma:imposs-imply-same-value}, we have $\imposs=\bar{f}\left(\ket{\psi_{0}}\right)=\left[l_{0},1-l_{0}\right]$,
where $l_{0}\in\left[0,\pi/2\right]$. A contradiction! \end{proof}
\begin{thm}\label{thm:arcsin-one-over-sqrt-three}Given $\theta_{0}\in\left[\arcsin\left(1/\sqrt{3}\right),\pi/4\right)$
and an interval-valued frame function~$\bar{f}:S^{2}\rightarrow\mathscr{I}$,
if $\bar{f}\left(\ket{\psi}\right)=\imposs$ for every $\ket{\psi}\in\angle^{-1}\left(\left[0,\theta_{0}\right]\right)$,
then $\unknown\in\mathscr{I}$.\end{thm} 
\begin{figure}
\begin{centering}
\includegraphics[scale=0.38]{figureb2} 
\par\end{centering}
\caption{\label{fig:three-dimensional-real-frame-2}This figure is used in
the proof of theorem~\ref{thm:arcsin-one-over-sqrt-three}. The latitude
of the yellow circle is $\theta_{0}\in\left[\arcsin\left(1/\sqrt{3}\right),\pi/4\right)$,
and the latitude of the green circle is $\pi/2-\theta_{0}$.}
\end{figure}

\begin{proof} We could illustrate the situation in figure~\ref{fig:three-dimensional-real-frame-2},
where the latitude of of the yellow circle is $\theta_{0}$ and the
latitude of of the green circle is $\pi/2-\theta_{0}$. Therefore,
$\bar{f}$ should map any state above the yellow circle to $\imposs$
and any states between the green circle and the equator to the same
interval, say, $\left[l_{0},1-l_{0}\right]$, where $0\le l_{0}\le\frac{1}{2}$
by lemma~\ref{lemma:imposs-imply-same-value}.

When $\theta_{0}\ge\arcsin\left(1/\sqrt{3}\right)$, the latitude
of the green circle is $\theta_{0}\le\pi/2-\arcsin\left(1/\sqrt{3}\right)$.
Now consider three states~$\ket{A}$, $\ket{B}$, and $\ket{C}$
whose latitude is $\pi/2-\arcsin\left(1/\sqrt{3}\right)$ so that
they are below the green circle, and these three states form an equilateral
triangle whose center is $H$. Since the latitude of this triangle
is $\pi/2-\arcsin\left(1/\sqrt{3}\right)$, we know $\overline{OH}=1/\sqrt{3}$,
where $O$ is the center of the sphere. By the Pythagorean theorem,
$\overline{HA}=\overline{HB}=\sqrt{\frac{2}{3}}$. Notice that $\triangle HAB$
is an isosceles triangle with $\angle AHB=2\pi/3$, we have $\overline{AB}=\sqrt{2}$.
Since $\overline{OA}=\overline{OB}=1$ and $\overline{AB}=\sqrt{2}$,
$\triangle OAB$ is an isosceles right triangle with $\angle AHB=\pi/2$,
i.e., $\ket{A}$ and $\ket{B}$ are orthogonal. Applying the same
reason on all $\ket{A}$, $\ket{B}$, and $\ket{C}$, $\left\{ \ket{A},\ket{B},\ket{C}\right\} $
is an orthonormal basis.

Since $\left\{ \ket{A},\ket{B},\ket{C}\right\} $ is an orthonormal
basis and $\bar{f}\left(\ket{A}\right)=\bar{f}\left(\ket{B}\right)=\bar{f}\left(\ket{C}\right)=\left[l_{0},1-l_{0}\right]$,
we have the following inequalities by theorem~\ref{thm:interval-valued-frame-function},
\begin{equation}
\left[1,1\right]-\left[l_{0},1-l_{0}\right]\subseteq\left[2l_{0},2-2l_{0}\right]\Rightarrow l_{0}\le0\,.
\end{equation}
Recall that $0\le l_{0}\le\frac{1}{2}$. Therefore, $l_{0}=0$, that
is, $\unknown=\left[l_{0},1-l_{0}\right]\in\mathscr{I}$. \end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Between Quantum and Classical}

This section discusses the relation between quantum and classical.
We first prove that any quantum phenomenon would not be observed if
we could only access commuting projectors since commuting projectors
could be measured simultaneously~\cite{Mermin_1993}. Based on this
fact, we show a quantum IVPM~$\bar{\mu}$ is convex if its restriction
on any orthonormal basis~$\Omega$, $\bar{\mu}^{\Omega}$, is a convex
classical IVPM. Together with the fact that every classical IVPM~$\bar{\mu}^{\Omega}$
is convex when the number of elements in $\Omega\le3$, every quantum
IVPM must be convex on a three-dimensional Hilbert space.

\subsection{Classical and Quantum Probability Measures}

Before diving into the lemmas and proofs, we start from two definitions.

\begin{definition}Given a Hilbert space $\Hilb$, and a set of projections
$\mathcal{S}\subseteq\events$, we define $\sum_{P\in\mathcal{S}}P=\mathbb{0}$
when $\mathcal{S}=\emptyset$ since $\mathbb{0}$ is the additive
identity in operators.\end{definition}

\begin{definition}\label{definition:quantum-classical}Given a Hilbert
space $\Hilb$ with an orthonormal basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$.
\begin{itemize}
\item For any quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$,
we define $\mu^{\Omega}:2^{\Omega}\rightarrow\left[0,1\right]$ by
$\mu^{\Omega}\left(E\right)=\mu\left(\sum_{\ket{j}\in E}\proj{j}\right)$.
\item For any quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$,
we define $\bar{\mu}^{\Omega}:2^{\Omega}\rightarrow\mathscr{I}$ by
$\bar{\mu}^{\Omega}\left(E\right)=\bar{\mu}\left(\sum_{\ket{j}\in E}\proj{j}\right)$.
\end{itemize}
\end{definition}

\begin{lemma}\label{lemma:classical-probability-measures}$\mu^{\Omega}$
and $\bar{\mu}^{\Omega}$ in the previous definition are classical
real-valued and interval-valued probability measures, respectively.\end{lemma}

\begin{proof}When $\mathscr{I}=\set{\left\{ a\right\} }{a\in\left[0,1\right]}$,
$\bar{\mu}$ and $\bar{\mu}^{\Omega}$ are essentially quantum and
classical real-valued probability measures, respectively. Hence, it
is sufficient to prove the interval-valued case.
\begin{itemize}
\item $\bar{\mu}^{\Omega}(\emptyset)=\bar{\mu}\left(\sum_{\ket{j}\in\emptyset}\proj{j}\right)=\bar{\mu}\left(\mathbb{0}\right)=[0,0]$. 
\item $\bar{\mu}^{\Omega}(\Omega)=\bar{\mu}\left(\sum_{\ket{j}\in\Omega}\proj{j}\right)=\bar{\mu}\left(\mathbb{1}\right)=[1,1]$. 
\item For any event $E\in2^{\Omega}$,
\begin{equation}
\bar{\mu}^{\Omega}\left(\Omega\backslash E\right)=\bar{\mu}\left(\mathbb{1}-\sum_{\ket{j}\in E}\proj{j}\right)=\left[1,1\right]-\bar{\mu}^{\Omega}\left(E\right)\,.
\end{equation}
\item For a collection $\left\{ E_{i}\right\} _{i=1}^{N}\subseteq2^{\Omega}$
of pairwise disjoint events, we have
\begin{equation}
\bar{\mu}^{\Omega}\left(\bigcup_{i=1}^{N}E_{i}\right)=\bar{\mu}\left(\sum_{i=1}^{N}\sum_{\ket{j}\in E_{i}}\proj{j}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}^{\Omega}\left(E_{i}\right)\,.
\end{equation}
\end{itemize}
\end{proof}

\begin{definition}\label{def:sub-event-space}Given a Hilbert space
$\Hilb$, a set of event~$\events'\subseteq\events$ is called a
subspace if it satisfies the follow conditions.
\begin{itemize}
\item $\mathbb{0}\in\events'$.
\item $\mathbb{1}\in\events'$.
\item For any projection $P\in\events'$, we have $\mathbb{1}-P\in\events'$.
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\events'$,
we have $\sum_{i=1}^{N}P_{i}\in\events'$.
\end{itemize}
By replacing every occurrence of $\events$ by $\events'$ in definition~\ref{def:QuantumCryptodeterministicMeasures},
\ref{def:QuantumProbabilitySpace}, and \ref{def:QuantumInterval-valuedProbability},
it is straightforward to define a quantum cryptodeterministic, real-valued
probability, and interval-valued probability measure on $\events'$.
Also, $\events'$ is called convex if we have $P_{1}P_{2}\in\events'$
for a pair of mutually commuting projections $P_{1}$ and $P_{2}\in\events'$.\end{definition}

\begin{thm}\label{thm:sub-event-space}Given a Hilbert space~$\Hilb$
and a mutually commuting sub-event space~$\events'\subseteq\events$,
there is a set~$\Omega$, another commuting sub-event space~$\widehat{\events}$
with $\events'\subseteq\widehat{\events}\subseteq\events$, and a
bijection~$\tau:\widehat{\events}\rightarrow2^{\Omega}$ such that:
\begin{enumerate}
\item \label{enu:sub-event-space-real-quantum2classical}If $\mu':\events'\rightarrow\left[0,1\right]$
is a quantum real-valued probability measure, then there is a classical
real-valued probability measure\nb$\mu:2^{\Omega}\rightarrow\left[0,1\right]$
such that $\mu'\left(P\right)=\mu\left(\tau\left(P\right)\right)$
for all $P\in\events'$.
\item \label{enu:sub-event-space-interval-quantum2classical}If $\bar{\mu}':\events'\rightarrow\mathscr{I}$
is a quantum interval-valued probability measure, then there is a
classical interval-valued probability measure\nb$\bar{\mu}:2^{\Omega}\rightarrow\mathscr{I}$
such that $\bar{\mu}'\left(P\right)=\bar{\mu}\left(\tau\left(P\right)\right)$
for all $P\in\events'$.
\item \label{enu:sub-event-space-real-classical2quantum}If $\mu:2^{\Omega}\rightarrow\left[0,1\right]$
is a classical real-valued probability measure, then $\mu\circ\tau:\widehat{\events}\rightarrow\left[0,1\right]$
is a QRVPM.
\item \label{enu:sub-event-space-Gleason}Given a QRVPM\nb$\widehat{\mu}:\widehat{\events}\rightarrow\left[0,1\right]$,
there is a density matrix~$\rho$ such that $\mu_{\rho}^{\mathrm{B}}\left(P\right)=\widehat{\mu}\left(P\right)$
for all $P\in\widehat{\events}$.
\item $\rho$ in the previous case may not be unique.
\item The bijection~$\tau:\widehat{\events}\rightarrow2^{\Omega}$ preserves
the following operations.
\begin{itemize}
\item $\tau\left(\mathbb{0}\right)=\emptyset$.
\item $\tau\left(\mathbb{1}\right)=\Omega$.
\item For any projection $P\in\widehat{\events}$, we have $\tau\left(\mathbb{1}-P\right)=\Omega\backslash\tau\left(P\right)$.
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\widehat{\events}$,
we have $\tau\left(\sum_{i=1}^{N}P_{i}\right)=\bigcup_{i=1}^{N}\tau\left(P_{i}\right)$.
\item For a pair of mutually commuting projections $P_{1}$ and $P_{2}\in\events'$,
we have $\tau\left(P_{1}P_{2}\right)=\tau\left(P_{1}\right)\cap\tau\left(P_{2}\right)$.
\end{itemize}
\end{enumerate}
\end{thm}

\begin{proof}
\begin{enumerate}
\item Since this case is a special case of the next case, it is sufficient
to prove the next case. 
\item Since $\events'$ is a set of mutually commuting projections, they
can be diagonalized by a common orthonormal basis\nb$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$.
Pick arbitrary projector\nb$P\in\events'$, its eigenvalues are $0$
or $1$. Select its eigenvectors whose eigenvalues are $1$ into a
set\nb$\tau\left(P\right)\subseteq\Omega$. Then, we have $P=\sum_{\ket{j}\in\tau\left(P\right)}\proj{j}$,
$\widehat{\events}=\set{\sum_{\ket{j}\in E}\proj{j}}{E\subseteq\Omega}$,
and $\tau$ is a bijection between $\widehat{\events}$ and $2^{\Omega}$.
By definition\nb\ref{definition:quantum-classical}, $\bar{\mu}'\left(P\right)=\bar{\mu}^{\Omega}\left(\tau\left(P\right)\right)$,
and $\bar{\mu}^{\Omega}$ is a classical interval-valued probability
measure by lemma\nb\ref{lemma:classical-probability-measures}.
\item Trivial.
\item If we pick $\rho=\sum_{j=1}^{d}\widehat{\mu}\left(\proj{j}\right)\proj{j}$,
then $\rho$ is a density matrix because\nb\cite{544199}
\begin{eqnarray}
\Tr\left(\rho\right) & = & \sum_{j=1}^{d}\widehat{\mu}\left(\proj{j}\right)\Tr\left(\proj{j}\right)=1\,.
\end{eqnarray}
Also, 
\begin{eqnarray}
 &  & \mu_{\rho}^{\mathrm{B}}\left(\proj{i}\right)=\sum_{j=1}^{N}\widehat{\mu}\left(\proj{j}\right)\mu_{\ket{j}}^{\mathrm{\mathrm{B}}}\left(\proj{i}\right)=\widehat{\mu}\left(\proj{i}\right)\textrm{ for all }i\,,\\
 &  & \mu_{\rho}^{\mathrm{B}}\left(P_{0}\right)+\mu_{\rho}^{\mathrm{B}}\left(P_{1}\right)\overset{\textrm{IH}}{=}\widehat{\mu}\left(P_{0}\right)+\widehat{\mu}\left(P_{1}\right)=\widehat{\mu}\left(P_{0}+P_{1}\right)
\end{eqnarray}
for orthogonal\nb$P_{0},P_{1}\in\events$. Therefore, we have $\mu_{\rho}^{\mathrm{B}}\left(P\right)=\widehat{\mu}\left(P\right)$
for all $P\in\widehat{\events}$ by induction.
\item Consider a Hilbert space\nb$\Hilb$ of dimension 2, a mutually commuting
sub-event space~$\widehat{\events}=\left\{ \mathbb{0},\proj{0},\proj{1},\mathbb{1}\right\} $,
and a QRVPM\nb$\widehat{\mu}:\widehat{\events}\rightarrow\left[0,1\right]$
such that $\widehat{\mu}\left(\proj{0}\right)=\widehat{\mu}\left(\proj{1}\right)=\frac{1}{2}$.
Then, $\proj{\ps}$ and $\proj{\minus}$ are both density matrices,
and 
\begin{eqnarray*}
 &  & \mu_{\ket{\ps}}^{\mathrm{B}}\left(\proj{0}\right)=\mu_{\ket{\minus}}^{\mathrm{B}}\left(\proj{0}\right)=\frac{1}{2}=\widehat{\mu}\left(\proj{0}\right)\,,\\
 &  & \mu_{\ket{\ps}}^{\mathrm{B}}\left(\proj{1}\right)=\mu_{\ket{\minus}}^{\mathrm{B}}\left(\proj{1}\right)=\frac{1}{2}=\widehat{\mu}\left(\proj{1}\right)\,.
\end{eqnarray*}
\item Trivial.
\end{enumerate}
\end{proof}

Given an orthonormal basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$,
if we consider sub-event space generated by $\Omega$, $\events^{\Omega}=\set{\sum_{\ket{j}\in E}\proj{j}}{E\subseteq\Omega}$,
then any quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
restricting on $\events^{\Omega}$, denoted by $\bar{\mu}|_{\events^{\Omega}}$,
behaves classically. In particular, if $\bar{\mu}$ is convex, each
$\bar{\mu}|_{\events^{\Omega}}$ is convex and has a non-empty core.
However, the core elements of each $\bar{\mu}|_{\events^{\Omega}}$
may not be glued together, and gives a core element of $\bar{\mu}$.
The inability of gluing core elements among commuting $\events^{\Omega}$
to non-commuting $\events$ is consistent with the idea of the Kochen-Specker
theorem, that each triple of commuting observables could be colored
well, but we cannot color all observables which are not all commuting.
This is also consistent with the assertion that mutually commuting
observables could be measured simultaneously~\cite{Mermin_1993}.
In another word, if Alice and Bob in Bell's theorem can only pick
from a set of mutually commuting observables, they could never observe
any ``quantum,`` and we cannot eliminate the local hidden variable
theory in this setting.

\subsection{Describe Classical World by Quantum Probability Measures with a Commuting
Sub-event Space}

Since there is a correspondence between classical probability measures
and quantum ones on a commuting sub-event space, we can use the quantum
language to describe the classical world given an orthonormal basis.

\begin{definition}Given a Hilbert space~$\Hilb$, an orthonormal
basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$, and a classical
real-valued probability measures~$\pmeas:2^{\Omega}\rightarrow\left[0,1\right]$
on the sample space~$\Omega$, the density matrix~$\sum_{j=0}^{d-1}\pmeas\left(\left\{ j\right\} \right)\proj{j}$
is called classical with respect to $\Omega$.\end{definition}

\begin{definition}\label{def:consistent-state}A state~$\rho$ is
consistent with $\bar{\mu}$ on a projector~$P$ if $\mu_{\rho}^{\mathrm{B}}\left(P\right)\in\bar{\mu}\left(P\right)$,
and denoted by $\rho\in\coreBorn\left(\bar{\mu},\left\{ P\right\} \right)$.
Moreover, $\rho\in\coreBorn\left(\bar{\mu},S\right)$ means $\rho$
is consistent with $\bar{\mu}$ on every projector in $S$. Notice
that $\rho\in\coreBorn\left(\bar{\mu},S\right)$ if and only if $\mu_{\rho}^{\mathrm{B}}\in\mathrm{core}\left(\bar{\mu},S\right)$.\end{definition}

Then, we can rephrase theorem~\ref{thm:Shapley}, in the language
of quantum IVPM with an orthonormal basis.

\begin{cor}For every convex quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$,
if a sub-event space~$\events'\subseteq\events$ commutes, then $\coreBorn\left(\bar{\mu},\events'\right)\ne\emptyset$.\end{cor}

\begin{proof}Since $\events'$ is a set of mutually commuting projections,
they can be diagonalized by a common orthonormal basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$.
By theorem~\ref{thm:sub-event-space}, we can find a bijection~$\tau:\widehat{\events}\rightarrow2^{\Omega}$
and a classical IVPM~$\bar{\mu}':2^{\Omega}\rightarrow\mathscr{I}$
such that $\bar{\mu}\left(P\right)=\bar{\mu}'\left(\tau\left(P\right)\right)$
for all $P\in\events'$. Since $\tau$ is a bijection, $\bar{\mu}'$
is convex. By theorem~\ref{thm:Shapley}, there is a classical probability
measure~$\pmeas:2^{\Omega}\rightarrow\left[0,1\right]$ such that
\begin{eqnarray*}
 &  & \pmeas\in\mathrm{core}\left(\bar{\mu}'\right)\\
 & \Rightarrow & \pmeas\left(E\right)\in\bar{\mu}'\left(E\right)\textrm{ for all }E\in2^{\Omega}\\
 & \Rightarrow & \pmeas\left(\tau\left(P\right)\right)\in\bar{\mu}'\left(\tau\left(P\right)\right)=\bar{\mu}\left(P\right)\textrm{ for all }P\in\widehat{\events}\\
 & \Rightarrow & \pmeas\left(\tau\left(P\right)\right)\in\bar{\mu}\left(P\right)\textrm{ for all }P\in\widehat{\events}
\end{eqnarray*}
By theorem~\ref{thm:sub-event-space}, there is a density matrix~$\rho$
such that $\mu_{\rho}^{\mathrm{B}}\left(P\right)=\mu'\left(\tau\left(P\right)\right)$
for all $P\in\events'$. Therefore, $\mu_{\rho}^{\mathrm{B}}\in\mathrm{core}\left(\bar{\mu},\events'\right)$,
i.e., $\rho\in\coreBorn\left(\bar{\mu},\events'\right)$.\end{proof}

\subsection{Application on Convexity}

The rest of the section focus on convexity.

\begin{lemma}\label{lemma:classical-quantum-convex}Given a quantum
IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$, if $\bar{\mu}^{\Omega}:2^{\Omega}\rightarrow\mathscr{I}$
are all convex for any orthonormal basis~$\Omega$, then $\bar{\mu}$
is convex.\end{lemma}

\begin{proof} Pick two commuting events~$P_{0},P_{1}\in\events$.
Since they are commuting, they can be diagonalized by a common orthonormal
basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$. Since they
are projectors, the eigenvalues of them are $0$ or $1$ so that we
can express them as $P_{0}=\sum_{\ket{j}\in E_{0}}\proj{j}$ and $P_{1}=\sum_{\ket{j}\in E_{1}}\proj{j}$,
where $E_{0},E_{1}\subseteq\Omega$. Then, 
\begin{equation}\eqalign{  
& \bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)= \bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)\\ 
\subseteq & \bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)= \bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\,. 
}\end{equation}
\end{proof}

\begin{lemma}\label{lemma:classical-convex-condition-lemma}Given
a classical IVPM~$\bar{\mu}^{\Omega}:2^{\Omega}\rightarrow\mathscr{I}$,
the following statements are true:
\begin{enumerate}
\item If $E_{0}\subseteq E_{1}$, then $\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)$.
\item If $\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)$,
then $\bar{\mu}^{\Omega}\left(\overline{E_{0}}\cup\overline{E_{1}}\right)+\bar{\mu}^{\Omega}\left(\overline{E_{0}}\cap\overline{E_{1}}\right)\subseteq\bar{\mu}^{\Omega}\left(\overline{E_{0}}\right)+\bar{\mu}^{\Omega}\left(\overline{E_{1}}\right)$,
where $\overline{E_{0}}=\Omega\backslash E_{0}$ and $\overline{E_{1}}=\Omega\backslash E_{1}$.
\end{enumerate}
\end{lemma}

\begin{proof}
\begin{enumerate}
\item $\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)=\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)$.
\item 
\begin{equation}\eqalign{ 
 & \bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\\
\Rightarrow & \bar{\mu}^{\Omega}\left(\Omega\backslash\left(\overline{E_{0}}\cap \overline{E_{1}}\right)\right)+\bar{\mu}^{\Omega}\left(\Omega\backslash\left(\overline{E_{0}}\cup \overline{E_{1}}\right)\right)\subseteq\bar{\mu}^{\Omega}\left(\Omega\backslash\left(\overline{E_{0}}\right)\right)+\bar{\mu}^{\Omega}\left(\Omega\backslash\left(\overline{E_{1}}\right)\right)\\
\Rightarrow & 2-\left[\bar{\mu}^{\Omega}\left(\overline{E_{0}}\cap \overline{E_{1}}\right)+\bar{\mu}^{\Omega}\left(\overline{E_{0}}\cup \overline{E_{1}}\right)\right]\subseteq2-\left[\bar{\mu}^{\Omega}\left(\overline{E_{0}}\right)+\bar{\mu}^{\Omega}\left(\overline{E_{1}}\right)\right]\\
\Rightarrow & \bar{\mu}^{\Omega}\left(\overline{E_{0}}\cap \overline{E_{1}}\right)+\bar{\mu}^{\Omega}\left(\overline{E_{0}}\cup \overline{E_{1}}\right)\subseteq\bar{\mu}^{\Omega}\left(\overline{E_{0}}\right)+\bar{\mu}^{\Omega}\left(\overline{E_{1}}\right)\,. 
}\end{equation}
\end{enumerate}
\end{proof}

\begin{lemma}\label{lemma:classical-3-elements}When the number of
elements in $\Omega$ is less than or equal to 3, every classical
IVPM~$\bar{\mu}^{\Omega}:2^{\Omega}\rightarrow\mathscr{I}$ is convex.\end{lemma}

\begin{proof}Given two events~$E_{0},E_{1}\subseteq\Omega$, we
want to verify 
\begin{equation}
\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\,.\label{eq:classical-convex}
\end{equation}
When the number of elements in $\Omega\le3$, one of the following
must happen
\begin{enumerate}
\item When $E_{0}\cap E_{1}=\emptyset$, equation~(\ref{eq:classical-convex})
is a special case of equation~(\ref{eq:classical-include}). 
\item When $E_{0}\cup E_{1}=\Omega$, we have $\overline{E_{0}}\cap\overline{E_{1}}=\emptyset$,
where $\overline{E_{0}}=\Omega\backslash E_{0}$ and $\overline{E_{1}}=\Omega\backslash E_{1}$.
By equation~(\ref{eq:classical-include}), we have $\bar{\mu}^{\Omega}\left(\overline{E_{0}}\cup\overline{E_{1}}\right)+\bar{\mu}^{\Omega}\left(\overline{E_{0}}\cap\overline{E_{1}}\right)\subseteq\bar{\mu}^{\Omega}\left(\overline{E_{0}}\right)+\bar{\mu}^{\Omega}\left(\overline{E_{1}}\right)$.
Then, lemma~\ref{lemma:classical-convex-condition-lemma} proves
equation~(\ref{eq:classical-convex}).
\item When $E_{0}\subseteq E_{1}$ or $E_{1}\subseteq E_{0}$, equation~(\ref{eq:classical-convex})
is also followed by lemma~\ref{lemma:classical-convex-condition-lemma}.
\end{enumerate}
so that $\bar{\mu}^{\Omega}$ is convex.\end{proof}

\begin{lemma}\label{lemma:quantum-convex-condition-lemma}Given a
QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$, the following statements
are true:
\begin{enumerate}
\item If $P_{1}=P_{0}+P_{2}$, then $\bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)$.
\item If $\bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)$,
then $\bar{\mu}\left(\overline{P_{0}}\vee\overline{P_{1}}\right)+\bar{\mu}\left(\overline{P_{0}}\overline{P_{1}}\right)\subseteq\bar{\mu}\left(\overline{P_{0}}\right)+\bar{\mu}\left(\overline{P_{1}}\right)$,
where $\overline{P_{0}}=\Omega\backslash P_{0}$ and $\overline{P_{1}}=\Omega\backslash P_{1}$.
\end{enumerate}
\end{lemma}

\begin{proof}The direct consequence of lemma~\ref{lemma:classical-convex-condition-lemma}
and theorem~\ref{thm:sub-event-space}.\end{proof}

\begin{thm}\label{thm:convex-3}Given a Hilbert space~$\Hilb$ of
dimension~$3$, every quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
is convex.\end{thm}

\begin{proof}The direct consequence of lemma~\ref{lemma:classical-quantum-convex}
and lemma~\ref{lemma:classical-3-elements}. \end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{From QRVPM to QIVPM\label{sec:Real-and-Interval}}

Recall that given a set of intervals\nb$\mathscr{I}$, we want to
know whether there is an QIVPM mapping to $\mathscr{I}$. If there
is a QIVPM\nb$\bar{\mu}:\events\rightarrow\mathscr{I}$, we want
to further verify whether $\bar{\mu}$ has a non-empty core. Usually,
the easiest way to prove the existence is to construct an example.
As in example~\ref{ex:three-dimensional-5-value}, constructing a
QIVPM directly is long and tedious. In contrast, a quantum probability
measure\nb$\mu:\events\rightarrow\left[0,1\right]$ can be easily
constructed by the Born rule. Therefore, it would be desirable if
we can construct a QIVPM\nb$\bar{\mu}$ from $\mu$, and expect $\mu$
is in the core of $\bar{\mu}$.

Notice that the range of $\mu$ and $\bar{\mu}$ are different. If
there is a ``homomorphism''\nb$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
preserving the desiring properties of the ranges, then we could easily
inducing a map from quantum real-valued probability measures to QIVPMs
by function composition\nb$\events\overset{\mu}{\rightarrow}\left[0,1\right]\overset{\iota}{\rightarrow}\mathscr{I}$.
This idea gives the following definition.

\begin{definition}Given $\mathscr{I}$ is a collection of intervals,
we say a interval-valued function\nb$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is additive if $\iota$ satisfies the following properties. 
\begin{itemize}
\item $\iota\left(0\right)=\left[0,0\right]$. 
\item $\iota\left(1\right)=\left[1,1\right]$. 
\item For any $x\in\left[0,1\right]$, $\iota\left(1-x\right)=\left[1,1\right]-\iota\left(x\right)$. 
\item For a set\nb$\left\{ x_{i}\right\} _{i=1}^{N}\subseteq\left[0,1\right]$,
if $\sum_{i=1}^{N}x_{i}\in\left[0,1\right]$, then $\iota\left(\sum_{i=1}^{N}x_{i}\right)\subseteq\sum_{i=1}^{N}\iota\left(x_{i}\right)$. 
\end{itemize}
Moreover, $\iota$ is called convex if for any three numbers\nb$x_{0}$,
$x_{1}$, and $x_{2}\in\left[0,1\right]$, $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$
implies $\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\subseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)$.\end{definition}

Although a interval-valued function\nb$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
may look exotic, its left-end and right-end\nb$\left[\iotal\left(x\right),\iotar\left(x\right)\right]=\iota\left(x\right)$
are just usual functions from a unit interval to another unit interval.
For an additive interval-valued function, $\iotal$ is superadditive,
i.e., $\iotal\left(x_{0}+x_{1}\right)\ge\iotal\left(x_{0}\right)+\iotal\left(x_{1}\right)$,
and $\iotar$ is subadditive, i.e., $\iotar\left(x_{0}+x_{1}\right)\le\iotar\left(x_{0}\right)+\iotar\left(x_{1}\right)$\nb\cite{PecaricTong1992}.
For a convex interval-valued function, $\iotal$ is Wright-convex,
i.e., $\iotal\left(x_{0}+x_{1}+x_{2}\right)+\iotal\left(x_{2}\right)\ge\iotal\left(x_{0}+x_{2}\right)+\iotal\left(x_{1}+x_{2}\right)$,
and $\iotar$ is Wright-concave, i.e., $\iotar\left(x_{0}+x_{1}+x_{2}\right)+\iotar\left(x_{2}\right)\le\iotar\left(x_{0}+x_{2}\right)+\iotar\left(x_{1}+x_{2}\right)$\nb\cite{Wright1954,RobertsVarberg1973,PecaricTong1992,Brzdek_2014}.
Also, if $\mathscr{I}$ contains only finite number of intervals,
$\iotal$ and $\iotar$ must be step functions. After we define interval-valued
function, we are ready to construct a QIVPM from a quantum probability
measure.

\begin{lemma}\label{lemma:real2interval}Consider an additive interval-valued
function\nb$\iota:\left[0,1\right]\rightarrow\mathscr{I}$. Then,
given a quantum probability measure\nb$\mu:\events\rightarrow\left[0,1\right]$,
a function\nb$\bar{\mu}:\events\rightarrow\mathscr{I}$ defined by
$\bar{\mu}\left(P\right)=\iota\left(\mu\left(P\right)\right)$ is
a quantum IVPM. Moreover, if $\iota$ is convex, then $\bar{\mu}$
is convex.\end{lemma}

\begin{proof}~
\begin{itemize}
\item $\bar{\mu}(\mathbb{0})=\iota\left(\mu(\mathbb{0})\right)=\iota\left(0\right)=\left[0,0\right]$.
\item $\bar{\mu}(\mathbb{1})=\iota\left(\mu(\mathbb{1})\right)=\iota\left(1\right)=\left[1,1\right]$.
\item For any projection $P$, $\bar{\mu}\left(\mathbb{1}-P\right)=\iota\left(1-\mu\left(P\right)\right)=\left[1,1\right]-\bar{\mu}\left(P\right)$.
\item For a set of mutually orthogonal projections\nb$\left\{ P_{i}\right\} _{i=1}^{N}$,
we have $\sum_{i=1}^{N}\mu\left(P_{i}\right)\in\left[0,1\right]$
so that $\bar{\mu}\left(\sum_{i=1}^{N}P_{i}\right)=\iota\left(\sum_{i=1}^{N}\mu\left(P_{i}\right)\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(P_{i}\right)$.
\item For convexity part, given a pair of commuting operators\nb$P_{0}$
and $P_{1}$, they and $P_{0}\vee P_{1}$ can be expressed by the
summation among $P_{0}\left(\mathbb{1}-P_{1}\right)$, $\left(\mathbb{1}-P_{0}\right)P_{1}$,
and $P_{0}P_{1}$. Hence, 
\begin{equation}\eqalign{ 
 & \bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\\
= & \iota\left(\mu\left(P_{0}\left(\mathbb{1}-P_{1}\right)\right)+\mu\left(\left(\mathbb{1}-P_{0}\right)P_{1}\right)+\mu\left(P_{0}P_{1}\right)\right)+\iota\left(\mu\left(P_{0}P_{1}\right)\right)\\
\subseteq&\iota\left(\mu\left(P_{0}\left(\mathbb{1}-P_{1}\right)\right)+\mu\left(P_{0}P_{1}\right)\right)+\iota\left(\mu\left(\left(\mathbb{1}-P_{0}\right)P_{1}\right)+\mu\left(P_{0}P_{1}\right)\right)\\
= & \bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right) 
}\end{equation}
\end{itemize}
\end{proof}

Since we want to consider QIVPMs induced by the Born rule, it is easier
to work with the states instead of the probability measures, which
gives us the following definition.

\begin{definition}A state\nb$\rho$ is consistent with $\bar{\mu}$
on a projector\nb$P$ if $\mu_{\rho}^{\mathrm{B}}\left(P\right)\in\bar{\mu}\left(P\right)$,
i.e., $\mu_{\rho}^{\mathrm{B}}\in\mathrm{core}\left(\bar{\mu}|_{\left\{ P\right\} }\right)$,
and denoted by $\rho\in\coreBorn\left(\bar{\mu},\left\{ P\right\} \right)$.
Moreover, $\rho\in\coreBorn\left(\bar{\mu},S\right)$ means $\rho$
is consistent with $\bar{\mu}$ on every projector in $S$, i.e.,
$\mu_{\rho}^{\mathrm{B}}\in\mathrm{core}\left(\bar{\mu}|_{S}\right)$.\end{definition}

\begin{lemma}Consider a set of intervals\nb$\mathscr{I}$ with an
additive interval-valued function\nb$\iota:\left[0,1\right]\rightarrow\mathscr{I}$.
For any normalized pure state\nb$\ket{\phi}$, $\iota\circ\mu_{\phi}^{\mathrm{\mathrm{B}}}$
is a quantum IVPM such that $\coreBorn\left(\iota\circ\mu_{\phi}^{\mathrm{\mathrm{B}}}\right)$
has an unique element. Obviously, if $\iota$ is convex, so does $\iota\circ\mu_{\phi}^{\mathrm{\mathrm{B}}}$.\end{lemma}

\begin{proof}$\left(\iota\circ\mu_{\phi}^{\mathrm{\mathrm{B}}}\right)\left(\proj{\phi}\right)=\iota\left(\mu_{\phi}^{\mathrm{\mathrm{B}}}\left(\proj{\phi}\right)\right)=\iota\left(1\right)=\left[1,1\right]$,
i.e., for any quantum probability measure~$\mu\in\coreBorn\left(\iota\circ\mu_{\phi}^{\mathrm{\mathrm{B}}}\right)$,
we have $\mu\left(\proj{\phi}\right)=1$. Therefore, $\mu=\mu_{\phi}^{\mathrm{\mathrm{B}}}$.\end{proof}

\begin{example}\label{ex:regular-pair}For any positive integer~$n$
and $\delta\in\left[0,1\right]$, there is an additive interval-valued
function\nb$\iota_{n,\delta}:\left[0,1\right]\rightarrow\mathscr{I}_{n,\delta}$
defined by
\begin{equation}\eqalign{ 
\eqalign{\mathscr{I}_{n,\delta} =& \set{\left[\frac{k-\delta}{n},\frac{k+\delta}{n}\right]}{k=1,\ldots,n-1}\\
&\cup\set{\left[\frac{k}{n},\frac{k+1}{n}\right]}{k=0,\ldots,n-1}\cup\left\{\left[0,0\right],\left[1,1\right]\right\}}\\ 
\iota_{n,\delta}\left(x\right) = \cases{ 
\left[x,x\right]\,, &if $x=0$ or $x=1$ ;\\ 
\left[\frac{k-\delta}{n},\frac{k+\delta}{n}\right]\,, &if $x=\frac{k}{n}\in\left(0,1\right)$ ;\\ 
\left[\frac{k}{n},\frac{k+1}{n}\right]\,, &if $\frac{k}{n}<x<\frac{k+1}{n}$ . 
}
}\end{equation}
When we use the same notation in other place in this paper, we may
write $\iota_{n}:\left[0,1\right]\rightarrow\mathscr{I}_{n}$ to represent
$\iota_{n,0}:\left[0,1\right]\rightarrow\mathscr{I}_{n,0}$. Notice
that the usual $\left\{ \necess,\imposs,\unknown\right\} $ is $\mathscr{I}_{1}$.
Also, let $\mathscr{I}_{\infty}=\set{\left\{ x\right\} }{x\in\left[0,1\right]}$
and $\iota_{\infty}\left(x\right)=\left\{ x\right\} $. Of course,
$\iota_{\infty}:\left[0,1\right]\rightarrow\mathscr{I}_{\infty}$
is a convex interval-valued function.\end{example}

\begin{proof}~
\begin{itemize}
\item Since $\delta\in\left[0,1\right]$ and $\iota_{n,\delta}\left(x\right)=\left[\frac{k-\delta}{n},\frac{k+\delta}{n}\right]$
only when $0<x=\frac{k}{n}<1$, we have $\iota_{n,\delta}\left(x\right)\subseteq\left[0,1\right]$.
\item $\iota_{n,\delta}\left(0\right)=\left[0,0\right]$. 
\item $\iota_{n,\delta}\left(1\right)=\left[1,1\right]$. 
\item When $x=0$ or $x=1$, we have $\iota_{n,\delta}\left(1-x\right)=\left[1-x,1-x\right]=\left[1,1\right]-\iota_{n,\delta}\left(x\right)$.\\
When $x=k/n\in\left(0,1\right)$, we have 
\begin{equation}
\iota_{n,\delta}\left(1-x\right)=\left[\frac{n-k-\delta}{n},\frac{n-k+\delta}{n}\right]=\left[1,1\right]-\iota_{n,\delta}\left(x\right)\,.
\end{equation}
When $k/n<x<\left(k+1\right)/n$, we have $1-k/n>1-x>1-\left(k+1\right)/n$
so that 
\begin{equation}
\iota_{n,\delta}\left(1-x\right)=\left[1-\frac{k+1}{n},1-\frac{k}{n}\right]=\left[1,1\right]-\iota_{n,\delta}\left(x\right)\,.
\end{equation}
\item It is sufficient to prove the case when $N=2$, i.e., for a set $\left\{ x_{1},x_{2}\right\} \subseteq\left[0,1\right]$,
if $x_{1}+x_{2}\in\left[0,1\right]$, then $\iota_{n,\delta}\left(x_{1}+x_{2}\right)\subseteq\iota_{n,\delta}\left(x_{1}\right)+\iota_{n,\delta}\left(x_{2}\right)$,
and the rest can be done by the induction. 
\begin{itemize}
\item When $x_{1}=0$, we have $\iota_{n,\delta}\left(x_{1}+x_{2}\right)=\iota_{n,\delta}\left(x_{2}\right)=\iota_{n,\delta}\left(x_{1}\right)+\iota_{n,\delta}\left(x_{2}\right)$.
\item When $x_{1}=k_{1}/n\in\left(0,1\right)$, we have 
\begin{equation}
\iota_{n,\delta}\left(x_{1}+x_{2}\right)\subseteq\left[\frac{k_{1}}{n},\frac{k_{1}}{n}\right]+\iota_{n,\delta}\left(x_{2}\right)\subseteq\iota_{n,\delta}\left(x_{1}\right)+\iota_{n,\delta}\left(x_{2}\right)\,,
\end{equation}
where $\iota_{n,\delta}\left(x_{1}+x_{2}\right)$ is a proper subset
of $\left[\frac{k_{1}}{n},\frac{k_{1}}{n}\right]+\iota_{n,\delta}\left(x_{2}\right)$
only when $x_{1}+x_{2}=1$.
\item When $k_{1}/n<x_{1}<\left(k_{1}+1\right)/n$ and $k_{2}/n<x_{2}<\left(k_{2}+1\right)/n$,
we have 
\begin{equation}
\frac{k_{1}+k_{2}}{n}<x_{1}+x_{2}<\frac{k_{1}+k_{2}+2}{n}
\end{equation}
so that 
\begin{equation}\eqalign{ 
& \iota_{n,\delta}\left(x_{1}+x_{2}\right)\subseteq\left[\frac{k_{1}+k_{2}}{n},\frac{k_{1}+k_{2}+2}{n}\right]\\ 
= &\left[\frac{k_{1}}{n},\frac{k_{1}+1}{n}\right]+\left[\frac{k_{2}}{n},\frac{k_{2}+1}{n}\right]=\iota_{n,\delta}\left(x_{1}\right)+\iota_{n,\delta}\left(x_{2}\right)\,. 
}\end{equation}
\end{itemize}
\end{itemize}
\end{proof}

\begin{example}\label{ex:equilateral}Consider the following intervals
\begin{equation}\eqalign{ 
\widehat{\mathscr{I}}_{n}= & \set{\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\\
 & \cup\{\imposs,\necess\}
}\end{equation}
which are overlapping intervals with length~$1/n$ except for the
intervals with length~$\left(2n\right)^{-1}$ at each end. As shown
in Table~\ref{tab:I_n}, for each interval~$[\ell,r]$, we define
$\textrm{Int}\left([\ell,r]\right)$, the non-overlapping interior
part of $[\ell,r]$.
\begin{table}
\caption{\label{tab:I_n}Intervals in $\widehat{\mathscr{I}}_{n}$, their lengths,
and their interiors, where $2\mid k$ means $k$ is even, and $2\nmid k$
means $k$ is odd. Although intervals~$[\ell,r]$ are overlapping,
their interior~$\textrm{Int}\left([\ell,r]\right)$ are non-overlapping.}
\centering{}%
\begin{tabular}{cccc}
\toprule 
\addlinespace
$k$ & $[\ell,r]$ & $r-\ell$ & $\textrm{Int}\left([\ell,r]\right)$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
 & $\imposs$ & $0$ & $\imposs$\tabularnewline\addlinespace
\midrule 
\addlinespace
$0$ & $\left[0,\tfrac{1}{2n}\right]$ & $\tfrac{1}{2n}$ & $\left(0,\tfrac{0.5}{2n}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
\begin{tabular}{c}
$0<k<2n$\tabularnewline
and $2\nmid k$\tabularnewline
\end{tabular} & $\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]$ & $\tfrac{1}{n}$ & $\left(\tfrac{k-0.5}{2n},\tfrac{k+0.5}{2n}\right)$\tabularnewline\addlinespace
\midrule 
\addlinespace
\begin{tabular}{c}
$0<k<2n$\tabularnewline
and $2\mid k$\tabularnewline
\end{tabular} & \multirow{1}{*}{$\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]$} & $\tfrac{1}{n}$ & $\left[\tfrac{k-0.5}{2n},\tfrac{k+0.5}{2n}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$2n$ & $\left[\tfrac{2n-1}{2n},1\right]$ & $\tfrac{1}{2n}$ & $\left[\tfrac{2n-0.5}{2n},1\right)$\tabularnewline\addlinespace
\midrule 
\addlinespace
 & $\necess$ & $0$ & $\necess$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
Since their interior parts are non-overlapping, we can define a function~$\widehat{\iota}_{n}:\left[0,1\right]\rightarrow\widehat{\mathscr{I}}_{n}$
by $\widehat{\iota}_{n}\left(x\right)=[\ell,r]$ if $x\in\textrm{Int}\left([\ell,r]\right)$
to map a real-valued probability to an interval-valued one. Table~\ref{tab:I_2}
lists a special case $\widehat{\mathscr{I}}_{2}$ to illustrate these
intervals and their interior parts more clearly.
\begin{table}
\caption{\label{tab:I_2}Intervals in $\widehat{\mathscr{I}}_{2}$, , their
lengths, and their interiors. Although intervals~$[\ell,r]$ are
overlapping, their interior~$\textrm{Int}\left([\ell,r]\right)$
are non-overlapping.}
\centering{}%
\begin{tabular}{cccc}
\toprule 
\addlinespace
$k$ & $[\ell,r]$ & $r-\ell$ & $\textrm{Int}\left([\ell,r]\right)$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
 & $\imposs$ & $0$ & $\imposs$\tabularnewline\addlinespace
\midrule 
\addlinespace
$0$ & $\left[0,\tfrac{1}{4}\right]$ & $\tfrac{1}{4}$ & $\left(0,\tfrac{1}{8}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$1$ & $\left[0,{\textstyle \tfrac{1}{2}}\right]$ & $\tfrac{1}{2}$ & $\left(\tfrac{1}{8},\tfrac{3}{8}\right)$\tabularnewline\addlinespace
\midrule 
\addlinespace
$2$ & $\left[\tfrac{1}{4},\tfrac{3}{4}\right]$ & $\tfrac{1}{2}$ & $\left[\tfrac{3}{8},\tfrac{5}{8}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$3$ & $\left[\tfrac{1}{2},1\right]$ & $\tfrac{1}{2}$ & $\left(\tfrac{5}{8},\tfrac{7}{8}\right)$\tabularnewline\addlinespace
\midrule 
\addlinespace
$4$ & $\left[\tfrac{3}{4},1\right]$ & $\tfrac{1}{4}$ & $\left[\tfrac{7}{8},1\right)$\tabularnewline\addlinespace
\midrule 
\addlinespace
 & $\necess$ & $0$ & $\necess$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
\end{example}

\begin{proof}~
\begin{itemize}
\item $\widehat{\iota}_{n}\left(0\right)=\left[0,0\right]$. 
\item $\widehat{\iota}_{n}\left(1\right)=\left[1,1\right]$. 
\item When $x=0$ or $x=1$, we have $\widehat{\iota}_{n}\left(1-x\right)=\left[1-x,1-x\right]=\left[1,1\right]-\widehat{\iota}_{n}\left(x\right)$.
Otherwise, since $2\mid k\Leftrightarrow2\mid\left(2n-k\right)$,
we always have
\begin{equation}
\widehat{\iota}_{n}\left(1-x\right)=\left[1-\frac{k+1}{2n},1-\frac{k-1}{2n}\right]\cap\left[0,1\right]=\left[1,1\right]-\widehat{\iota}_{n}\left(x\right)\,.
\end{equation}
\item It is sufficient to prove the case when $N=2$, i.e., for a set $\left\{ x_{1},x_{2}\right\} \subseteq\left[0,1\right]$,
if $x_{1}+x_{2}\in\left[0,1\right]$, then $\widehat{\iota}_{n}\left(x_{1}+x_{2}\right)\subseteq\widehat{\iota}_{n}\left(x_{1}\right)+\widehat{\iota}_{n}\left(x_{2}\right)$,
and the rest can be done by the induction. 
\begin{itemize}
\item When $x_{1}=0$, we have $\widehat{\iota}_{n}\left(x_{1}+x_{2}\right)=\widehat{\iota}_{n}\left(x_{2}\right)=\widehat{\iota}_{n}\left(x_{1}\right)+\widehat{\iota}_{n}\left(x_{2}\right)$.
\item When $x_{1}=1$, we must have $x_{2}=0$ which reduces to the previous
case.
\item When $\widehat{\iota}_{n}\left(x_{1}\right)=\left[\frac{k_{1}-1}{2n},\frac{k_{1}+1}{2n}\right]\cap\left[0,1\right]$
and $\widehat{\iota}_{n}\left(x_{2}\right)=\left[\frac{k_{2}-1}{2n},\frac{k_{2}+1}{2n}\right]\cap\left[0,1\right]$,
we have $k_{1}-0.5\le2nx_{1}\le k_{1}+0.5$ and $k_{2}-0.5\le2nx_{2}\le k_{2}+0.5$
so that $k_{1}+k_{2}-1\le2n\left(x_{1}+x_{2}\right)\le k_{1}+k_{2}+1$.
Hence, no matter whether $x_{1}+x_{2}=1$ or not, we always have 
\begin{equation}\eqalign{ 
& \widehat{\iota}_{n}\left(x_{1}+x_{2}\right)\subseteq\left[\frac{k_{1}+k_{2}-2}{2n},\frac{k_{1}+k_{2}+2}{2n}\right]\cap\left[0,1\right]\\ 
= &\left[\frac{k_{1}-1}{2n},\frac{k_{1}+1}{2n}\right]\cap\left[0,1\right]+\left[\frac{k_{2}-1}{2n},\frac{k_{2}+1}{2n}\right]\cap\left[0,1\right]\\
= &\widehat{\iota}_{n}\left(x_{1}\right)+\widehat{\iota}_{n}\left(x_{2}\right)\,. 
}\end{equation}
\end{itemize}
\end{itemize}
\end{proof}

Although a convex QIVPM can be constructed by a convex interval-valued
function easily, its range either is $\left\{ \necess,\imposs,\unknown\right\} $
or contains infinitely many intervals. The reason is that a Wright-convex
function is midpoint-convex or Jensen-convex, i.e., 
\begin{equation}
\frac{\iotal\left(x_{0}\right)+\iotal\left(x_{1}\right)}{2}\ge\iotal\left(\frac{x_{0}+x_{1}}{2}\right)\,,
\end{equation}
and a midpoint-convex function is continuous on an open interval if
we cannot use the Axiom of Choices\nb\cite{Blumberg_1919,Sierpinski1920,RobertsVarberg1973}.

Even if an additive interval-valued function\nb$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is not convex, it may still induce a convex QIVPM. In particular,
every QIVPM on a three-dimensional Hilbert space is convex by theorem~\ref{thm:convex-3}.
The next natural question is whether a non-convex interval-valued
function always induces some non-convex QIVPM, and the answer is ``yes''
by the following lemma.

\begin{lemma}Given an additive but non-convex interval-valued function\nb$\iota:\left[0,1\right]\rightarrow\mathscr{I}$,
there is a non-convex quantum IVPM\nb$\bar{\mu}:\events\rightarrow\mathscr{I}$.\end{lemma}

\begin{proof}Since $\left(\mathscr{I},\iota\right)$ is additive
but non-convex, there are three numbers\nb$x_{0}$, $x_{1}$, and
$x_{2}\in\left[0,1\right]$ such that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$
but $\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)$.
Consider a 4-dimensional Hilbert space and a density matrix
\begin{equation}
\rho=x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}+\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}\,.
\end{equation}
Let $P_{0}=\proj{0}+\proj{2}$, $P_{1}=\proj{1}+\proj{2}$, and $\bar{\mu}\left(P\right)=\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(P\right)\right)$.
Then,
\begin{equation}\eqalign{ 
 & \bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\\
= & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\\ 
= & \iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)
\nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)\\ 
= & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)
=\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\,. 
}\end{equation}
\end{proof}

In conclusion, we may still use an additive but non-convex interval-valued
function to construct QIVPMs, but we need to verify the convexity
of the constructed QIVPM one-by-one which might be long and tedious.

\subsection{Convex but Infinite Number of Intervals}

Since we cannot restrict our convex interval-valued function mapping
to only finite number of intervals, let's consider another family
of natural interval-valued functions.

\begin{lemma}Given a function\nb$c:\left[0,1\right]\rightarrow\left[0,1\right]$
satisfying the following conditions:
\begin{itemize}
\item For all $x\in\left[0,1\right]$, $c\left(x\right)\le x$.
\item For all $x\in\left[0,1\right]$, $c\left(1-x\right)=c\left(x\right)$.
\item $c$ is Wright-concave\nb\cite{Brzdek_2014}, i.e., for any three
numbers\nb$x_{0}$, $x_{1}$, and $x_{2}\in\left[0,1\right]$, $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$
implies $c\left(x_{0}+x_{1}+x_{2}\right)+c\left(x_{2}\right)\le c\left(x_{0}+x_{2}\right)+c\left(x_{1}+x_{2}\right)$.
\end{itemize}
Then, there is a convex interval-valued function\nb$\iota_{c}':\left[0,1\right]\rightarrow\mathscr{I}_{c}'$
defined by $\iota_{c}'\left(x\right)=\left[x-c\left(x\right),x+c\left(x\right)\right]$,
where $\mathscr{I}_{c}'=\set{\iota_{c}'\left(x\right)}{x\in\left[0,1\right]}$.\end{lemma}

\begin{proof}~
\begin{itemize}
\item Since $c\left(x\right)\le x$, $0\le x-c\left(x\right)$. Also, $c\left(x\right)=c\left(1-x\right)\le1-x$
implies $x+c\left(x\right)\le1$. Therefore, $\iota_{c}'\left(x\right)=\left[x-c\left(x\right),x+c\left(x\right)\right]\subseteq\left[0,1\right]$.
\item For any $x\in\left[0,1\right]$, $\iota_{c}'\left(1-x\right)=\left[1,1\right]-\left[x-c\left(1-x\right),x+c\left(1-x\right)\right]=\left[1,1\right]-\iota_{c}'\left(x\right)$.
\item Since $c\left(0\right)\in\left[0,1\right]$ and $c\left(0\right)\le0$,
we have $c\left(0\right)=0$. Therefore, $\iota_{c}'\left(0\right)=\left[0,0\right]$,
and $\iota_{c}'\left(1\right)=\left[1,1\right]-\iota_{c}'\left(0\right)=\left[1,1\right]$. 
\item Given any three numbers\nb$x_{0}$, $x_{1}$, and $x_{2}\in\left[0,1\right]$,
if $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, we have
\begin{equation}\eqalign{ 
 & c\left(x_{0}+x_{1}+x_{2}\right)+c\left(x_{2}\right)\le c\left(x_{0}+x_{2}\right)+c\left(x_{1}+x_{2}\right)\\
\Rightarrow & \left[x_{0}+x_{1}+x_{2}+c\left(x_{0}+x_{1}+x_{2}\right)\right]+\left[x_{2}+c\left(x_{2}\right)\right]\\
 & \le\left[x_{0}+x_{2}+c\left(x_{0}+x_{2}\right)\right]+\left[x_{1}+x_{2}+c\left(x_{1}+x_{2}\right)\right]\,.
}\end{equation}
Also, 
\begin{equation}\eqalign{ 
 & c\left(x_{0}+x_{1}+x_{2}\right)+c\left(x_{2}\right)\le c\left(x_{0}+x_{2}\right)+c\left(x_{1}+x_{2}\right)\\
\Rightarrow & -c\left(x_{0}+x_{2}\right)-c\left(x_{1}+x_{2}\right)\le-c\left(x_{0}+x_{1}+x_{2}\right)-c\left(x_{2}\right)\\
\Rightarrow & \left[x_{0}+x_{2}-c\left(x_{0}+x_{2}\right)\right]+\left[x_{1}+x_{2}-c\left(x_{1}+x_{2}\right)\right]\\
 & \le\left[x_{0}+x_{1}+x_{2}-c\left(x_{0}+x_{1}+x_{2}\right)\right]+\left[x_{2}-c\left(x_{2}\right)\right]\,.
}\end{equation}
The above two equations implies $\iota_{c}'\left(x_{0}+x_{1}+x_{2}\right)+\iota_{c}'\left(x_{2}\right)\subseteq\iota_{c}'\left(x_{0}+x_{2}\right)+\iota_{c}'\left(x_{1}+x_{2}\right)$.
\item For a set\nb$\left\{ x_{i}\right\} _{i=1}^{N}\subseteq\left[0,1\right]$,
if $\sum_{i=1}^{N}x_{i}\in\left[0,1\right]$, then $\iota_{c}'\left(\sum_{i=1}^{N}x_{i}\right)\subseteq\sum_{i=1}^{N}\iota_{c}'\left(x_{i}\right)$
by inducting on the convex condition listed previously with $x_{2}=0$.
\end{itemize}
\end{proof}

\subsection{Sates Consistent with a QIVPM Induced by the Born Rule}

When we work with mixed states or density matrices, we also need a
norm to compute the distance among them.

\begin{definition}\label{def:norm}Given a operator or matrix\nb$A$,
the norm of $A$ is defined by\nb\cite{544199}
\begin{equation}
\left\Vert A\right\Vert =\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{A}{\psi}\right|\,.
\end{equation}
\end{definition}

Notice that the norm in definition~\ref{def:norm} is not one of
the common matrix norms.

\begin{example}Consider $A=\left(\begin{array}{cc}
1 & 0\\
1 & 1
\end{array}\right)$. It is easy to verify $\left\Vert A\right\Vert =3/2$. On the other
hand, its Frobenius norm or Hilbert-Schmidt norm is\nb\cite{GolubVanLoan1996,Jaeger2007}
\begin{equation}
\left\Vert A\right\Vert _{\mathrm{F}}=\Tr\left(A^{\dagger}A\right)=\left(\sum_{i,j}\left|a_{ij}\right|^{2}\right)^{1/2}=\sqrt{3}\,,
\end{equation}
where $a_{ij}$ are the $\left(i,j\right)$-th entry of $A$. If we
consider the operator $p$-norms\nb\cite{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996}
\begin{equation}
\left\Vert A\right\Vert _{p}=\sup_{\left\Vert \ket{\psi}\right\Vert _{p}=1}\left\Vert A\ket{\psi}\right\Vert _{p}=\sup_{\ket{\psi}\ne0}\frac{\left\Vert A\ket{\psi}\right\Vert _{p}}{\left\Vert \ket{\psi}\right\Vert _{p}}\,,
\end{equation}
we have $\left\Vert A\right\Vert _{1}=\left\Vert A\right\Vert _{\infty}=2$
and $\left\Vert A\right\Vert _{2}=\left[\left(3+\sqrt{5}\right)/2\right]^{1/2}$,
where $\ket{\psi}=\sum_{i=1}^{d}\alpha_{i}\ket{i}$, $\left\Vert \ket{\psi}\right\Vert _{\infty}=\max_{i}\left|\alpha_{i}\right|$,
and $\left\Vert \ket{\psi}\right\Vert _{p}=\left(\sum_{i=1}^{d}\left|\alpha_{i}\right|^{p}\right)^{1/p}$
for $1\le p<\infty$. The common matrix norms of $A$ are all different
from $3/2$ so that our norm defined in definition~\ref{def:norm}
is different from all of them.\end{example}

\begin{lemma}\label{thm:Finite-precision-Gleason}Given a Hilbert
space with the set of all projectors\nb$\events$, and an additive
interval-valued function\nb$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
with the maximum length of intervals in $\mathscr{I}$ is $\delta$.
For two mixed states\nb$\rho$ and $\rho'$, if $\rho'\in\coreBorn\left(\iota\circ\mu_{\rho}^{\mathrm{B}},\events\right)$,
then $\left\Vert \rho-\rho'\right\Vert \le\delta$.\end{lemma}

\begin{proof}By definition, we have $\mu_{\rho'}^{\mathrm{B}}\left(P\right)\in\iota\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$
for all $P\in\events$. By substitute $P=\proj{\psi}$, we have 
\begin{equation}
\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)\in\iota\left(\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)\right)\,.\label{eq:consistent2another_state}
\end{equation}
Since the maximum length of intervals in $\mathscr{I}$ is $\delta$,
we can move $\rho$ and $\rho'$ to the same side and rewrite equation~(\ref{eq:consistent2another_state})
as
\begin{equation}
\delta\ge\left|\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)-\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)\right|=\left|\melem{\psi}{\rho'-\rho}{\psi}\right|\,.
\end{equation}
Applying maximum on both side, we have
\begin{equation}
\delta\ge\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{\rho'-\rho}{\psi}\right|=\left\Vert \rho'-\rho\right\Vert \,.
\end{equation}
\end{proof}

By applying the previous lemma on example~\ref{ex:equilateral},
we can estimate the states consistent with a QIVPM more tightly.

\begin{example}Given the additive interval-valued function\nb$\widehat{\iota}_{n}:\left[0,1\right]\rightarrow\widehat{\mathscr{I}}_{n}$,
the maximum length of intervals in $\widehat{\mathscr{I}}_{n}$ is
$1/n$. Therefore, for any totally mixed state\nb$\rho$ on a $d$-dimensional
Hilbert space, if $\rho'\in\coreBorn\left(\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)$,
then $\left\Vert \rho-\rho'\right\Vert \le1/n$.

For example, when $n=2$, two states~$\rho_{0}=\frac{1}{3}\cdot\proj{0}+\frac{1}{3}\cdot\proj{1}+\frac{1}{3}\cdot\proj{2}$
and $\rho_{1}=\frac{7}{24}\cdot\proj{0}+\frac{1}{3}\cdot\proj{1}+\frac{3}{8}\cdot\proj{2}$
can induce two QIVPMs~$\widehat{\iota}_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$
and $\widehat{\iota}_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}}$. Given
any projector $P$ in the minimal subspace of events containing $\proj{0}$,
$\proj{1}$, and $\proj{2}$, $\events_{C}$, the interval-valued
probability of the induced QIVPMs, $\widehat{\iota}_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)\right)$
and $\widehat{\iota}_{2}\left(\mu_{\rho_{1}}^{\mathrm{B}}\left(P\right)\right)$,
can be computed straightforwardly in Table~\ref{tab:Compute-the-QIVPM}.
\begin{table}
\caption{\label{tab:Compute-the-QIVPM}Compute $\widehat{\iota}_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$
and $\widehat{\iota}_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}}$ based on
$\mu_{\rho_{0}}^{\mathrm{B}}$ and $\mu_{\rho_{1}}^{\mathrm{B}}$
on $\events_{C}$.}
\centering{}%
\begin{tabular}{ccccc}
\toprule 
\addlinespace
$P$ & $\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)$ & $\widehat{\iota}_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)\right)$ & $\mu_{\rho_{1}}^{\mathrm{B}}\left(P\right)$ & $\widehat{\iota}_{2}\left(\mu_{\rho_{1}}^{\mathrm{B}}\left(P\right)\right)$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
$\mathbb{0}$ & $0$ & $\imposs$ & $0$ & $\imposs$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\proj{0}$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$ & $\frac{7}{24}$ & $\left[0,\frac{1}{2}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\proj{1}$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\proj{2}$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$ & $\frac{3}{8}$ & $\left[\tfrac{1}{4},\tfrac{3}{4}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\mathbb{1}-\proj{2}$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$ & $\frac{5}{8}$ & $\left[\tfrac{1}{4},\tfrac{3}{4}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\mathbb{1}-\proj{1}$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\mathbb{1}-\proj{0}$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$ & $\frac{17}{24}$ & $\left[\frac{1}{2},1\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\mathbb{1}$ & $1$ & $\necess$ & $1$ & $\necess$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
Table~\ref{tab:Compute-the-QIVPM} shows that $\rho_{0}$ is consistent
with $\widehat{\iota}_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}}$ and $\rho_{1}$
is consistent with $\widehat{\iota}_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$
both on every projectors in $\events_{C}$. Indeed, they are consistent
with each other on every projectors in $\events$ that implies $\left\Vert \rho_{0}-\rho_{1}\right\Vert \le\frac{1}{2}$
by lemma~\ref{thm:Finite-precision-Gleason}. This is true because
the exact value of $\left\Vert \rho_{0}-\rho_{1}\right\Vert $ is
$\frac{1}{24}$~\cite{544199}.\end{example}

Previously, we proved that if $\rho'$ is consistent with $\iota\circ\mu_{\rho}^{\mathrm{B}}$,
then $\rho'$ is close to $\rho$. However, if $\rho'$ is close to
$\rho$, there is no way to guarantee $\rho'$ is consistent with
$\iota\circ\mu_{\rho}^{\mathrm{B}}$ because no matter how close between
two real numbers~$x$ and $y$, we can never guarantee $y\in\iota\left(x\right)$
as in the following example.

\begin{example}Given an additive interval-valued function\nb$\iota:\left[0,1\right]\rightarrow\mathscr{I}$,
consider let $x_{0}=\sup\set{x\in\left[0,1\right]}{\iota\left(x\right)=\left[0,0\right]}$.
Since $\iota\left(1\right)=\left[1,1\right]$, $x_{0}<1$. For any
$\varepsilon>0$, $\left|\left(x+\frac{\varepsilon}{2}\right)-x\right|=\frac{\varepsilon}{2}<\varepsilon$,
and $x+\frac{\varepsilon}{2}\ge\frac{\varepsilon}{2}>0$ implies $x+\frac{\varepsilon}{2}\notin\left[0,0\right]=\iota\left(x\right)$.

Notice that even if we explicitly exclude $0$ and $1$, the previous
argument applies on any boundary points among intervals. Therefore,
there is no way to avoid this problem if we have more than three intervals.\end{example}

\subsection{Category Theory}

It looks like $\iota$ might be a morphism in a category. Let's try
to check if it is true here.

\begin{definition}Given $\mathscr{I}$ is a collection of intervals,
we say a interval-valued function\nb$\iota:\mathscr{I}'\rightarrow\mathscr{I}''$
is additive if $\iota$ satisfies the following properties. 
\begin{itemize}
\item $\iota\left(\imposs\right)=\imposs$. 
\item $\iota\left(\necess\right)=\necess$. 
\item For any $\left[\ell,r\right]\in\mathscr{I}'$, $\iota\left(\necess-\left[\ell,r\right]\right)=\necess-\iota\left(\left[\ell,r\right]\right)$. 
\item For a pair\nb$\left[\ell_{0},r_{0}\right]$ and $\left[\ell_{1},r_{1}\right]\in\mathscr{I}'$,
if $\left[\ell_{0},r_{0}\right]\subseteq\left[\ell_{1},r_{1}\right]$,
then $\iota\left(\left[\ell_{0},r_{0}\right]\right)\subseteq\iota\left(\left[\ell_{1},r_{1}\right]\right)$. 
\item For a pair\nb$\left[\ell_{0},r_{0}\right]$ and $\left[\ell_{1},r_{1}\right]\in\mathscr{I}'$,
if $\left[\ell_{0},r_{0}\right]+\left[\ell_{1},r_{1}\right]\cap\left[0,1\right]\neq\emptyset$,
then $\left[\ell_{0},r_{0}\right]+\left[\ell_{1},r_{1}\right]\in\mathscr{I}'$
and $\iota\left(\left[\ell_{0},r_{0}\right]+\left[\ell_{1},r_{1}\right]\right)\subseteq\iota\left(\left[\ell_{0},r_{0}\right]\right)+\iota\left(\left[\ell_{1},r_{1}\right]\right)$. 
\end{itemize}
Moreover, $\iota$ is called convex if for any three intervals\nb$\left[\ell_{0},r_{0}\right]$,
$\left[\ell_{1},r_{1}\right]$, and $\left[\ell_{2},r_{2}\right]\in\mathscr{I}'$,
their sum $\left[\ell_{0},r_{0}\right]+\left[\ell_{1},r_{1}\right]+\left[\ell_{2},r_{2}\right]\in\mathscr{I}'$
implies $\iota\left(\left[\ell_{0},r_{0}\right]+\left[\ell_{1},r_{1}\right]+\left[\ell_{2},r_{2}\right]\right)+\iota\left(\left[\ell_{2},r_{2}\right]\right)\subseteq\iota\left(\left[\ell_{0},r_{0}\right]+\left[\ell_{2},r_{2}\right]\right)+\iota\left(\left[\ell_{1},r_{1}\right]+\left[\ell_{2},r_{2}\right]\right)$.\end{definition}

\begin{lemma}\label{lemma:interval2interval}Consider an additive
interval-valued function\nb$\iota:\mathscr{I}'\rightarrow\mathscr{I}''$.
Then, given a QIVPM\nb$\bar{\mu}':\events\rightarrow\mathscr{I}'$,
a function\nb$\bar{\mu}'':\events\rightarrow\mathscr{I}''$ defined
by $\bar{\mu}''\left(P\right)=\iota\left(\bar{\mu}'\left(P\right)\right)$
is a QIVPM.\end{lemma}

\begin{proof}~
\begin{itemize}
\item $\bar{\mu}''\left(\mathbb{0}\right)=\iota\left(\bar{\mu}'\left(\mathbb{0}\right)\right)=\iota\left(\imposs\right)=\imposs$.
\item $\bar{\mu}''\left(\mathbb{1}\right)=\iota\left(\bar{\mu}'\left(\mathbb{1}\right)\right)=\iota\left(\necess\right)=\necess$.
\item For any projection $P$, $\bar{\mu}''\left(\mathbb{1}-P\right)=\iota\left(\necess-\bar{\mu}'\left(P\right)\right)=\necess-\bar{\mu}''\left(P\right)$.
\item For a pair of mutually orthogonal projections\nb$P_{0}$ and $P_{1}\in\events$,
since $\emptyset\neq\bar{\mu}'\left(P_{0}+P_{1}\right)\cap\left[0,1\right]\subseteq\bar{\mu}'\left(P_{0}\right)+\bar{\mu}'\left(P_{1}\right)\cap\left[0,1\right]$,
we have $\bar{\mu}'\left(P_{0}\right)+\bar{\mu}'\left(P_{1}\right)\in\mathscr{I}'$
and
\[
\bar{\mu}''\left(P_{0}+P_{1}\right)=\iota\left(\bar{\mu}'\left(P_{0}+P_{1}\right)\right)\subseteq\iota\left(\bar{\mu}'\left(P_{0}\right)+\bar{\mu}'\left(P_{1}\right)\right)\subseteq\iota\left(\bar{\mu}'\left(P_{0}\right)\right)+\iota\left(\bar{\mu}'\left(P_{1}\right)\right)=\bar{\mu}''\left(P_{0}\right)+\bar{\mu}''\left(P_{1}\right)\,.
\]
\end{itemize}
\end{proof}

\section{From Interval-valued to Real-valued}

Previously we discussed how to construct a quantum IVPM from a quantum
real-valued probability measure; we then switch our focus to how to
construct a quantum real-valued probability measure from a special
kind of quantum IVPMs. Since the only properties we use in this section
is whether its interval-valued or real-valued, the results in this
section applies on every subspace of events\nb$\events'$. In particular,
when $\events'$ is commuting, any quantum results will be classical
as shown in theorem~\ref{thm:sub-event-space}.

\begin{definition}\label{def:delta-deterministic}An interval-valued
probability measure\nb$\bar{\mu}:\events'\rightarrow\mathscr{I}$
is called $\left(1/n,\delta\right)$-deterministic if for every $\left[l,r\right]\in\mathscr{I}$,
there is a $k\in\mathbb{Z}$ such that $\left[l,r\right]\subseteq\left[k/n-\delta,k/n+\delta\right]$.\end{definition}

Recall $\mathscr{I}_{\mathrm{D}}=\left\{ \imposs,\necess\right\} $
so that a $\mathscr{I}_{\mathrm{D}}$-valued QIVPM is $\left(1,0\right)$-deterministic.
For another example, given a three-dimensional Hilbert space, 
\begin{equation}
\iota_{\infty}\circ\mu_{\frac{\mathbb{1}}{3}}^{\mathrm{B}}:\events'\rightarrow\left\{ \imposs,\left[\frac{1}{3},\frac{1}{3}\right],\left[\frac{2}{3},\frac{2}{3}\right],\necess\right\} 
\end{equation}
is $\left(1,\frac{1}{3}\right)$-deterministic and $\left(\frac{1}{3},0\right)$-deterministic.
However, we will prove there is generally no $\left(1/n,\delta\right)$-deterministic
probability measures when $\delta$ is much smaller than $1/n$.

\begin{lemma}\label{lemma:QIVPM2QCM}Given a Hilbert space of dimension\nb$d$,
and a quantum $\left(1/n,\delta\right)$-deterministic probability
measure\nb$\bar{\mu}:\events'\rightarrow\mathscr{I}$ with $\delta<1/\left(3n\right)$,
if $\bar{\mu}(P)\subseteq\left[k/n-\delta,k/n+\delta\right]$, let
$\mu(P)=k/n$, then $\mu:\events'\rightarrow\left[0,1\right]$ is
a quantum probability measure. In another word, $\iota_{\infty}\circ\mu$
is $\left(1/n,0\right)$-deterministic by lemma~\ref{lemma:real2interval}.\end{lemma}

\begin{proof}~
\begin{itemize}
\item Since $\bar{\mu}(\mathbb{0})=\imposs$, we have $\mu(\mathbb{0})=0$.
\item Since $\bar{\mu}(\mathbb{1})=\necess$, we have $\mu(\mathbb{1})=1$.
\item For every event $P$, we have 
\begin{equation}\eqalign{ 
& \pmeas\left(P\right)=\frac{k}{n}\Rightarrow\bar{\mu}(P)\subseteq\left[\frac{k}{n}-\delta,\frac{k}{n}+\delta\right]\\
\Rightarrow & \bar{\mu}(\mathbb{1}-P)=\left[1,1\right]-\bar{\mu}(P)\subseteq\left[1-\frac{k}{n}-\delta,1-\frac{k}{n}+\delta\right]\\
\Rightarrow & \pmeas\left(\mathbb{1}-P\right)=1-\frac{k}{n}=1-\pmeas\left(P\right)\,.
}\end{equation}
\item Given a pair of orthogonal operators\nb$P_{0}$ and $P_{1}$, let
$\pmeas\left(P_{0}\right)=k_{0}/n$ and $\pmeas\left(P_{1}\right)=k_{1}/n$.
Then, 
\begin{equation}
\bar{\mu}\left(P_{0}+P_{1}\right)\subseteq\bar{\mu}(P_{0})+\bar{\mu}(P_{1})\subseteq\left[\frac{k_{0}+k_{1}}{n}-2\delta,\frac{k_{0}+k_{1}}{n}+2\delta\right]\,.
\end{equation}
Since $\delta<1/\left(3n\right)$, we have
\begin{equation}
\frac{k_{0}+k_{1}-1}{n}+\delta<\frac{k_{0}+k_{1}}{n}-2\delta\le\frac{k_{0}+k_{1}}{n}+2\delta<\frac{k_{0}+k_{1}+1}{n}-\delta\,.
\end{equation}
The previous inequality implies that if $\bar{\mu}\left(P_{0}+P_{1}\right)$
overlap with any interval of the form $\left[k/n-\delta,k/n+\delta\right]$
for some $k$, $k$ must be $k_{0}+k_{1}$, i.e., 
\begin{equation}
\bar{\mu}\left(P_{0}+P_{1}\right)\subseteq\left[\frac{k_{0}+k_{1}}{n}-\delta,\frac{k_{0}+k_{1}}{n}+\delta\right]\,.
\end{equation}
Therefore, 
\begin{equation}
\pmeas\left(P_{0}+P_{1}\right)=\frac{k_{0}+k_{1}}{n}=\pmeas\left(P_{0}\right)+\pmeas\left(P_{1}\right)\,.
\end{equation}
By induction, we have $\pmeas\left(\bigcup_{i=1}^{N}P_{i}\right)=\sum_{i=1}^{N}\pmeas(P_{i})$
for any mutually orthogonal events\nb$\left\{ P_{i}\right\} _{i=1}^{N}$. 
\end{itemize}
\end{proof}

Together with the Kochen-Specker theorem, we will have the following
corollary.

\begin{cor}\label{cor:Kochen-Specker-IVPM}Given a Hilbert space
of dimension\nb$d\ge3$, there is no quantum $\left(1,\delta\right)$-deterministic
probability measure for $\delta<\frac{1}{3}$.\end{cor}

In general~\cite{Redhead1987-REDINA}, Gleason's theorem asserts
that a quantum probability measure must be continuous. Therefore,
a $\left(1/n,0\right)$-deterministic QIVPM exists if and only if
$d\mid n$, and so does a $\left(1/n,\delta\right)$-deterministic
QIVPM for $\delta<1/\left(3n\right)$.

Lemma~\ref{lemma:QIVPM2QCM} may be extended in another trivial way.

\begin{lemma}If every interval in $\mathscr{I}$ is a singleton set,
given a QIVPM\nb$\bar{\mu}:\events'\rightarrow\mathscr{I}$, then
$\mu:\events'\rightarrow\left[0,1\right]$ defined by $\bar{\mu}\left(P\right)=\left\{ \mu\left(P\right)\right\} $
is a quantum probability measure.\end{lemma}

\begin{proof}~
\begin{itemize}
\item Since $\bar{\mu}(\mathbb{0})=\imposs$, we have $\mu(\mathbb{0})=0$.
\item Since $\bar{\mu}(\mathbb{1})=\necess$, we have $\mu(\mathbb{1})=1$.
\item For every event $P$, we have
\begin{eqnarray*}
 &  & \bar{\mu}\left(\mathbb{1}-P\right)=\left[1,1\right]-\bar{\mu}\left(P\right)\\
 & \Rightarrow & \left\{ \mu\left(\mathbb{1}-P\right)\right\} =\left[1,1\right]-\left\{ \mu\left(P\right)\right\} =\left\{ 1-\mu\left(P\right)\right\} \\
 & \Rightarrow & \mu\left(\mathbb{1}-P\right)=1-\mu\left(P\right)\,.
\end{eqnarray*}
\item Given a pair of orthogonal operators\nb$P_{0}$ and $P_{1}$, we
have
\begin{eqnarray*}
 &  & \bar{\mu}\left(P_{0}+P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\\
 & \Rightarrow & \left\{ \mu\left(P_{0}+P_{1}\right)\right\} \subseteq\left\{ \mu\left(P_{0}\right)\right\} +\left\{ \mu\left(P_{1}\right)\right\} =\left\{ \mu\left(P_{0}\right)+\mu\left(P_{1}\right)\right\} \\
 & \Rightarrow & \mu\left(P_{0}+P_{1}\right)=\mu\left(P_{0}\right)+\mu\left(P_{1}\right)\,.
\end{eqnarray*}
\end{itemize}
\end{proof}

Together with example~\ref{ex:regular-pair}, we know QIVPMs with
$\mathscr{I}_{\infty}$ is one-to-one correspondence to quantum real-valued
probability measures. Moreover, since $\iota_{\infty}:\left[0,1\right]\rightarrow\mathscr{I}_{\infty}$
is convex, we have the following corollary.

\begin{cor}If every interval in $\mathscr{I}$ is a singleton set,
then every QIVPM\nb$\bar{\mu}:\events'\rightarrow\mathscr{I}$ is
convex.\end{cor}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Examples for Quantum Interval-valued Probability Measures\label{sec:Examples-for-QuantumIVPM}}

\begin{figure}
\begin{centering}
\includegraphics[scale=0.38]{figureE1} 
\par\end{centering}
\caption{\label{fig:three-dimensional-5-value}This figure illustrates example~\ref{ex:three-dimensional-5-value}
plotted in $\mathbb{R}^{3}$. The red and green dotted vectors are
$\ket{0}$ and $\ket{\ps}$ respectively. All possible real vectors
of the subspaces orthogonal to $\ket{0}$ or $\ket{\ps}$ are drawn
in the red and green circles, respectively.}
\end{figure}

\begin{example}[Three-dimensional quantum 5-interval-valued probability
measure]\label{ex:three-dimensional-5-value} Given a three dimensional
Hilbert space with an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Let\emph{ }$\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$, and
$\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess,\midd\right\} $,
where $\midd$ represents $\left[\frac{1}{4},\frac{3}{4}\right]$.
The definition of $\bar{\mu}$ below refers to figure~\ref{fig:three-dimensional-5-value}
which plots the 1-dimensional projectors: 
\begin{enumerate}
\item Let: 
\begin{equation}\eqalign{ 
\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs\,,\\
\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps})=\necess\,, 
}\end{equation}
where $\ket{0}$ and $\ket{\ps}$ are plotted as the red and green
dotted vectors, respectively. 
\item The red and green circles are the states orthogonal to $\ket{0}$
and $\ket{\ps}$, respectively. We define $\bar{\mu}\left(\proj{\psi}\right)=\bar{\mu}\left(\mathbb{1}-\proj{\psi}\right)=\midd$,
where $\ip{\psi}{0}=0$ or $\ip{\psi}{\ps}=0$.
\item Otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$ and $\bar{\mu}(\mathbb{1}-\proj{\psi})=\likely$. 
\end{enumerate}
By theorem~\ref{thm:interval-valued-frame-function}, to check $\bar{\mu}$
a quantum interval-valued probability measure is equivalent to check
$\bar{f}$ satisfying 
\begin{equation}\eqalign{ 
\left[1,1\right]-\bar{f}\left(\ket{\psi_0}\right)\subseteq\bar{f}\left(\ket{\psi_1}\right)+\bar{f}\left(\ket{\psi_2}\right)\,,\\
\left[1,1\right]-\bar{f}\left(\ket{\psi_1}\right)\subseteq\bar{f}\left(\ket{\psi_2}\right)+\bar{f}\left(\ket{\psi_0}\right)\,,\\
\left[1,1\right]-\bar{f}\left(\ket{\psi_2}\right)\subseteq\bar{f}\left(\ket{\psi_0}\right)+\bar{f}\left(\ket{\psi_1}\right)
}\label{eq:non-additive-vectors}\end{equation}
for every orthonormal basis~$\left\{ \ket{\psi_{0}},\ket{\psi_{1}},\ket{\psi_{2}}\right\} $,
where $\bar{f}\left(\ket{\psi}\right)=\bar{\mu}\left(\proj{\psi}\right)$.
We are going to enumerate all possible orthonormal bases to verify
equation~(\ref{eq:non-additive-vectors}). Also, to simplify the
notation, we denote $\set{\ket{\psi}}{\ip{\psi}{0}\ne0\textrm{ and }\ip{\psi}{\ps}\ne0}$
by $\mathcal{T}$.
\begin{enumerate}
\item When $\ket{\psi_{0}}$ is $\ket{0}$, then $\ip{\psi_{1}}{0}=\ip{\psi_{2}}{0}=0$.
Equation~(\ref{eq:non-additive-vectors}) can then be verified as
follow. 
\begin{equation}\eqalign{ 
\left[1,1\right]-\bar{f}\left(\ket{0}\right)=\necess\subseteq\midd+\midd=\bar{f}\left(\ket{\psi_1}\right)+\bar{f}\left(\ket{\psi_2}\right)\,,\\
\left[1,1\right]-\bar{f}\left(\ket{\psi_1}\right)=\midd=\midd+\imposs=\bar{f}\left(\ket{\psi_2}\right)+\bar{f}\left(\ket{0}\right)\,,
}\end{equation}
Similarly, when $\ket{\psi_{0}}$ is $\ket{\ps}$, equation~(\ref{eq:non-additive-vectors})
holds. 
\item When $\ip{\psi_{0}}{0}=\ip{\psi_{1}}{\ps}=0$ and $\ket{\psi_{2}}\in\mathcal{T}$,
we have 
\begin{equation}\eqalign{ 
\left[1,1\right]-\bar{f}\left(\ket{\psi_0}\right)=\midd\subseteq\midd+\unlikely=\bar{f}\left(\ket{\psi_1}\right)+\bar{f}\left(\ket{\psi_2}\right)\,,\\
\left[1,1\right]-\bar{f}\left(\ket{\psi_2}\right)=\likely\subseteq\midd+\midd=\bar{f}\left(\ket{\psi_0}\right)+\bar{f}\left(\ket{\psi_1}\right)\,.
}\end{equation}
\item When $\ip{\psi_{0}}{0}=0$, and $\ket{\psi_{1}}$, $\ket{\psi_{2}}\in\mathcal{T}$,
we have 
\begin{equation}\eqalign{ 
\left[1,1\right]-\bar{f}\left(\ket{\psi_0}\right)=\midd\subseteq\unlikely+\unlikely=\bar{f}\left(\ket{\psi_1}\right)+\bar{f}\left(\ket{\psi_2}\right)\,,\\
\left[1,1\right]-\bar{f}\left(\ket{\psi_1}\right)=\likely\subseteq\unlikely+\midd=\bar{f}\left(\ket{\psi_2}\right)+\bar{f}\left(\ket{\psi_0}\right)\,.
}\end{equation}
Similarly, when $\ip{\psi_{0}}{\ps}=0$, $\ket{\psi_{1}}\in\mathcal{T}$,
and $\ket{\psi_{2}}\in\mathcal{T}$, equation~(\ref{eq:non-additive-vectors})
holds. 
\item When $\ket{\psi_{0}}$, $\ket{\psi_{1}}$, and $\ket{\psi_{2}}\in\mathcal{T}$,
i.e., the ``otherwise'' case. Then, equation~(\ref{eq:non-additive-vectors})
can easily be verified. 
\begin{equation}
\left[1,1\right]-\bar{f}\left(\ket{\psi_{0}}\right)=\likely\subseteq\unlikely+\unlikely=\bar{f}\left(\ket{\psi_{1}}\right)+\bar{f}\left(\ket{\psi_{2}}\right)\,.
\end{equation}
\end{enumerate}
Since we proved that $\bar{\mu}$ is a quantum interval-valued probability
measure, $\bar{\mu}$ is convex by theorem~\ref{thm:convex-3}.

The following argument establishes that $\bar{\mu}$ has an empty
core. Assume there is a real-valued probability measure satisfying
$\mu_{\rho}^{\mathrm{\mathrm{B}}}(P)\in\bar{\mu}(P)$ for all $P\in\events$.
Because $\mu_{\rho}^{\mathrm{\mathrm{B}}}(\proj{0})\in\bar{\mu}(\proj{0})=\imposs$
and $\mu_{\rho}^{\mathrm{\mathrm{B}}}(\proj{\ps})\in\bar{\mu}(\proj{\ps})=\imposs$,
we must have $\mu_{\rho}^{\mathrm{\mathrm{B}}}(\proj{0})=\mu_{\rho}^{\mathrm{\mathrm{B}}}(\proj{\ps})=0$
so that $\mu_{\rho}^{\mathrm{\mathrm{B}}}=\mu_{\ket{2}}^{\mathrm{\mathrm{B}}}$.
However, 
\begin{equation}
\mu_{\ket{2}}^{\mathrm{\mathrm{B}}}(\proj{2})=1\notin\unlikely=\bar{\mu}(\proj{2})\,.
\end{equation}
\end{example}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Expectation Values for IVPMs}

\subsection{Expectation Values for Classical IVPMs}

We may need to understand how to define expectation values of an observable
for quantum IVPMs. Recall an observable is the quantum version of
a random variable. Hence, in order to define expectation values of
an observable, we need to understand the definition of expectation
values of a random variable for an classical IVPM. Since the expectation
value of random variable~$X$ is the integral $\int X\rmd\pmeas$~\cite{Kolmogorov1950},
we need to understand the meaning of integration on IVPMs. In the
rest of the section, we will consider the real-valued and interval-valued
probability measure defined in example~\ref{ex1} and \ref{ex3},
and see how to compute the expectation value of different random variables.
To simplify the discussion later, we will only consider a random variable~$X:\Omega\rightarrow\mathbb{R}$
which is a step function, i.e., $X=\sum_{i=1}^{N}x_{i}\mathbf{1}_{E_{i}}$,
where $E_{i}\in\events$ and $\mathbf{1}_{E}$ is the indicator function
defined by 
\begin{equation} 
\mathbf{1}_{A}(x)=\cases{ 
1 & if $x\in A$ ;\\ 
0 & if $x\notin A$. 
}
\end{equation}

Given a random variable~$X$ and a probability measure~$\pmeas$,
the expectation value of $X$ is 
\begin{eqnarray}
\int X\rmd\pmeas & = & \sum_{i=1}^{N}x_{i}\pmeas\left(E_{i}\right)\,.\label{eq:expectation}
\end{eqnarray}
Intuitively, the expectation value is weighted average of the possible
outcomes of $X$ with the weight equal to the probability of each
outcome could happen. For example, consider the random variable~$X_{0}$
representing the number of tails. For the probability measure~$\pmeas$
in example~\ref{ex1}, the expectation value of $X_{0}$ is
\begin{equation}\eqalign{ 
\int X_{0}\rmd\pmeas & = 0\cdot\pmeas(\{HH\})+1\cdot\pmeas(\{HT,TH\})+2\cdot\pmeas(\{TT\})\\
 & = 0\cdot\frac{1}{3}+1\cdot\frac{2}{3}+2\cdot0=\frac{2}{3}\,. 
}\end{equation}

There is a subtlety happens when we want to define the expectation
value of a random variable on IVPMs, and will be demonstrated by the
following example.

\begin{example}\label{ex:expectation}Consider an CIVPM\nb$\bar{\mu}$
and an event\nb$E$ such that $\bar{\mu}(E)=[0.2,0.3]$ and $\bar{\mu}(\overline{E})=[0.7,0.8]$.
Given a random variable
\begin{equation} 
X\left(\omega\right)=\cases{ 
1 & , if $\omega\in E$ ;\\
2 & , if $\omega\notin E$ ,
}
\end{equation}
its expectation value with respect to $\bar{\mu}$ might be naïvely
defined as
\begin{equation}\eqalign{ 
 & 1\cdot\bar{\mu}(E)+2\cdot\bar{\mu}(\overline{E})\\
= & 1\cdot[0.2,0.3]+2\cdot[0.7,0.8]\\
= & [1\cdot0.2+2\cdot0.7,1\cdot0.3+2\cdot0.8]\\
= & [1.6,1.9]\,.
}\end{equation}
As we discussed when introducing the core, the true probability of
$\bar{\mu}(E)$, $\mu(E)$, can be anywhere in the range $[0.2,0.3]$
and the true probability of $\bar{\mu}(\overline{E})$ is $\mu(\overline{E})=1-\mu(E)$,
that is, $\mu\in\mathrm{core}\left(\bar{\mu}\right)$. If $\mu(E)=0.2$,
the expectation value of $X$ with respect to $\mu$ is 
\begin{equation}
1\cdot\mu(E)+2\cdot\mu(\overline{E})=1\cdot0.2+2\cdot0.8=1.8\,;
\end{equation}
if $\mu(E)=0.3$, the expectation value of $X$ with respect to $\mu$
is
\begin{equation}
1\cdot\mu(E)+2\cdot\mu(\overline{E})=1\cdot0.3+2\cdot0.7=1.7\,.
\end{equation}
In another word, no matter how we pick $\mu(E)$, the expectation
value of $X$ is in $\left[1.7,1.8\right]$ which is a more realistic
expectation value of $X$ with respect to $\bar{\mu}$. The interval~$\left[1.7,1.8\right]$
is narrower than the naïve expectation value~$[1.6,1.9]$ because
the naïve one combines contradicting situations in the computation.\end{example}

In general, the computation of the realistic expectation value, called
the Choquet integral, is more complex than what we discussed because
a random variable may have more than two values. Despite its complexity,
it is still a weight average among the values of a random variable
as given in the following definition.

\begin{definition}[Choquet integral~\cite{Vitali1925,Choquet1954,GilboaSchmeidler1994,Grabisch2016}]\label{def:Choquet}~
\begin{itemize}
\item Consider a classical IVPM~$\bar{\mu}$ such that $\mul:\events\rightarrow\left[0,1\right]$
and $\mur:\events\rightarrow\left[0,1\right]$ are the left-end and
the right-end of $\bar{\mu}$, respectively, i.e., $\bar{\mu}\left(E\right)=\left[\mul\left(E\right),\mur\left(E\right)\right]$.
\item We still decompose a random variable~$X$ into step functions, but
we separate the positive and negative terms and write like $\sum_{i=1}^{N^{-}}x_{i}^{-}\mathbf{1}_{E_{i}^{-}}+\sum_{i=1}^{N^{+}}x_{i}^{+}\mathbf{1}_{E_{i}^{+}}$,
i.e., 
\begin{equation}
x_{N^{-}}^{-}\le\ldots\le x_{1}^{-}\le0\le x_{1}^{+}\le\ldots\le x_{N^{+}}^{+}
\end{equation}
 and $\left\{ E_{i}^{-}\right\} _{i=1}^{N^{-}}$ and $\left\{ E_{i}^{+}\right\} _{i=1}^{N^{+}}$
are all disjoint.
\item Let $E_{i}^{s\prime}=\bigcup_{j\ge i}E_{j}^{s}$ and $\Delta\mu^{*}\left(E_{i}^{s\prime}\right)=\mu^{*}\left(E_{i}^{s\prime}\right)-\mu^{*}\left(E_{i+1}^{s\prime}\right)$,
where $s\in\left\{ +,-\right\} $ and $\mu^{*}\in\left\{ \mul,\mur\right\} $.
\end{itemize}
Then, the Choquet integral of $X$ with respect to $\bar{\mu}$ is
\begin{equation}\eqalign{ 
\int X\rmd\bar{\mu} = & \left[\sum_{i=1}^{N^{+}}x_{i}^{+}\Delta\mul\left(E_{i}^{+\prime}\right)+\sum_{i=1}^{N^{-}}x_{i}^{-}\Delta\mur\left(E_{i}^{-\prime}\right),\right.\\  
& \left.\sum_{i=1}^{N^{+}}x_{i}^{+}\Delta\mur\left(E_{i}^{+\prime}\right)+\sum_{i=1}^{N^{-}}x_{i}^{-}\Delta\mul\left(E_{i}^{-\prime}\right)\right]\,.
}\label{eq:interval-expectation}\end{equation}
\end{definition}

\begin{table}
\caption{\label{table:Choquet}
\[
\int X\rmd\bar{\mu}=\left[1\cdot0.3+2\cdot0.7,1\cdot0.2+2\cdot0.8\right]=\left[1.7,1.8\right]\,.
\]
}

\begin{indented}
\item\begin{tabular}{|c|c|c|c|c|c|c|c|} 
\br 
$i$ & $x_{i}^{+}$ & $E_{i}^{+}$ & $E_{i}^{+\prime}$ & $\mul\left(E_{i}^{+\prime}\right)$ & $\Delta\mul\left(E_{i}^{+\prime}\right)$ & $\mur\left(E_{i}^{+\prime}\right)$ & $\Delta\mur\left(E_{i}^{+\prime}\right)$\tabularnewline 
\mr 
$1$ & $1$ & $E$ & $\Omega$ & $1$ & $0.3$ & $1$ & $0.2$\tabularnewline 
\mr 
$2$ & $2$ & $\overline{E}$ & $\overline{E}$ & $0.7$ & $0.7$ & $0.8$ & $0.8$\tabularnewline 
\mr 
$3$ &  &  & $\emptyset$ & $0$ &  & $0$ & \tabularnewline 
\br 
\end{tabular}
\end{indented}
\end{table}
Table~\ref{table:Choquet} follows the definition to compute $\int X\rmd\bar{\mu}$.
Notice that equation~(\ref{eq:interval-expectation}) looks complex.
On the one hand, it would be a little bit easier to think its right-hand
side as 
\begin{equation}
\sum_{s\in\left\{ +,-\right\} }\sum_{i=1}^{N^{s}}x_{i}^{s}\left[\Delta\mul\left(E_{i}^{s\prime}\right),\Delta\mur\left(E_{i}^{s\prime}\right)\right]\,,\label{eq:interval-expectation-1}
\end{equation}
$\Delta\mul\left(E_{i}^{s\prime}\right)$ can be larger than $\Delta\mur\left(E_{i}^{s\prime}\right)$
as in the second row in table~\ref{table:Choquet}. Therefore, if
we want to use this notation, we need to stick to this convention
that not reverse the two ends of intervals, even if the left-end is
larger than the right-end. On the other hand, similar to equation~(\ref{eq:expectation}),
equation~(\ref{eq:interval-expectation-1}) can be considered as
weighted average of the possible outcomes of $X$ as well, despite
the ``weight''
\begin{equation}
\left[\Delta\mul\left(E_{i}^{s\prime}\right),\Delta\mur\left(E_{i}^{s\prime}\right)\right]\subseteq\left[\mul\left(E_{i}^{s}\right),\mur\left(E_{i}^{s}\right)\right]=\bar{\mu}\left(E_{i}^{s}\right)\textrm{ for }s\in\left\{ +,-\right\} 
\end{equation}
is tighter than the interval-valued probability of the corresponding
event. This makes it easier to remember.

Also, $\left[1.7,1.8\right]$ in example~\ref{ex:expectation} is
exactly the same as the expectation value computed in table~\ref{table:Choquet}.
This nature property can be generated to the following theorem.

\begin{thm}\label{thm:Rosenmuller}\cite{Rosenmuller1971,GilboaSchmeidler1994,Grabisch2016}For
every classical IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$ and
any random variable~$X:\Omega\rightarrow\mathbb{R}$, $\bar{\mu}$
is convex if and only if
\begin{equation}
\int X\rmd\bar{\mu}=\left[\min_{\mu\in\mathrm{core}\left(\bar{\mu}\right)}\int X\rmd\mu,\max_{\mu\in\mathrm{core}\left(\bar{\mu}\right)}\int X\rmd\mu\right]\,.\label{eq:Rosenmuller}
\end{equation}
\end{thm}

\yutsung{{Example for when }$\bar{\mu}$ is not convex, Eq.~(\ref{eq:Rosenmuller})
doesn't hold. This can justify why we need convex...}

The next lemmas shows our definition is exactly the same as in the
literature, and is consistent with the CRVPM.

\begin{lemma}Follow the same notation in the definition, we also
define $X^{+}=\sum_{i=1}^{N^{+}}x_{i}^{+}\mathbf{1}_{E_{i}^{+}}$
and $X^{-}=\sum_{i=1}^{N^{-}}x_{i}^{-}\mathbf{1}_{E_{i}^{-}}$. Then,
$\int X\rmd\bar{\mu}=\int X^{+}\rmd\bar{\mu}-\int\left(-X^{-}\right)\rmd\bar{\mu}$.\end{lemma}

\begin{proof}
\begin{eqnarray*}
\int X\rmd\bar{\mu} & = & \sum_{i=1}^{N^{+}}x_{i}^{+}\left[\Delta\mul\left(E_{i}^{+\prime}\right),\Delta\mur\left(E_{i}^{+\prime}\right)\right]+\sum_{i=1}^{N^{-}}x_{i}^{-}\left[\Delta\mul\left(E_{i}^{-\prime}\right),\Delta\mur\left(E_{i}^{-\prime}\right)\right]\\
 & = & \int X^{+}\rmd\bar{\mu}+\left[\sum_{i=1}^{N^{-}}x_{i}^{-}\Delta\mul\left(E_{i}^{-\prime}\right),\sum_{i=1}^{N^{-}}x_{i}^{-}\Delta\mur\left(E_{i}^{-\prime}\right)\right]\\
 & = & \int X^{+}\rmd\bar{\mu}-\left[-\sum_{i=1}^{N^{-}}x_{i}^{-}\Delta\mur\left(E_{i}^{-\prime}\right),-\sum_{i=1}^{N^{-}}x_{i}^{-}\Delta\mul\left(E_{i}^{-\prime}\right)\right]\\
 & = & \int X^{+}\rmd\bar{\mu}-\left[\sum_{i=1}^{N^{-}}\left(-x_{i}^{-}\right)\Delta\mur\left(E_{i}^{-\prime}\right),\sum_{i=1}^{N^{-}}\left(-x_{i}^{-}\right)\Delta\mul\left(E_{i}^{-\prime}\right)\right]\\
 & = & \int X^{+}\rmd\bar{\mu}-\sum_{i=1}^{N^{-}}\left(-x_{i}^{-}\right)\left[\Delta\mur\left(E_{i}^{-\prime}\right),\Delta\mul\left(E_{i}^{-\prime}\right)\right]\\
 & = & \int X^{+}\rmd\bar{\mu}-\int\left(-X^{-}\right)\rmd\bar{\mu}\,.
\end{eqnarray*}
\end{proof}

\begin{lemma}\label{lemma:Choquet-Lebesgue}\cite{Grabisch2016}When
$\mul:\events\rightarrow\left[0,1\right]$ and $\mur:\events\rightarrow\left[0,1\right]$
are real-valued probability measures, we have $\mul=\mur$ and $\int X\rmd\bar{\mu}=\left\{ \int X\rmd\mul\right\} =\left\{ \int X\rmd\mur\right\} $.\end{lemma}

\subsection{Expectation Values for $\delta$-deterministic CIVPMs}

Finally, we are going to focus on $\delta$-deterministic CIVPMs.

\begin{lemma}\label{lemma:deterministic-CIVPM}Consider a sample
space~$\Omega$ with $d$ elements, and $\delta<1/d$. Given a $\delta$-deterministic
CIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$, the following statements
are true:
\begin{enumerate}
\item If $\bar{\mu}$ is $0$-deterministic, its left-end~$\mul$ and right-end~$\mur$
are the same CRVPM, and will be simply denoted by $\pmeas:\events\rightarrow\left\{ 1,0\right\} $.
\item There is a unique element~$\omega_{0}\in\Omega$ such that $\bar{\mu}\left(\left\{ \omega_{0}\right\} \right)\subseteq\left[1-\delta,1\right]$,
denoted by $\bar{\mu}^{-1}\left(\left[1-\delta,1\right]\right)$.
In particular, if $\bar{\mu}$ is $0$-deterministic, $\bar{\mu}\left(\left\{ \omega_{0}\right\} \right)=\left\{ \pmeas\left(\left\{ \omega_{0}\right\} \right)\right\} =\necess$,
$\omega_{0}$ could be denoted by $\pmeas^{-1}\left(1\right)$ or
$\bar{\mu}^{-1}\left(\necess\right)$.
\end{enumerate}
\end{lemma}

\begin{proof}~
\begin{enumerate}
\item The direct consequence of lemma~\ref{lemma:QIVPM2QCM}.
\item We have either $\bar{\mu}\left(\left\{ \omega\right\} \right)\subseteq\left[0,\delta\right]$
or $\bar{\mu}\left(\left\{ \omega\right\} \right)\subseteq\left[1-\delta,1\right]$.
\begin{casenv}
\item If $\bar{\mu}\left(\left\{ \omega\right\} \right)\subseteq\left[0,\delta\right]$
for all $\omega\in\Omega$, then $\necess=\bar{\mu}\left(\Omega\right)\subseteq\sum_{\omega\in\Omega}\bar{\mu}\left(\left\{ \omega\right\} \right)\subseteq\left[0,d\delta\right]\Rightarrow d\delta\ge1$.
A contradiction!
\item If there are more than one~$\omega\in\Omega$ such that $\bar{\mu}\left(\left\{ \omega\right\} \right)\subseteq\left[1-\delta,1\right]$,
then $\necess=\bar{\mu}\left(\Omega\right)\subseteq\sum_{\omega\in\Omega}\bar{\mu}\left(\left\{ \omega\right\} \right)\subseteq\left[2-2\delta,d\right]\Rightarrow2\delta\ge1$.
A contradiction!
\end{casenv}
Therefore, there is a unique element~$\omega_{0}\in\Omega$ such
that $\bar{\mu}\left(\left\{ \omega_{0}\right\} \right)\subseteq\left[1-\delta,1\right]$.
\end{enumerate}
\end{proof}

\begin{lemma}\label{lemma:deterministic-CIVPM-expectation}For any
CIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\mathrm{D}}$ and
a random variable~$X=\sum_{i=1}^{N}x_{i}\mathbf{1}_{E_{i}}$, then
$\int X\rmd\bar{\mu}=\left\{ X\left(\bar{\mu}^{-1}\left(\necess\right)\right)\right\} $.\end{lemma}

\begin{proof}By adopting the notation in lemma~\ref{lemma:deterministic-CIVPM},
we have
\begin{eqnarray*}
 &  & \int X\rmd\bar{\mu}\stackrel{\textrm{Lemma }\ref{lemma:Choquet-Lebesgue}}{=}\left\{ \int X\rmd\mu\right\} \\
 & \stackrel{\textrm{Equation }(\ref{eq:expectation})}{=} & \left\{ X\left(\pmeas^{-1}\left(1\right)\right)\cdot1\right\} =\left\{ X\left(\bar{\mu}^{-1}\left(\necess\right)\right)\right\} \,.
\end{eqnarray*}
\end{proof}

Given two singleton sets~$\left\{ x_{1}\right\} $ and $\left\{ x_{2}\right\} $,
it is intuitive to define their multiplication of to be $\left\{ x_{1}x_{2}\right\} $
which gives the following lemma.

\begin{lemma}\label{lemma:deterministic-CIVPM-product}For any CIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\mathrm{D}}$
and a sequence of random variable~$\left\{ X_{j}\right\} $,
\begin{equation}
\prod_{j}\int X_{j}\rmd\bar{\mu}=\int\prod_{j}X_{j}\rmd\bar{\mu}\,.\label{eq:Kochen-Specker-Rules-classical}
\end{equation}
\end{lemma}

\begin{proof}
\begin{eqnarray*}
 &  & \prod_{j}\int X_{j}\rmd\bar{\mu}\stackrel{\textrm{Lemma }\ref{lemma:deterministic-CIVPM-expectation}}{=}\prod_{j}\left\{ X_{j}\left(\bar{\mu}^{-1}\left(\necess\right)\right)\right\} \\
 & = & \left\{ \prod_{j}X_{j}\left(\bar{\mu}^{-1}\left(\necess\right)\right)\right\} \stackrel{\textrm{Lemma }\ref{lemma:deterministic-CIVPM-expectation}}{=}\int\prod_{j}X_{j}\rmd\bar{\mu}\,.
\end{eqnarray*}
\end{proof}

\subsection{Expectation Values for Quantum IVPMs}

Before discussing quantum IVPMs, we first review various notations
brought forward from conventional quantum theory. Assume that a physical
system is represented by a density matrix~$\rho$, and that we wish
to measure a physical quantity~${\cal O}$ of this system, with corresponding
observable~$\mathbf{O}$, whose spectral decomposition is given by
$\mathbf{O}=\sum_{i=1}^{N}\lambda_{i}P_{i}$, where $P_{i}$ is the
projector onto the eigenspace of the eigenvalue~$\lambda_{i}$. Then,
the Born rule of CQT tells us that the expectation value of the observable~$\mathbf{O}$
is given by~\cite{544199,Jaeger2007}
\begin{equation}
\expval{\mathbf{O}}_{\mu_{\rho}^{\mathrm{B}}}=\Tr\left(\rho\mathbf{O}\right)=\sum_{i=1}^{N}\lambda_{i}\mu_{\rho}^{\mathrm{B}}\left(P_{i}\right)\ .\label{eq:quantum-expectation}
\end{equation}

Consider a quantum IVPM~$\bar{\mu}$ such that $\mul:\events\rightarrow\left[0,1\right]$
and $\mur:\events\rightarrow\left[0,1\right]$ are the left-end and
the right-end of $\bar{\mu}$, respectively, i.e., $\bar{\mu}\left(P\right)=\left[\mul\left(P\right),\mur\left(P\right)\right]$
for all $P\in\events$. This time, when we express the observable
as its spectral decomposition
\begin{equation}
\mathbf{O}=\sum_{s\in\left\{ +,-\right\} }\sum_{i=1}^{N^{s}}\lambda_{i}^{s}P_{i}^{s}\,,\label{eq:spectrum-decomposition}
\end{equation}
we separate the positive and negative eigenvalues and order its eigenvalues
from the smallest to the largest 
\begin{equation}
\lambda_{N^{-}}^{-}\le\ldots\le\lambda_{1}^{-}\le0\le\lambda_{1}^{+}\le\ldots\le\lambda_{N^{+}}^{+}\,.
\end{equation}
Let $P_{i}^{s\prime}=\sum_{j\ge i}P_{j}^{s}$ and $\Delta\mu^{*}\left(P_{i}^{s\prime}\right)=\mu^{*}\left(P_{i}^{s\prime}\right)-\mu^{*}\left(P_{i+1}^{s\prime}\right)$,
where $s\in\left\{ +,-\right\} $ and $\mu^{*}\in\left\{ \mul,\mur\right\} $.
We can combine the classical Choquet integral~\cite{Vitali1925,Choquet1954,GilboaSchmeidler1994,Grabisch2016}
with Eq.~(\ref{eq:quantum-expectation}) and define the expectation
value of $\mathbf{O}$ with respect to $\bar{\mu}$ as
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\sum_{s\in\left\{ +,-\right\} }\sum_{i=1}^{N^{s}}\lambda_{i}^{s}\left[\Delta\mul\left(P_{i}^{s\prime}\right),\Delta\mur\left(P_{i}^{s\prime}\right)\right]\,.\label{eq:quantum-interval-expectation}
\end{equation}

\subsection{Between Expectation Values for Classical and Quantum IVPMs\label{subsec:Between-Expectation-Values-Classical-Quantum}}

Since the similarity between classical and quantum expectation values,
we would prove some lemmas\footnote{\yutsung{functor?}} which can
help us move the classical results to quantum.

\begin{definition}Given an observable~$\mathbf{O}$ with its spectral
decomposition~(\ref{eq:spectrum-decomposition}). A subspace of events~$\events'\subseteq\events$
is called a subspace of events induced by $\mathbf{O}$ if it is the
minimal subspace of events satisfying $\left\{ P_{i}^{-}\right\} _{i=1}^{N^{-}}\cup\left\{ P_{i}^{+}\right\} _{i=1}^{N^{+}}\subseteq\events'$.
Obvious, $\events'$ is convex.\end{definition}

Notice that a sub-event space induced by any observable commutes.
By theorem~\ref{thm:sub-event-space}, we can find a set~$\Omega$,
a function~$\tau:\events'\rightarrow2^{\Omega}$ to pull quantum
real-valued or interval-valued probability measures to classical ones.
Although it is a powerful tool, it would be more powerful if we can
pull an observable to a random variable as well.

\begin{definition}\label{def:observable2random-variable}Given an
observable~$\mathbf{O}$ with its spectral decomposition~(\ref{eq:spectrum-decomposition})
and its inducing subspace of events\nb$\events'$, set~$\Omega$,
and function~$\tau:\events'\rightarrow2^{\Omega}$. Then, the random
variable~$X:\Omega\rightarrow\mathbb{R}$ defined by $X=\sum_{s\in\left\{ +,-\right\} }\sum_{i=1}^{N^{s}}\lambda_{i}^{s}\mathbf{1}_{\tau\left(P_{i}^{s}\right)}$
is called the random variable induced by $\mathbf{O}$.\end{definition}

The following lemma shows that .

\begin{lemma}\label{lemma:observable2random-variable}Given an observable~$\mathbf{O}$
with an orthonormal basis\nb$\Omega=\left\{ \ket{k}\right\} _{k=0}^{d-1}$,
$\widehat{\events}=\set{\sum_{\ket{k}\in E}\proj{k}}{E\subseteq\Omega}$,
and the bijection~$\tau:\widehat{\events}\rightarrow2^{\Omega}$
defined in theorem~\ref{thm:sub-event-space}. By definition~\ref{def:observable2random-variable},
$\mathbf{O}$ induces a random variable\nb$X:\Omega\rightarrow\mathbb{R}$,
if $\mathbf{O}=\sum_{k=0}^{d-1}\lambda_{k}\proj{k}$, then $X=\sum_{k=0}^{d-1}\lambda_{k}\mathbf{1}_{\left\{ \ket{k}\right\} }$.\end{lemma}

\begin{proof}Since $\mathbf{O}$ can be diagonalized by $\Omega$,
each projector in the spectral decomposition\nb$P_{i}^{s}$ is the
summation of some basis vectors\nb$\sum_{k\in S_{i}^{s}}\proj{k}$.
Since
\begin{equation}
\mathbf{1}_{\tau\left(P_{i}^{s}\right)}=\mathbf{1}_{\tau\left(\sum_{k\in S_{i}^{s}}\proj{k}\right)}=\mathbf{1}_{\bigcup_{k\in S_{i}^{s}}\tau\left(\proj{k}\right)}=\sum_{k\in S_{i}^{s}}\mathbf{1}_{\left\{ \ket{k}\right\} }\,,
\end{equation}
we have $X=\sum_{k=0}^{d-1}\lambda_{k}\mathbf{1}_{\ket{k}}$.\end{proof}

\begin{lemma}\label{lemma:real-expectation-value-between-classical-quantum}Adopting
the definitions in this subsection. Given a QRVPM~$\mu:\events\rightarrow\left[0,1\right]$,
there is a classical real-valued probability measure~$\mu':2^{\Omega}\rightarrow\left[0,1\right]$
such that $\mu|_{\events'}=\mu'\circ\tau$. Then, we have $\expval{\mathbf{O}}_{\mu}=\int X\rmd\mu'$.\end{lemma}

\begin{proof}Since $\mu:\events\rightarrow\left[0,1\right]$ is a
QRVPM, $\mu|_{\events'}$ is a QRVPM on the commuting sub-event space\nb$\events'$.
By case\nb\ref{enu:sub-event-space-real-quantum2classical} in theorem~\ref{thm:sub-event-space},
there is a CRVPM such that $\mu|_{\events'}=\mu'\circ\tau$. 
\begin{eqnarray*}
\expval{\mathbf{O}}_{\mu} & \stackrel{\textrm{Equation }(\ref{eq:quantum-expectation})}{=} & \sum_{s\in\left\{ +,-\right\} }\sum_{i=1}^{N^{s}}\lambda_{i}^{s}\mu\left(P_{i}^{s}\right)\\
 & = & \sum_{s\in\left\{ +,-\right\} }\sum_{i=1}^{N^{s}}\lambda_{i}^{s}\mu'\left(\tau\left(P_{i}^{s}\right)\right)\\
 & \stackrel{\textrm{Equation }(\ref{eq:expectation})}{=} & \int X\rmd\mu'\,.
\end{eqnarray*}
\end{proof}

\begin{lemma}\label{lemma:interval-expectation-value-between-classical-quantum}Adopting
the definitions in this subsection. Given a quantum IVPM~$\bar{\mu}$,
there is a classical IVPM~$\bar{\mu}':2^{\Omega}\rightarrow\mathscr{I}$
such that $\bar{\mu}=\bar{\mu}'\circ\tau$. Then, we have $\expval{\mathbf{O}}_{\bar{\mu}}=\int X\rmd\bar{\mu}'$.\end{lemma}

\begin{proof}Let $\mul[\prime]:\events'\rightarrow\left[0,1\right]$
and $\mur[\prime]:\events'\rightarrow\left[0,1\right]$ are the left-end
and the right-end of $\bar{\mu}'$, respectively, i.e., $\bar{\mu}'\left(E\right)=\left[\mul[\prime]\left(E\right),\mur[\prime]\left(E\right)\right]$
for all $E\in2^{\Omega}$. Then, 
\begin{eqnarray*}
\expval{\mathbf{O}}_{\bar{\mu}} & \stackrel{\textrm{Equation }(\ref{eq:quantum-interval-expectation})}{=} & \sum_{s\in\left\{ +,-\right\} }\sum_{i=1}^{N^{s}}\lambda_{i}^{s}\left[\Delta\mul\left(P_{i}^{s\prime}\right),\Delta\mur\left(P_{i}^{s\prime}\right)\right]\\
 & = & \sum_{s\in\left\{ +,-\right\} }\sum_{i=1}^{N^{s}}\lambda_{i}^{s}\left[\Delta\mul[\prime]\left(\tau\left(P_{i}^{s\prime}\right)\right),\Delta\mur[\prime]\left(\tau\left(P_{i}^{s\prime}\right)\right)\right]\\
 & \stackrel{\textrm{Definition }\ref{def:Choquet}}{=} & \int X\rmd\bar{\mu}'\,.
\end{eqnarray*}
\end{proof}

\subsection{Between Expectation Values for Quantum Real-valued and Interval-valued
Probability Measures}

Intuitively, the expectation value of an observable with respect to
an IVPM should be the minimum and maximum expectation value of the
same observable with respect to its consistent states. We found this
is true when we require our IVPM to be convex, and on a commuting
sub-event space.

\begin{thm}\label{thm:quantum-Rosenmuller}Given a Hilbert space,
for every quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
and any observable~$\mathbf{O}$ inducing a sub-event space~$\events'$,
if $\bar{\mu}$ is convex, then
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\left[\min_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\mu_{\rho}^{\mathrm{B}}},\max_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\mu_{\rho}^{\mathrm{B}}}\right]\,.\label{eq:quantum-interval-expectation-core}
\end{equation}
Conversely, if Eq.~(\ref{eq:quantum-interval-expectation-core})
holds for every observable, then $\bar{\mu}$ is convex.\end{thm}

\begin{proof}Adopting the definitions in subsection~\ref{subsec:Between-Expectation-Values-Classical-Quantum},
it is easy to prove
\begin{eqnarray*}
\expval{\mathbf{O}}_{\bar{\mu}} & \stackrel{\textrm{Lemma }\ref{lemma:real-expectation-value-between-classical-quantum}}{=} & \int X\rmd\bar{\mu}'\\
 & \stackrel{\textrm{Theorem }\ref{thm:Rosenmuller}}{=} & \left[\min_{\mu'\in\mathrm{core}\left(\bar{\mu}'\right)}\int X\rmd\mu',\max_{\mu'\in\mathrm{core}\left(\bar{\mu}'\right)}\int X\rmd\mu'\right]\\
 & \stackrel{\textrm{Lemma }\ref{lemma:interval-expectation-value-between-classical-quantum}}{=} & \left[\min_{\mu'\in\mathrm{core}\left(\bar{\mu}'\right)}\expval{\mathbf{O}}_{\mu_{\rho}^{\mathrm{B}}},\min_{\mu'\in\mathrm{core}\left(\bar{\mu}'\right)}\expval{\mathbf{O}}_{\mu_{\rho}^{\mathrm{B}}}\right]\,.
\end{eqnarray*}
Therefore, equation~(\ref{eq:quantum-interval-expectation-core})
is equivalent to convexity if $\mathrm{core}\left(\bar{\mu}'\right)$
and $\coreBorn\left(\bar{\mu},\events'\right)$ are one-to-one correspondence
so that
\begin{equation}
\left[\min_{\mu'\in\mathrm{core}\left(\bar{\mu}'\right)}\expval{\mathbf{O}}_{\mu_{\rho}^{\mathrm{B}}},\min_{\mu'\in\mathrm{core}\left(\bar{\mu}'\right)}\expval{\mathbf{O}}_{\mu_{\rho}^{\mathrm{B}}}\right]=\left[\min_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\mu_{\rho}^{\mathrm{B}}},\max_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\mu_{\rho}^{\mathrm{B}}}\right]\,.
\end{equation}

Notice that lemma~\ref{lemma:real-expectation-value-between-classical-quantum}
guarantees that for any $\rho\in\coreBorn\left(\bar{\mu},\events'\right)$,
there exists $\mu'\in\mathrm{core}\left(\bar{\mu}'\right)$ such that
$\mu_{\rho}^{\mathrm{B}}=\mu'\circ\tau$. On the other hand, for any
$\mu'\in\mathrm{core}\left(\bar{\mu}'\right)$, we have
\begin{eqnarray}
 &  & \mu'\left(E\right)\in\bar{\mu}'\left(E\right)\textrm{ for all }E\in2^{\Omega}\\
 & \Rightarrow & \mu'\left(\tau\left(P\right)\right)\in\bar{\mu}'\left(\tau\left(P\right)\right)=\bar{\mu}\left(P\right)\textrm{ for all }P\in\events'
\end{eqnarray}
By theorem~\ref{thm:sub-event-space}, there is a density matrix~$\rho$
such that $\mu_{\rho}^{\mathrm{B}}\left(P\right)=\mu'\left(\tau\left(P\right)\right)$
for all $P\in\events'$. Therefore, $\mu_{\rho}^{\mathrm{B}}\in\mathrm{core}\left(\bar{\mu},\events'\right)$,
i.e., $\rho\in\coreBorn\left(\bar{\mu},\events'\right)$.\end{proof}

Our next lemma discusses the quantum version of lemma~\ref{lemma:Choquet-Lebesgue}.

\begin{lemma}\label{lemma:expectation-singleton}Given a Hilbert
space of $d\ge3$, when $\mul:\events\rightarrow\left[0,1\right]$
and $\mur:\events\rightarrow\left[0,1\right]$ are real-valued probability
measures, we have $\mul=\mur=\mu$. Then, $\expval{\mathbf{O}}_{\bar{\mu}}=\left\{ \expval{\mathbf{O}}_{\mu}\right\} $.\end{lemma}

\begin{proof}[Proof 1]Adopting the definitions in subsection~\ref{subsec:Between-Expectation-Values-Classical-Quantum},
we have
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}\stackrel{\textrm{Lemma }\ref{lemma:real-expectation-value-between-classical-quantum}}{=}\int X\rmd\bar{\mu}'\stackrel{\textrm{Lemma }\ref{lemma:Choquet-Lebesgue}}{=}\left\{ \int X\rmd\mu'\right\} \stackrel{\textrm{Lemma }\ref{lemma:interval-expectation-value-between-classical-quantum}}{=}\left\{ \expval{\mathbf{O}}_{\mu}\right\} \,,
\end{equation}
where $\mu'=\mul[\prime]=\mur[\prime]$.\end{proof}

\begin{proof}[Proof 2]Let $\events'$ be the subspace of events
induced by $\mathbf{O}$. For any state\nb$\rho\in\coreBorn\left(\bar{\mu},\events'\right)$,
we have $\mu_{\rho}^{\mathrm{B}}\left(P\right)\in\bar{\mu}\left(P\right)=\left\{ \mu\left(P\right)\right\} $
for all $P\in\events'$. Thus, $\expval{\mathbf{O}}_{\mu_{\rho}^{\mathrm{B}}}=\expval{\mathbf{O}}_{\mu}$.
By theorem~\ref{thm:quantum-Rosenmuller}, we have
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\left[\min_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\mu_{\rho}^{\mathrm{B}}},\max_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\mu_{\rho}^{\mathrm{B}}}\right]=\left\{ \expval{\mathbf{O}}_{\mu}\right\} 
\end{equation}
\end{proof}

\subsection{Expectation Values for $\delta$-deterministic QIVPMs\label{subsec:Quantum-I_D-valued-Probability}}

After we defined the expectation value, we can provide another proof
of corollary~\ref{cor:Kochen-Specker-IVPM} which doesn't refer the
Kochen-Specker theorem directly. Assume there is a $\delta$-deterministic
QIVPM~$\bar{\mu}$, we will prove $\bar{\mu}$ has such bizarre properties
that it cannot exist for small $\delta$.

\begin{lemma}\label{lemma:deterministic-QIVPM}Given a Hilbert space
of dimension~$d$, $\delta<1/d$, and a $\delta$-deterministic QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$,
the following statements are true:
\begin{enumerate}
\item If $\bar{\mu}$ is $0$-deterministic, its left-end~$\mul$ and right-end~$\mur$
are the same QRVPM, and will be simply denoted by $\pmeas:\events\rightarrow\left\{ 1,0\right\} $.
\item Given an orthonormal basis~$\Omega$, there is a unique element~$\ket{\psi_{i_{0}}}\in\Omega$
such that $\bar{\mu}\left(\proj{\psi_{i_{0}}}\right)\subseteq\left[1-\delta,1\right]$,
denoted by $\bar{\mu}^{-1}\left(\left[1-\delta,1\right]\right)\cap\Omega$.
In particular, if $\bar{\mu}$ is $0$-deterministic, $\bar{\mu}\left(\proj{\psi_{i_{0}}}\right)=\left\{ \pmeas\left(\proj{\psi_{i_{0}}}\right)\right\} =\necess$,
$\proj{\psi_{i_{0}}}$ could be denoted by $\pmeas^{-1}\left(1\right)\cap\Omega$
or $\bar{\mu}^{-1}\left(\necess\right)\cap\Omega$.
\end{enumerate}
\end{lemma}

\begin{proof}~
\begin{enumerate}
\item The direct consequence of lemma~\ref{lemma:QIVPM2QCM}.
\end{enumerate}
\end{proof}

\begin{lemma}For any quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\mathrm{D}}$,
$\expval{\mathbf{O}}_{\bar{\mu}}$ is a singleton set~$\left\{ \lambda\right\} $,
where $\lambda$ is an eigenvalue of $\mathbf{O}$.\end{lemma}

\begin{proof}[Proof 1]Adopt the definitions in subsection~\ref{subsec:Between-Expectation-Values-Classical-Quantum}.
By lemma~\ref{lemma:QIVPM2QCM}, $\mul$ and $\mur$ are the same
real-valued probability measures, and is simply denoted by $\mu:\events\rightarrow\left\{ 1,0\right\} $.
By lemma~\ref{lemma:expectation-singleton}, $\expval{\mathbf{O}}_{\bar{\mu}}=\left\{ \expval{\mathbf{O}}_{\mu}\right\} $.
Since $\mu$ is a cryptodeterministic measure, there is only one pair
of $\left(s_{0},i_{0}\right)$ such that $\mu\left(P_{i_{0}}^{s_{0}}\right)=1$.
Therefore, 
\begin{equation}
\expval{\mathbf{O}}_{\mu}\stackrel{\textrm{Equation }(\ref{eq:quantum-expectation})}{=}\sum_{s\in\left\{ +,-\right\} }\sum_{i=1}^{N^{s}}\lambda_{i}^{s}\mu\left(P_{i}^{s}\right)=\lambda_{i_{0}}^{s_{0}}\,.\label{eq:expectation-value-eigenvalue}
\end{equation}
\end{proof}

\begin{proof}[Proof 2]Adopt the definitions in subsection~\ref{subsec:Between-Expectation-Values-Classical-Quantum},
we have
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\int X\rmd\bar{\mu}'\stackrel{\textrm{Lemma }\ref{lemma:deterministic-CIVPM-expectation}}{=}\left\{ X\left(\bar{\mu}^{-1}\left(\necess\right)\right)\right\} \,,
\end{equation}
where $X\left(\bar{\mu}^{-1}\left(\necess\right)\right)$ is an eigenvalue
of $\mathbf{O}$ by the definition of $X$.\end{proof}

Given two singleton sets~$\left\{ \lambda_{1}\right\} $ and $\left\{ \lambda_{2}\right\} $,
it is intuitive to define their multiplication of to be $\left\{ \lambda_{1}\lambda_{2}\right\} $
which gives the following lemma.

\begin{lemma}For a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\mathrm{D}}$
and a sequence of commuting observables~$\left\{ \mathbf{O}_{j}\right\} $,
\begin{equation}
\prod_{j}\expval{\mathbf{O}_{j}}_{\bar{\mu}}=\expval{\prod_{j}\mathbf{O}_{j}}_{\bar{\mu}}\,.\label{eq:Kochen-Specker-Rules}
\end{equation}
\end{lemma}

\begin{proof}[Proof 1]Since $\left\{ \mathbf{O}_{j}\right\} $ is
a sequence of commuting observables, they can be diagonalized by a
common orthonormal basis\nb$\Omega=\left\{ \ket{k}\right\} _{k=0}^{d-1}$.
By theorem~\ref{thm:sub-event-space}, there is a commuting sub-event
space\nb$\widehat{\events}$ with $\widehat{\events}\subseteq\events$
and a bijection\nb$\tau:\widehat{\events}\rightarrow2^{\Omega}$
such that there is a CIVPM\nb$\bar{\mu}:2^{\Omega}\rightarrow\mathscr{I}_{\mathrm{D}}$
such that $\bar{\mu}\left(P\right)=\bar{\mu}'\left(\tau\left(P\right)\right)$
for all $P\in\widehat{\events}$.

Let $\events'_{j}$ be the subspace of events induced by $\mathbf{O}_{j}=\sum_{k=0}^{d-1}\lambda_{j,k}\proj{k}$.
Clearly, $\events'_{j}\subseteq\widehat{\events}$. Therefore, restricting
$\tau$ on each $\events'_{j}$ gives a function\nb$\tau|_{\events'_{j}}:\events'_{j}\rightarrow2^{\Omega}$.
Then, we can follow lemma~\ref{lemma:observable2random-variable}
to induce random variables\nb$X_{j}=\sum_{k=0}^{d-1}\lambda_{j,k}\mathbf{1}_{\left\{ \ket{k}\right\} }$,
and have $\expval{\mathbf{O}_{j}}_{\bar{\mu}}=\int X_{j}\rmd\bar{\mu}'$
by lemma~\ref{lemma:interval-expectation-value-between-classical-quantum}.
Therefore, 
\begin{equation}\eqalign{ 
 & \prod_{j}\expval{\mathbf{O}_{j}}_{\bar{\mu}}=\prod_{j}\int X_{j}\rmd\bar{\mu}'\stackrel{\textrm{Lemma }\ref{lemma:deterministic-CIVPM-product}}{=}\int\prod_{j}X_{j}\rmd\bar{\mu}'\\
= & \int\prod_{j}\sum_{k=0}^{d-1}\lambda_{j,k}\mathbf{1}_{\left\{ \ket{k}\right\} }\rmd\bar{\mu}'=\int\sum_{k=0}^{d-1}\prod_{j}\lambda_{j,k}\mathbf{1}_{\left\{ \ket{k}\right\} }\rmd\bar{\mu}'\\
= & \expval{\sum_{k=0}^{d-1}\prod_{j}\lambda_{j,k}\proj{k}}_{\bar{\mu}}=\expval{\prod_{j}\mathbf{O}_{j}}_{\bar{\mu}}\,. 
}\end{equation}
\end{proof}

\begin{proof}[Proof 2]Since $\left\{ \mathbf{O}_{j}\right\} $ is
a sequence of commuting observables, they can be diagonalized by a
common orthonormal basis\nb$\Omega=\left\{ \ket{k}\right\} _{k=0}^{d-1}$,
i.e., $\mathbf{O}_{j}=\sum_{k=0}^{d-1}\lambda_{j,k}\proj{k}$ and
$\prod_{j}\mathbf{O}_{j}=\sum_{k=0}^{d-1}\prod_{j}\lambda_{j,k}\proj{k}$.
By lemma~\ref{lemma:QIVPM2QCM}, $\mul$ and $\mur$ are the same
real-valued probability measures, and is simply denoted by $\mu:\events\rightarrow\left\{ 1,0\right\} $.
Since $\mu$ is a cryptodeterministic measure, there is only one $\proj{k_{0}}$
such that $\mu\left(\proj{k_{0}}\right)=1$.
\begin{equation}\eqalign{ 
 & \prod_{j}\expval{\mathbf{O}_{j}}_{\bar{\mu}}\stackrel{\textrm{Lemma }\ref{lemma:expectation-singleton}}{=}\prod_{j}\left\{ \expval{\mathbf{O}_{j}}_{\mu}\right\} =\prod_{j}\left\{ \lambda_{j,k_{0}}\right\} \\
= & \left\{ \prod_{j}\lambda_{j,k_{0}}\right\} =\expval{\prod_{j}\mathbf{O}_{j}}_{\mu}\stackrel{\textrm{Lemma }\ref{lemma:expectation-singleton}}{=}\expval{\prod_{j}\mathbf{O}_{j}}_{\bar{\mu}}\,. 
}\end{equation}
\end{proof}

Based on the previous lemmas, we could follow Mermin's idea to prove
the Kochen-Specker theorem again~\cite{Mermin1990Simple,peres1995quantum}.

\begin{thm}\label{thm:Kochen-Specker-expectation}Given a Hilbert
space of dimension~$d\ge3$, there is no quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\mathrm{D}}$,
where $\mathscr{I}_{\mathrm{D}}=\left\{ \imposs,\necess\right\} $
as before.\end{thm}

\begin{proof}Since the expectation value~$\expval{\cdot}_{\bar{\mu}}$
satisfying the previous lemmas, the following observables~\cite{Mermin1990Simple,peres1995quantum}:
\begin{equation}
\begin{array}{ccc}
\mathbb{1}\otimes\sigma_{z} & \sigma_{z}\otimes\mathbb{1} & \sigma_{z}\otimes\sigma_{z}\\
\sigma_{x}\otimes\mathbb{1} & \mathbb{1}\otimes\sigma_{x} & \sigma_{x}\otimes\sigma_{x}\\
\sigma_{x}\otimes\sigma_{z} & \sigma_{z}\otimes\sigma_{x} & \sigma_{y}\otimes\sigma_{y}
\end{array}\label{eq:MerminSquare}
\end{equation}
must have the the expectation value~$1$ or $-1$. In each row and
column, operators commutes, and each operator is the product of the
two others, except in the third column $\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$.
Therefore, it is clear impossible to find a IVPM~$\bar{\mu}$ such
that 
\begin{equation}
\begin{array}{ccc}
\expval{\mathbb{1}\otimes\sigma_{z}}_{\bar{\mu}} & \expval{\sigma_{z}\otimes\mathbb{1}}_{\bar{\mu}} & \expval{\sigma_{z}\otimes\sigma_{z}}_{\bar{\mu}}\\
\expval{\sigma_{x}\otimes\mathbb{1}}_{\bar{\mu}} & \expval{\mathbb{1}\otimes\sigma_{x}}_{\bar{\mu}} & \expval{\sigma_{x}\otimes\sigma_{x}}_{\bar{\mu}}\\
\expval{\sigma_{x}\otimes\sigma_{z}}_{\bar{\mu}} & \expval{\sigma_{z}\otimes\sigma_{x}}_{\bar{\mu}} & \expval{\sigma_{y}\otimes\sigma_{y}}_{\bar{\mu}}
\end{array}\label{eq:MerminSquare-values}
\end{equation}
satisfying equation~(\ref{eq:Kochen-Specker-Rules}).\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{comment}
 \bibliographystyle{plain}
\bibliography{prop}
\end{comment}
\printbibliography
\end{document}